### 2014-08-21 manuscript.Rmd
init commit

---
layout: preprint
title: "An introduction to Docker for reproducible research and collaboration"
author: 
  - name: Carl Boettiger
    affiliation: cstar
    email: cboettig(at)gmail.com
    footnote: Corresponding author
address: 
  - code: cstar
    address: | 
      Center for Stock Assessment Research,
      110 Shaffer Rd,
      Santa Cruz, CA 95050, USA 
abstract: | 
          Write the abstract of the paper here

bibliography: components/references.bib
csl: components/ecology.csl
documentclass: components/elsarticle

output: 
  pdf_document:
    template: components/elsarticle.latex
    keep_tex: true
    fig_caption: true
---

```{r caching, include=FALSE}
library("methods")
library("knitr")
basename <- "manuscript"
opts_chunk$set(fig.path = paste("components/figure/", basename, "-", sep=""),
               cache.path = paste("components/cache/", basename, "/", sep=""))
opts_chunk$set(cache = 2)
opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, 
               comment = NA, verbose = TRUE, echo=FALSE)

```

```{r libraries, include=FALSE, cache = FALSE}
library("rmarkdown")
```

Wouldn't it be nice if you could just revert your computer to how it was an hour ago or a day ago when everything was still working?  Wouldn't it be nice if your collaborators could be running an identical computer environment to your own?

# Reproducible computational research and the challenge of software environments

- Reproducible research is an important issue
- Computational research should in principle be the least difficult to make completely reproducible, but in practice this is not the case (see this entire special issue). 
- Computational research relies on an ever more complex set of software and dependencies for that software
- These dependencies frequently frustrate reproducibility (see Titus study on failing to install xx% of research software, etc, NESCENT reproducibility study, etc)

# Typical solution approaches

- Documentation
- Standard environments
- Virtual machines & Cloud computing

# Challenges with common approaches

# Emerging approaches: DevOpts

- DevOpts: Scripting your environment
- Vagrant

# Docker

- Introduction and hype

- Technologies from operating systems research: LXC containers, virtualization of OS, git-like verioning

# Using Docker for Reproducible Research

- Example:

```{r}
# Read in an external data file 
# Load some libraries with external dependencies
# Compute some statistics, plot 
```

- Interactive use, scripted use (Dockerfiles)
- Reverting to a previous state
- Integration with local development environment
- Integration with cloud computing
- Transparent & extensible
- serial and linking containers

- Best practices are still taking shape

# Walking the talk

- This paper is a dynamic document with code integrated into the text.  
- The computational environment of this paper is provided as a version controlled Docker image

### 2014-09-09 provision.sh
update notes

@@ -35,7 +35,7 @@ docker run --name='drone' -d -p 8080:80 --privileged mattgruter/drone
 
 
 ## Deploy Gitlab on 10080
-docker run --name='gitlab' -it --rm \
+docker run --name='gitlab' -d \
   -p 10022:22 -p 10080:80 \
   -e 'GITLAB_PORT=10080' -e 'GITLAB_SSH_PORT=10022' \
   sameersbn/gitlab:7.2.1-1

### 2014-09-10 manuscript.Rmd
writing

@@
 abstract: |
-          Write the abstract of the paper here
+  Academic research
 
@@
 
-```{r caching, include=FALSE}
-library("methods")
-library("knitr")
-basename <- "manuscript"
-opts_chunk$set(fig.path = paste("components/figure/", basename, "-", sep=""),
-               cache.path = paste("components/cache/", basename, "/", sep=""))
-opts_chunk$set(cache = 2)
-opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, 
-               comment = NA, verbose = TRUE, echo=FALSE)
+<!--
 
-```
+Suggested paper topics include, but are not limited to:
 
-```{r libraries, include=FALSE, cache = FALSE}
-library("rmarkdown")
-```
+  * novel ways to manage systems research artifacts, dependencies, workflows,
+    and data collection
+  * testbeds for repeatable research and/or publishing executable artifacts
+  * "rich publications" that connect papers to software and data
+  * repeatability in distributed and nondeterministic systems
+  * mitigating obstacles to artifact sharing, repeatability, and reproduction
+  * repeatability challenges specific to systems
+  * measurement bias in systems experiments
+  * experiences and outcomes from systems repetition and reproduction studies
+  * improving scientific practice in systems research
+  * incentives for producing and sharing high-quality research artifacts
+
+Papers are expected to present significant results, insights, and/or directions
+for future work, and they are expected to include substantial material that has
+not already been published.
+
+SUBMISSION INSTRUCTIONS
+
+Submissions must be no longer than ten (10) pages; shorter papers are allowed
+and encouraged.  Submissions must be written in English and follow the standard
+formatting guidelines for papers appearing in Operating Systems Review (see
+<http://www.sigops.org/osr.html>).  Papers must be submitted in PDF format via
+the submission web site.  Submitted papers will be reviewed by the guest editor
+and the review committee.  Papers will be evaluated on technical quality,
+originality, relevance, and presentation.  Accepted papers will be published in
+the January 2015 issue of Operating Systems Review.
+
+-->
+
+Reproducible research, and in particular, computational reproducibility,
+has recieved an increasing level of attention throughout the scientific
+community as all steps of the scientific process, from data collection
+and processing, to analyses, visualizations and conclusions depend ever
+more on computation and algorithms.  Though in principle this algorithmic
+dependence should make such research easier to reproduce -- computer codes
+being both more portable and potentially more precise to exchange and run
+than experimental methods -- in practice this has led to an ever larger
+and more complex black box that stands between what was actually done
+and what is described in the literature.  Crucial scientific processes
+such as replicating the results, extending the approach or testing the
+conclusions in other contexts, or even merely installing the software
+used by the original researchers can become immensely time-consuming if
+not impossible.
+
+At the heart of the computational reproducibility challenge is the emmense, 
+complex, diverse, and rapidly changing nature of the computer software
+
+It is possible, and indeed often desirable, to tackle this problem
+at a much higher level than that of the operating system itself.
+For instance, it is often sufficient to address these concerns at the
+level of a particular scripting language rather than the level of the
+operating system. Most major scripting languages have dedicated tools for
+handling dependencies and versioning (e.g. Ruby gems, R packages, Python
+eggs, perl modules, javascript node package manager); with some languages
+having the fortune or misfortune to have multiple systems. Though systems
+differ somewhat in their strategies and their effectiveness, all are
+generally vulnerable to changes in their lower-level dependencies (such as
+compilers, C libraries, and so forth) that may differ between platforms.
+More pointedly, researchers whose work is not easily isolated to a single
+such environment but combines tools from different systems (including
+combining those tools used for formatting & visual display, such as
+LaTeX or pandoc, with scripting or analysis tools such as R). Moreover,
+addressing reproducibility at the level of the operating system avoids
+constructing approaches unique to a specific language or tool.
+
+Two dominant paradigms have emerged to address these issues so far:
+workflow software and cloud computing on virtual machines.  Here I
+consider how a popular emerging technology, Docker, based on Linux
+containers, offers a third route that addresses pratical and theoretical
+concerns for reproducible research not addressed by either of these
+paradigms.
+
+It is worth observing from the outset that the primary barrier to
+computational reproducibility in many domain sciences has nothing
+to do with the technology discussed here, but stems rather from a
+reluctance to publish the code used to generating the results in the
+first place [@Barnes2010]. Despite extensive evidence to the contrary
+[e.g. @Ince2012], many researchers and journals continue to assume that
+summary descriptions or pseudo-code provide a sufficient description of
+computational methods used in data gathering, processing, simulation,
+visualization, or analysis.  Until such code is available in the first
+place, we cannot even begin to encounter the problems that the approaches
+discussed here set out to solve. As a result, few domain researchers
+may be fully aware of the challenges involved in effectively re-using
+published code.
+
+Yet this coin has two sides: By lowering the barriers
+these barrier to re-use through the appropriate infrastructure first, the research
+community as a whole may be better positioned to take advantage of code 
+sharing practices when they are adopted.  This has cultural implications as well:
+many researchers may be reluctant to publish code today because they fear
+a primarily one-way street: others will have the technical savvy to benefit
+from their hard work, while they may not benefit from the work produced by others.
+Lowering the technical barriers to reuse provides immediate practical benefits
+that make this exchanged into a more balanced, two-way street.
+
+The issue most domain researchers do not recognize as an issue.  If computational reprod
+
+Before we begin, it is important to distinguish between the nature of the problem we seek to address here and other elements of computationally reproducible research.
+While this approach may address some of the same challenges as workflow software, dynamic documents / executable papers and ad hoc packaging systems, it can always been seen to supplement rather than replace such solutions.
+A good example comes from dynamic documents, which seek to embed the code required to re-generate the results within the manuscript itself.
+
+http://doi.ieeecomputersociety.org/10.1109/MCSE.2012.62
 
@@
 
-# Walking the talk
+My original motivation was more on the research side, primarily from NESCent's reproducibility study and Rich's reproducibility ropensci blog post, where you might recall Docker being discussed in the comments.
 
-- This paper is a dynamic document with code integrated into the text.  
-- The computational environment of this paper is provided as a version controlled Docker image
+Hilmar highlighted 3 reasons why reproducibility efforts fail even for papers that seem to do the right things: 1) dependency hell frustrating installation, 2) small holes in documentation can be major barriers for "novices" (even if novices are experts in nearby fields and nearby languages), 3) software dependency evolution breaks existing code (code rot).
+
+Rich highlighted about 4 reasons: 1) dependency hell / incomplete documentation on versions, 2) all the scripts & software for installing managing the 'reproducibility' part of their analysis (travis, etc) were as complex as the analysis itself.
+(more than half the software dependencies were things like RCurl,  knitr & packrat etc).
+3) Running the analysis on Travis CI means it has to finish in 50 minutes, 4) the learning curve for Github + Travis + etc is prohibitive.
+
+Docker provides a "DevOpts" approach, replacing user documentation on how to install with scripts that handle provisioning.
+This pretty much solves Hilmar's #1 & #2.
+You can maintain a timestamped binary of the Docker image to address Hilmar's #3.
+
+Rich's #2 is largely handled by providing an image where this stuff is all set up for you, so it building on the ropensci docker image, you need to add only what is necessary for your actual analysis.
+His #3 is addressed bc instead of running on Travis, you can run this locally (on Windows/Mac/Linux) or on a cloud of your choice with as much horsepower as you need.
+I think this lowers the learning curve (#4) some, (for instance, you can use git in the usual way but you can also fully interact without knowing git).
+
+My hope is, that instead of the rather giant set-up Rich et al needed, a user could just have their paper as say, an R script or rmarkdown file, develop & run it locally in the docker-ropensci environment, and then anyone having trouble getting it to run in their native R should at least be able to run it in the docker-ropensci environment.
+Docker is super-easy to extend: they could make their own minimal Dockerfile like this: https://github.com/ropensci/RNeXML/blob/devel/manuscripts/Dockerfile There's a few other cool things about Docker (The git-like way it handles images etc) too.
+
+# Acknowledgements
+
+- NSF
+- rOpenSci
+- Dirk Eddlebuettel, Rich FitzJohn, Yihui Xie, Titus Brown, John Stanton-Geddes and others for helpful discussions about reproducibility, virtualization, and Docker that have helped shape this manuscript.

### 2014-09-12 manuscript.Rmd
writing away

@@
 
-Reproducible research, and in particular, computational reproducibility,
-has recieved an increasing level of attention throughout the scientific
-community as all steps of the scientific process, from data collection
-and processing, to analyses, visualizations and conclusions depend ever
-more on computation and algorithms.  Though in principle this algorithmic
-dependence should make such research easier to reproduce -- computer codes
-being both more portable and potentially more precise to exchange and run
+Reproducible research has recieved an increasing level of attention
+throughout the scientific community [The Economist 2014]. A all steps
+of the scientific process, from data collection and processing, to
+analyses, visualizations and conclusions depend ever more on computation
+and algorithms, _computational reproducibility_ has recieved particular
+attention [Merali 2010].  Though in principle this algorithmic dependence
+should make such research easier to reproduce -- computer codes being
+both more portable and potentially more precise to exchange and run
 than experimental methods -- in practice this has led to an ever larger
@@
 
-At the heart of the computational reproducibility challenge is the emmense, 
-complex, diverse, and rapidly changing nature of the computer software
+## A technical and cultural problem
+
+It is worth observing from the outset that the primary barrier to
+computational reproducibility in many domain sciences has nothing
+to do with the technological approaches discussed here, but stems rather from a
+reluctance to publish the code used to generating the results in the
+first place ([Barnes 2010]). Despite extensive evidence to the contrary
+(e.g. [Ince 2012]), many researchers and journals continue to assume that
+summary descriptions or pseudo-code provide a sufficient description of
+computational methods used in data gathering, processing, simulation,
+visualization, or analysis.  Until such code is available in the first
+place, we cannot even begin to encounter the problems that the approaches
+discussed here set out to solve. As a result, few domain researchers
+may be fully aware of the challenges involved in effectively re-using
+published code.
+
+A lack of requirements or incentives no doubt plays a crucial role
+in discouraging sharing ([Barnes 2010]). Nevertheless, it is easy to
+underestimate the siginficant barriers raised by a lack of familiar,
+intuitive, and widely adopted tools for addressing the challenges of
+computational reproducibility. Surveys and case studies find that a lack
+of time, more than innate opposition to sharing, discourages researchers
+from providing code ([Stodden 2010], [FitzJohn 2014]).
+
+<!-- Time savings. Personal reproducibility. Collaboration. Teaching. -->
+
+<!-- Figure 1: technical reproducibility stack
+1. Operating system
+2. Workflow systems 
+3. Package managers 
+4. Dynamic documents
+
+-->
+
+<!-- Operating system - level approach -->
+At the heart of the computational reproducibility challenge is the 
+complexity, diversity, and rapidly changing nature of the computer software
+that underpin such work.
+
+<!-- Current approaches at the Operating Systems level -->
+
+Two dominant paradigms have emerged to address these issues so far:
+workflow software (e.g. [Atintas 2004], [Hull 2006])  and
+cloud computing on virtual machines (e.g. [Dudley 2013], [Howe 2012]). 
+Workflow software provides very elegant technical solutions to the
+challenges of communication between diverse software tools, capturing
+provenance in graphically driven interfaces, and handling issues from of
+versioning dependencies to data access.  Workflow solutions are often 
+built by well-funded collaborations between domain scientists and computer
+scientists, and can be very successful in the communities within which 
+they recieve substatianal adoption. Nonetheless, most workflow systems 
+struggle with relatively low total adoption overall ([Gin 2007, Dudley 2013]).
+
+[Dudley 2013] gives several reasons for this:
+
+> (i) efforts are not rewarded by the current academic research and
+funding environment; (ii) commercial software vendors tend to protect
+their markets through proprietary formats and interfaces; (iii)
+investigators naturally tend to want to ‘own’ and control their
+research tools; (iv) even the most generalized software will not be
+able to meet the needs of every researcher in a field; and finally (v)
+the need to derive and publish results as quickly as possible precludes
+the often slower standards-based development path.
+
+In short, workflow software expects a new approach to computational research
+
+In contrast, virtual machines (VMs) offer a far more brute-force approach.  If the goal
+is to reproduce everything as it ran on the researcher's own computer, well 
+then we can just take a snapshot of the entire computer: data, software, 
+operating system and all.
+
+To make this practical, [Dudley 2013] and [Howe 2012]
+propose using virtual machine images that will run on the cloud, such as Amazon's
+EC2 system, which is already based upon this kind of virtualization.
+
+Critics of the use of VMs to support reproducibility highlight that the 
+approach is too much of a black box and ill suited for
+
+Here I consider how a popular emerging technology, Docker, based on
+Linux containers, offers a third route that addresses pratical and
+theoretical concerns for reproducible research not addressed by either
+of these paradigms.
+
+Wouldn't it be nice if you could just revert your computer to how it was an hour ago or a day ago when everything was still working?
+Wouldn't it be nice if your collaborators could be running an identical computer environment to your own?
+
+# Reproducible computational research and the challenge of software environments
+
+- Reproducible research is an important issue
+- Computational research should in principle be the least difficult to make completely reproducible, but in practice this is not the case (see this entire special issue). 
+- Computational research relies on an ever more complex set of software and dependencies for that software
+- These dependencies frequently frustrate reproducibility (see Titus study on failing to install xx% of research software, etc, NESCENT reproducibility study, etc)
+
+## Alternatives and other considerations
+
+<!-- Package managers -->
 
 It is possible, and indeed often desirable, to tackle this problem
-at a much higher level than that of the operating system itself.
+at a much higher level than that of the operating system itself (e.g. [Ooms 2014]).
 For instance, it is often sufficient to address these concerns at the
@@
 
-Two dominant paradigms have emerged to address these issues so far:
-workflow software and cloud computing on virtual machines.  Here I
-consider how a popular emerging technology, Docker, based on Linux
-containers, offers a third route that addresses pratical and theoretical
-concerns for reproducible research not addressed by either of these
-paradigms.
+<!-- Synergistic with other reproducible-research technology -->
 
-It is worth observing from the outset that the primary barrier to
-computational reproducibility in many domain sciences has nothing
-to do with the technology discussed here, but stems rather from a
-reluctance to publish the code used to generating the results in the
-first place [@Barnes2010]. Despite extensive evidence to the contrary
-[e.g. @Ince2012], many researchers and journals continue to assume that
-summary descriptions or pseudo-code provide a sufficient description of
-computational methods used in data gathering, processing, simulation,
-visualization, or analysis.  Until such code is available in the first
-place, we cannot even begin to encounter the problems that the approaches
-discussed here set out to solve. As a result, few domain researchers
-may be fully aware of the challenges involved in effectively re-using
-published code.
+The approaches discussed here should be seen as largely synergistic
+with, rather than an alternative to, other topics in computationally
+reproducible research. For example, consider the dynamic documents
+approach ([Leisch 2002], [Peng 2011], [Xie 2013]), which seek to embed
+the code required to re-generate the results within the manuscript
+itself. This process serves as a way to distribute computational
+code with publications, to ensure that methodological discriptions,
+parameter values, figures, and so forth remain synchronized with the code,
+and to provide a way for other researchers to both tweak the code and
+'execute' the paper to reproduce the results.  While these features
+are very desirable from the perspective of reproducible research,
+they inevitably increase the number of software dependencies involved,
+including those tools that support combining code with text (e.g. [Xie
+2013]) as well as tools merely for the handling the formating and layout
+(e.g. LaTeX, pandoc).  Consequently, such a paper can fail to execute
+successfully for reasons having nothing to do with the computations
+involved in the research, but rather due to software responsible for
+implementing these dynamic features.  This problem is largely resolved by
+providing a consistent development environmment, making the approaches
+discussed here a natural complement to that of dynamic documents.
+A similar argument can be made regarding the use of workflow systems,
+which can introduce additional softare dependencies even as they seek
+to make it easier to work across them.
 
-Yet this coin has two sides: By lowering the barriers
-these barrier to re-use through the appropriate infrastructure first, the research
-community as a whole may be better positioned to take advantage of code 
-sharing practices when they are adopted.  This has cultural implications as well:
-many researchers may be reluctant to publish code today because they fear
-a primarily one-way street: others will have the technical savvy to benefit
-from their hard work, while they may not benefit from the work produced by others.
-Lowering the technical barriers to reuse provides immediate practical benefits
-that make this exchanged into a more balanced, two-way street.
+<!-- Cultural sharing practices -->
+We noted at the outset that cultural expecations responsible for a
+lack of code sharing practices in many fields are a far more extensive
+primary barrier to reproducibility than the technical barriers we have
+discussed here. Nevertheless, it may be worth considering how solutions
+to these technical barriers can influence the cultural landscape as well.
+Many researchers may be reluctant to publish code today because they
+fear a it will be primarily a one-way street: more technical savvy
+researchers then themselves can benefit from their hard work, while
+they may not benefit from the work produced by others.  Lowering the
+technical barriers to reuse provides immediate practical benefits
+that make this exchanged into a more balanced, two-way street. Another
+concern is that the difficulty imposed in preparing code to be shared,
+such as providing even semi-adequate documentation or support for other
+users to be able to install and run it in the first place is too high
+[Stodden 2010]. By lowering the barriers these barrier to re-use through
+the appropriate infrastructure first, we may also reduce certain cultural
+barriers to sharing.
 
-The issue most domain researchers do not recognize as an issue.  If computational reprod
-
-Before we begin, it is important to distinguish between the nature of the problem we seek to address here and other elements of computationally reproducible research.
-While this approach may address some of the same challenges as workflow software, dynamic documents / executable papers and ad hoc packaging systems, it can always been seen to supplement rather than replace such solutions.
-A good example comes from dynamic documents, which seek to embed the code required to re-generate the results within the manuscript itself.
-
-http://doi.ieeecomputersociety.org/10.1109/MCSE.2012.62
-
-Wouldn't it be nice if you could just revert your computer to how it was an hour ago or a day ago when everything was still working?
-Wouldn't it be nice if your collaborators could be running an identical computer environment to your own?
-
-# Reproducible computational research and the challenge of software environments
-
-- Reproducible research is an important issue
-- Computational research should in principle be the least difficult to make completely reproducible, but in practice this is not the case (see this entire special issue). 
-- Computational research relies on an ever more complex set of software and dependencies for that software
-- These dependencies frequently frustrate reproducibility (see Titus study on failing to install xx% of research software, etc, NESCENT reproducibility study, etc)
-
-# Typical solution approaches
-
-- Documentation
-- Standard environments
-- Virtual machines & Cloud computing
-
-# Challenges with common approaches
-
-# Emerging approaches: DevOpts
-
-- DevOpts: Scripting your environment
-- Vagrant
+[Vandewalle 2009]: http://doi.org/10.1109/MSP.2009.932122
+[Merali 2010]: http://doi.org/10.1038/467775a "Computational Science ... error: why scientific programming does not compute"
+[Gil 2007]: http://doi.org/10.1109/MC.2007.421 "Examining the Challenges of Scientific Workflows"
+[Hull 2006]: http://doi.org/10.1093/nar/gkl320 "Taverna: a tool for building and running workflows of services"
+[Altintas 2004]: http://doi.org/10.1109/SSDM.2004.1311241 "Kepler: an extensible system for design and execution of scientific workflows"
+[Howe 2012]: http://doi.org/10.1109/MCSE.2012.62
+[Dudley 2013]: http://doi.org/10.1038/nbt1110-1181 
+[Clark 2014]: https://berkeley.app.box.com/s/w424gdjot3tgksidyyfl
+[Ooms 2013]: http://arxiv.org/abs/1303.2140v2
+[Ooms 2014]: http://arxiv.org/abs/1406.4806
+[Peng 2011]: http://doi.org/10.1126/science.1213847
+[Barnes 2010]: http://doi.org/10.1038/467753a
+[Ince 2012]: http://doi.org/10.1038/nature10836
+[Stodden 2014]: http://doi.org/10.3233/SJI-140818
+[Stodden 2010]: http://doi.org/10.2139/ssrn.1550193
 
@@
 
-My original motivation was more on the research side, primarily from NESCent's reproducibility study and Rich's reproducibility ropensci blog post, where you might recall Docker being discussed in the comments.
+Why computational reproducibility is difficult
 
-Hilmar highlighted 3 reasons why reproducibility efforts fail even for papers that seem to do the right things: 1) dependency hell frustrating installation, 2) small holes in documentation can be major barriers for "novices" (even if novices are experts in nearby fields and nearby languages), 3) software dependency evolution breaks existing code (code rot).
+Recent studies have highlighted 5 reasons
 
-Rich highlighted about 4 reasons: 1) dependency hell / incomplete documentation on versions, 2) all the scripts & software for installing managing the 'reproducibility' part of their analysis (travis, etc) were as complex as the analysis itself.
-(more than half the software dependencies were things like RCurl,  knitr & packrat etc).
-3) Running the analysis on Travis CI means it has to finish in 50 minutes, 4) the learning curve for Github + Travis + etc is prohibitive.
+1) "Dependency Hell"
+  a) Imprecise documentation. Small holes in the documentation can be major barriers for "novices" (Sensu ..., where novices may still be experts in nearby langauages). 
+  b) Code rot. Dependency evolution breaks existing code
 
-Docker provides a "DevOpts" approach, replacing user documentation on how to install with scripts that handle provisioning.
-This pretty much solves Hilmar's #1 & #2.
-You can maintain a timestamped binary of the Docker image to address Hilmar's #3.
+Solution to this is scripted installations: DevOpts. At the Operating System level!
 
-Rich's #2 is largely handled by providing an image where this stuff is all set up for you, so it building on the ropensci docker image, you need to add only what is necessary for your actual analysis.
-His #3 is addressed bc instead of running on Travis, you can run this locally (on Windows/Mac/Linux) or on a cloud of your choice with as much horsepower as you need.
-I think this lowers the learning curve (#4) some, (for instance, you can use git in the usual way but you can also fully interact without knowing git).
+2) Complexity of existing reproducibility options 
+  a) Virtual Machines
+  b) Workflow software
+  c) Continuous Integration / software development practices
+Barriers to entry / learning
+Barriers to reuse / remix
 
-My hope is, that instead of the rather giant set-up Rich et al needed, a user could just have their paper as say, an R script or rmarkdown file, develop & run it locally in the docker-ropensci environment, and then anyone having trouble getting it to run in their native R should at least be able to run it in the docker-ropensci environment.
-Docker is super-easy to extend: they could make their own minimal Dockerfile like this: https://github.com/ropensci/RNeXML/blob/devel/manuscripts/Dockerfile There's a few other cool things about Docker (The git-like way it handles images etc) too.
+Docker does 4 new things:
+
+1) Remix.  Titus has an excellent post, "Virtual Machines Considered
+Harmful for reproducibility" [1] , essentially pointing out that "you
+can't install an image for every pipeline you want...".  In contrast,
+Docker containers are designed to work exactly like that -- reusable
+building blocks you can link together with very little overhead in
+disk space or computation.  This more than anything else sets Docker
+apart from the standard VM approach.
+
+[Clark 2014]
+
+> In scientific computing the environment was commonly managed via Makfiles & Unix-y hacks, or alternatively with monolithic software like Matlab.
+More recently, centralized package management has provided curated tools that work well together.
+But as more and more essential functionality is built out across a variety of systems and languages, the value -- and also the difﬁculty -- of coordinating multiple tools continues to increase.
+Whether we are producing research results or web services, it is becoming increasingly essential to set up new languages, libraries, databases, and more.
+
+> Documentation for complex software environments is stuck between two opposing demands.
+To make things easier on novice users, documentation must explain details relevant to factors like different operating systems.
+Alternatively, to save time writing and updating documentation, developers like to abstract over such details.
+
+## A "DevOpts" approach
+
+[Clark 2014]
+
+ Provisioning scripts. Docker images are not 'black boxes'. A
+"Dockerfile" is a simple make-like script which installs all the
+software necessary to re-create ("provision") the image.  This has
+many advantages: (a) The script is much smaller and more portable than
+the image. (b) the script can be version managed (c) the script gives
+a human readable (instead of binary) description of what software is
+installed and how. This also avoids pitfalls of traditional
+documentation of dependencies that may be too vague or out-of-sync.
+(d) Other users can build on, modify, or extend the script for their
+own needs.  All of this is what we call the he "DevOpts" approach to
+provisioning, and can be done with AMIs or other virtual machines
+using tools like Ansible, Chef, or Puppet coupled with things like
+Packer or Vagrant (or clever use of make and shell scripts).
+
+For a much better overview of this "DevOpts" approach in the
+reproducible research context and a gentle introduction to these
+tools, I highly recommend taking a look at Clark et al [2].
+
+3) You can run the docker container *locally*.  I think this is huge.
+In my experience, most researchers do their primary development
+locally.  By running RStudio-server on your laptop, it isn't necessary
+for me to spin up an EC2 instance (with all the knowledge & potential
+cost that requires).  By sharing directories between Docker and the
+host OS, a user can still use everything they already know -- their
+favorite editor, moving files around with the native OS
+finder/browser, using all local configurations, etc, while still
+having the code execution occur in the container where the software is
+precisely specified and portable.  Whenever you need more power, you
+can then deploy the image on Amazon, DigitalOcean, a bigger desktop,
+your university HPC cluster, your favorite CI platform, or wherever
+else you want your code to run.  (On Mac & Windows, this uses
+something called boot2docker, and was not very seamless early on.  It
+has gotten much better and continues to improve.)
+
+Related challenges:  RStudio and a consistent environment. 
+- Linked volumes 
+- linked containers 
+- signed containers?
+
+Hub / image delivery
+
+4) Versioned images.  In addition to version managing the Dockerfile,
+the images themselves are versioned using a git-like hash system
+(check out: docker commit, docker push/pull, docker history, docker
+diff, etc).  They have metadata specifying the date, author, parent
+image, etc.  We can roll back an image through the layers of history
+of its construction, then build off an earlier layer.  This also
+allows docker to do all sorts of clever things, like avoiding
+downloading redundant software layers from the docker hub.  (If you
+pull a bunch of images that all build on ubuntu, you don't get n
+copies of ubuntu you have to download and store).  Oh yeah, and
+hosting your images on Docker hub is free (no need to pay for an S3
+bucket... for now?) and supports automated builds based on your
+dockerfiles, which acts as a kind of CI for your environment.
+Versioning and diff'ing images is a rather nice reproducibility
+feature.
+
+5) Linking images 
+Docker is not a virtual machine; containers are designed expressly to be remixable blocks.
+You can put an R engine on one container and a mysql database on another and connect them.
+Docker philosophy aims at one function per container to maximize this reuse.
+of course it's up to you to build this way rather than a single monolithic dockerfile, but the idea of linking containers is a technical concept at the heart of docker that offers a second and very different way to address the 'remix' problem of VMs.
+
+Cultural challenges
 

### 2014-09-12 manuscript.Rmd
writing

@@
 layout: preprint
-title: "An introduction to Docker for reproducible research and collaboration"
+title: "An introduction to Docker for reproducible research, with examples from the R environment"
 author: 
@@
 abstract: |
-  Academic research
+  Docker is a promising solution to
 
@@
 Reproducible research has recieved an increasing level of attention
-throughout the scientific community [The Economist 2014]. A all steps
-of the scientific process, from data collection and processing, to
-analyses, visualizations and conclusions depend ever more on computation
-and algorithms, _computational reproducibility_ has recieved particular
+throughout the scientific community (e.g. [Peng 2011], [Nature editorial])
+and the public at large (e.g. [The Economist 2014]). All steps of the
+scientific process, from data collection and processing, to analyses,
+visualizations and conclusions depend ever more on computation and
+algorithms, _computational reproducibility_ has recieved particular
 attention [Merali 2010].  Though in principle this algorithmic dependence
@@
 
-## A technical and cultural problem
+# A cultural problem
 
@@
 
-A lack of requirements or incentives no doubt plays a crucial role
-in discouraging sharing ([Barnes 2010]). Nevertheless, it is easy to
-underestimate the siginficant barriers raised by a lack of familiar,
-intuitive, and widely adopted tools for addressing the challenges of
-computational reproducibility. Surveys and case studies find that a lack
-of time, more than innate opposition to sharing, discourages researchers
-from providing code ([Stodden 2010], [FitzJohn 2014]).
+A lack of requirements or incentives no doubt plays a crucial role in
+discouraging sharing ([Barnes 2010], [Stodden 2013]). Nevertheless,
+it is easy to underestimate the siginficant barriers raised by a lack
+of familiar, intuitive, and widely adopted tools for addressing the
+challenges of computational reproducibility. Surveys and case studies find
+that a lack of time, more than innate opposition to sharing, discourages
+researchers from providing code ([Stodden 2010], [FitzJohn 2014]).
+
+# The four technical challenges of computational reproducibility
+
+By restricting ourselves to studies of where code has been made
+available, we may sidestep for the moment the cultural challenges
+to reproducibility so that we may focus on the technical ones; in
+particular, those challenges for which improved tools and techniques
+rather than merely norms of behavior can contribute substantially to
+improved reproducibility.
+
+Studies focusing on code that has been made available with scientific
+publications regularly find the same common issues that pose substantial
+barriers to reproducing the original results or building on that code
+(e.g. [Lapp 2014], [Garijo 2013], [Gilbert 2012], [Collberg 2014],
+[Brown 2014]), which I attempt to summarize as follows.
+
+<!-- Is this a permitted term? -->
+
+1. **"Dependency Hell"**
+
+A recent study by researchers at the University of Arizona found that
+less than 50% of software could even be successfully built or installed
+[Collberg 2014], and in an ongoing effort by other researchers to
+replicate that study [Brown 2014]. Installing or building software
+necessary to run the code in question assumes the ability to recreate
+the computational environment of the original researchers.
+
+<!-- FIXME more stuff on what this means-->
 
@@
 
-<!-- Operating system - level approach -->
+<!-- Omit this? -->
+Differences in numerical evaluation, such as arise in floating point
+arithmetic or computers or even ambiguities in standardized programming
+languages ("order-of-evaluation" problems) can be responsible for
+differing results between or even within the same computational platform
+([Ince 2012]). Such issues make it difficult to restrict the true dependencies
+of the code to higher level environments such as that of a given scripting
+language, indpendent of the underlying OS or even hardware itself.
+
+2.  **Imprecise documentation**
+
+Documentation on how to install and run code associated with published
+research is another frequent barrier to replication.  [Lapp 2014] found
+this impairs a researcher's ability to install and build the software
+necessary, as even small holes in the documentation were found to be
+major barriers, particularly for "novices" (Sensu [Garijo 2013] --
+where novices may be experts in nearby langauages but unfamiliar with
+the package managers and other tools of the language involved). This
+same problem is discussed in [Clark 2014].  Imprecise documentation
+goes well beyond issues of the software environment itself: incomplete
+documentation of parameters involved meant as few as 30% of analyses
+(n=34) using the popular software STRUCTURE could be reporoduced in the
+study of [Gilbert 2012].
+
+3.  **Code rot**
+
+Software dependencies are not static elements, but recieve regular
+updates that may fix bugs, add new features or deprecate old features
+(or even entire dependencies themselves).  Any of these changes can
+potentially change the results generated by the code.  As some of these
+changes may indeed resolve valid bugs or earlier problems with underlying
+code, it will often be insufficient to demonstrate that results can be
+reproduced when using the original versions, a problem sometimes known
+as "code rot."  Researchers will want to know if the results are robust
+to the changes. The case studies in [Lapp 2014] provide examples of
+these problems.
+
+4. **Barriers to adoption and reuse in existing solutions**
+
+Technological solutions such as workflow software, virtual machines,
+continuous integration services, and best practices from software
+development would address many of the issues frequently frustrating
+reproducibility. However, researchers face siginificant barriers to entry
+in learning these tools and approaches which are not part of their typical
+curriculum, or lack incentives co-measurate with the effort required
+([Joppa 2013], [FitzJohn 2014]).
+
+Existing proposals to address the challenges 1-3 either involve adopting
+an relatively new workflow that may be foreign to both the researcher and
+their collaborators, as in the case of many workflow software solutions,
+or adopt a relatively black-box approach that cannot be easily extended
+by other researchers, as in the case of virtual machines.
+
+------------------------------------
+
+<!--
+# Docker
+
 At the heart of the computational reproducibility challenge is the 
@@
 that underpin such work.
+
+-->
+
+# The alternatives
 
@@
 
-[Dudley 2013] gives several reasons for this:
+[Dudley 2013] gives several reasons that such comprehensive workflow
+systems have not been more successful:
 
@@
 their markets through proprietary formats and interfaces; (iii)
-investigators naturally tend to want to ‘own’ and control their
+investigators naturally tend to want to 'own' and control their
 research tools; (iv) even the most generalized software will not be
@@
 then we can just take a snapshot of the entire computer: data, software, 
-operating system and all.
-
-To make this practical, [Dudley 2013] and [Howe 2012]
+operating system and all. To make this practical, [Dudley 2013] and [Howe 2012]
 propose using virtual machine images that will run on the cloud, such as Amazon's
@@
 Critics of the use of VMs to support reproducibility highlight that the 
-approach is too much of a black box and ill suited for
+approach is too much of a black box and thus ill suited for reproducibility [Watson 2012].
+While the approach sidesteps the need to either install or even document the
+dependencies, this also makes it more difficult for other researchers to
+understand, evaluate, or alter those dependencies.  Moreover, other research
+cannot easily build on the virtual machine in a consistent and scalable way. 
+If each study provided it's own virtual machine, any pipeline combining the 
+tools of multiple studies would quickly become impractical or impossible 
+to implement.
 
-Here I consider how a popular emerging technology, Docker, based on
-Linux containers, offers a third route that addresses pratical and
-theoretical concerns for reproducible research not addressed by either
-of these paradigms.
+## A "DevOpts" approach
 
-Wouldn't it be nice if you could just revert your computer to how it was an hour ago or a day ago when everything was still working?
-Wouldn't it be nice if your collaborators could be running an identical computer environment to your own?
+The problems highlighted here are not unique to academic software, but
+impact software development in general. While the academic research 
+literature has frequently focused on the development of workflow software
+dedicated to particular domains, or otherwise to the use of virtual machines,
+the software development community has recently emphasized a philosophy 
+(rather than a particular tool), known as _Development_ and Systems _Operation_, or more 
+frequently just "DevOpts." The approach is characterized by scripting,
+rather than documenting, a description of the necessary dependencies 
+for software to run, usually from the Operating System (OS) on up.
 
-# Reproducible computational research and the challenge of software environments
+[Clark 2014] describes the DevOpts approach
+along with both its relevance to reproducible research and examples of
+its use in the academic research context.
 
-- Reproducible research is an important issue
-- Computational research should in principle be the least difficult to make completely reproducible, but in practice this is not the case (see this entire special issue). 
-- Computational research relies on an ever more complex set of software and dependencies for that software
-- These dependencies frequently frustrate reproducibility (see Titus study on failing to install xx% of research software, etc, NESCENT reproducibility study, etc)
+Their paper provides both a consise statement of the problem we
+have already discussed and the solution provided under this banner:
 
-## Alternatives and other considerations
+> In scientific computing the environment was commonly managed via
+Makfiles & Unix-y hacks, or alternatively with monolithic software
+like Matlab. More recently, centralized package management has provided
+curated tools that work well together. But as more and more essential
+functionality is built out across a variety of systems and languages,
+the value -- and also the difficulty -- of coordinating multiple tools
+continues to increase. Whether we are producing research results or web
+services, it is becoming increasingly essential to set up new languages,
+libraries, databases, and more.
 
-<!-- Package managers -->
+> Documentation for complex software environments is stuck between two
+opposing demands. To make things easier on novice users, documentation
+must explain details relevant to factors like different operating
+systems. Alternatively, to save time writing and updating documentation,
+developers like to abstract over such details.
 
-It is possible, and indeed often desirable, to tackle this problem
-at a much higher level than that of the operating system itself (e.g. [Ooms 2014]).
-For instance, it is often sufficient to address these concerns at the
-level of a particular scripting language rather than the level of the
-operating system. Most major scripting languages have dedicated tools for
-handling dependencies and versioning (e.g. Ruby gems, R packages, Python
-eggs, perl modules, javascript node package manager); with some languages
-having the fortune or misfortune to have multiple systems. Though systems
-differ somewhat in their strategies and their effectiveness, all are
-generally vulnerable to changes in their lower-level dependencies (such as
-compilers, C libraries, and so forth) that may differ between platforms.
-More pointedly, researchers whose work is not easily isolated to a single
-such environment but combines tools from different systems (including
-combining those tools used for formatting & visual display, such as
-LaTeX or pandoc, with scripting or analysis tools such as R). Moreover,
-addressing reproducibility at the level of the operating system avoids
-constructing approaches unique to a specific language or tool.
+The authors contrast this to the DevOpts approach:
 
-<!-- Synergistic with other reproducible-research technology -->
+> A DevOps approach to "documenting" an application might consist of
+providing brief descriptions of various install paths, along with scripts
+or "recipes" that automate setup.
 
-The approaches discussed here should be seen as largely synergistic
-with, rather than an alternative to, other topics in computationally
-reproducible research. For example, consider the dynamic documents
-approach ([Leisch 2002], [Peng 2011], [Xie 2013]), which seek to embed
-the code required to re-generate the results within the manuscript
-itself. This process serves as a way to distribute computational
-code with publications, to ensure that methodological discriptions,
-parameter values, figures, and so forth remain synchronized with the code,
-and to provide a way for other researchers to both tweak the code and
-'execute' the paper to reproduce the results.  While these features
-are very desirable from the perspective of reproducible research,
-they inevitably increase the number of software dependencies involved,
-including those tools that support combining code with text (e.g. [Xie
-2013]) as well as tools merely for the handling the formating and layout
-(e.g. LaTeX, pandoc).  Consequently, such a paper can fail to execute
-successfully for reasons having nothing to do with the computations
-involved in the research, but rather due to software responsible for
-implementing these dynamic features.  This problem is largely resolved by
-providing a consistent development environmment, making the approaches
-discussed here a natural complement to that of dynamic documents.
-A similar argument can be made regarding the use of workflow systems,
-which can introduce additional softare dependencies even as they seek
-to make it easier to work across them.
+which elegantly addresses both the demand for simplicity of use (one
+executes a script instead of manually managing the environmental setup)
+and comprehensiveness of implementation. [Clark 2014] are careful to note
+that this is not so much a technological shift as a philosophical one:
 
-<!-- Cultural sharing practices -->
-We noted at the outset that cultural expecations responsible for a
-lack of code sharing practices in many fields are a far more extensive
-primary barrier to reproducibility than the technical barriers we have
-discussed here. Nevertheless, it may be worth considering how solutions
-to these technical barriers can influence the cultural landscape as well.
-Many researchers may be reluctant to publish code today because they
-fear a it will be primarily a one-way street: more technical savvy
-researchers then themselves can benefit from their hard work, while
-they may not benefit from the work produced by others.  Lowering the
-technical barriers to reuse provides immediate practical benefits
-that make this exchanged into a more balanced, two-way street. Another
-concern is that the difficulty imposed in preparing code to be shared,
-such as providing even semi-adequate documentation or support for other
-users to be able to install and run it in the first place is too high
-[Stodden 2010]. By lowering the barriers these barrier to re-use through
-the appropriate infrastructure first, we may also reduce certain cultural
-barriers to sharing.
+> The primary shift that's required is not one of new tooling, as most
+developers already have the basic tooling they need. Rather, the needed
+shift is one of philosophy.
 
-[Vandewalle 2009]: http://doi.org/10.1109/MSP.2009.932122
-[Merali 2010]: http://doi.org/10.1038/467775a "Computational Science ... error: why scientific programming does not compute"
-[Gil 2007]: http://doi.org/10.1109/MC.2007.421 "Examining the Challenges of Scientific Workflows"
-[Hull 2006]: http://doi.org/10.1093/nar/gkl320 "Taverna: a tool for building and running workflows of services"
-[Altintas 2004]: http://doi.org/10.1109/SSDM.2004.1311241 "Kepler: an extensible system for design and execution of scientific workflows"
-[Howe 2012]: http://doi.org/10.1109/MCSE.2012.62
-[Dudley 2013]: http://doi.org/10.1038/nbt1110-1181 
-[Clark 2014]: https://berkeley.app.box.com/s/w424gdjot3tgksidyyfl
-[Ooms 2013]: http://arxiv.org/abs/1303.2140v2
-[Ooms 2014]: http://arxiv.org/abs/1406.4806
-[Peng 2011]: http://doi.org/10.1126/science.1213847
-[Barnes 2010]: http://doi.org/10.1038/467753a
-[Ince 2012]: http://doi.org/10.1038/nature10836
-[Stodden 2014]: http://doi.org/10.3233/SJI-140818
-[Stodden 2010]: http://doi.org/10.2139/ssrn.1550193
+Nevertheless, a growing suite of tools designed explicitly for this
+purpose have rapidly replaced the use of general purpose tools (such
+as Makefiles, bash scripts) to become synonomous with the "DevOpts"
+philosophy.  [Clark 2014] reviews many of these "DevOpts" tools, their
+different roles, and their application in reproducible research.
+
+I focus the remainder of this paper on one of the most recent and rapidly
+growing among these, called Docker, which offers several promising
+developments above and beyond the tools highlighted in [Clark 2014]
+in providing a DevOpts approach to reproducible research. My goal
+in focusing on this example is not to promote a particular solution, 
+but to anchor the discussion of technical solutions to reproducibility
+challenges in concrete examples.
+
+<!-- Link on how to install docker -->
+<!-- Examples here are not meant as a tutorial to docker, but 
+rather to illustrate concepts a reader should be able to follow whether 
+or not they have any previous knowledge of docker and its commands -->
 
@@
 
-- Introduction and hype
+Docker is an open source (Apache Version 2) software platform written in
+the Go programming language.
 
-- Technologies from operating systems research: LXC containers, virtualization of OS, git-like verioning
+Docker builds on many long-familiar technologies from operating
+systems research: LXC containers, virtualization of the OS, and
+a hash-based or git-like verioning and differencing system, among
+others. Researchers familiar with LXC may be tempted to view the Docker
+system as little more than a creative application and marketing of an
+existing technology. However, the Docker approach adds several features
+crucial for its use in reproducible research (adapted from [Docker
+FAQ](https://docs.docker.com/faq/#what-does-docker-add-to-just-plain-lxc)):
 
-# Using Docker for Reproducible Research
+- Portable deployment: Docker defines a format for bundling and exchanging
+containers such that they run identically across platforms.
 
-- Example:
+- Automatic builds: Docker defines a format (Dockerfiles) that provide
+a simple, scripted way to construct containers, following the DevOpts
+philosophy.
 
-```{r}
-# Read in an external data file 
-# Load some libraries with external dependencies
-# Compute some statistics, plot 
+- Versioning: Docker provides git-like capibilities for tracking changes
+to a container, inspecting the history, metadata, or diff between
+versions, committing versions, rolling back changes, and incremential
+uploads and downloads that send only the differences (like `git push`
+or `git pull`).
+
+- Component reuse: Any container can be used as the base for any other.
+
+- Sharing: Docker provides a public registry (Docker Hub) where any
+user can make images avaialble to any other. (The registry itself is
+also open source so that it can be deployed privately).
+
+Each of these elements have clear implications for reproducible research.
+
+# Using Docker in an R environment
+
+Perhaps the most important feature of a reproducible research tool is
+that it be easy to learn and fit relatively seamlessly into exsiting
+workflow patterns of domain researchers.  As we have seen, this, more
+than any other concern, can explain the relatively low uptake of previously
+proposed solutions.  <!-- too strong? -->
+
+Being a new and unfamiliar tool to most domain scientists, Docker is far
+from immune to the same critique.  Nevertheless, Docker takes us several 
+key steps towards an approach that can be easily adopted in a research 
+context.
+
+Primary among these is the ability to run Docker locally on all major
+operating systems with minimal prior expertise.  On systems not already
+based on the Linux Kernel (such as Mac or Windows clients), this is 
+accomplished through the use of a small VirtualBox-based VM running on
+the host OS called `boot2docker`.
+
+Local development offers several important advantages in encouraging 
+adoption.
+
+-----
+
+In the R environment, the open source RStudio IDE provides another key component
+in making this system accessible to most domain scientists. Because the RStudio 
+IDE is written completely with standard web languages: javascript, css, and html,
+its web-based client, RStudio-Server, looks and feels identical to its popular 
+desktop IDE.  RStudio Server provides a way to interact with R on a remote environment
+without the latency, X tunnelling, etc typically involved in running R on a remote
+server.
+
+For our purposes, it provides users a way to interact with R, the filesystem,
+git, text editors, and graphics running on a Docker container.  Users already
+familiar with the popular IDE can thus benefit from the reproduciblity and 
+portability features provided by running R in a container environment without
+having to adapt to a new workflow for managing these things that is usually 
+necessary when moving to a cloud environment.
+
+RStudio-Server is not the only way to provide a familiar developer environment
+to users working with a development environment running in a container on their
+local computer.  Docker supports the sharing of volumes (directories) between
+the container and the host OS.  This allows a user to rely on the familiar tools
+of the host OS for roles such as text editing, file browsing, or version control,
+while still allowing code execution to occur inside the controlled development 
+environment of the container. In this example, we launch an interactive R console
+in a container that is linked to our current working directory:
+
+```bash
+docker run -v $(pwd):/ -it eddelbuettel/docker-ubuntu-r /usr/bin/R
 ```
+
+(Clearly a similar command could be used just as well with interactive
+shells from other languages such as `ipython` or `irb`). The resulting
+system behaves almost identically[^2] to running R on the command line
+of the host OS.
+
+[^2]: External windows, such graphics windows cannot be opened directly from the container in this setup.
+Graphics would have to be saved to desk as raster or vector files and viewed on the host OS, until a better solution can be found.
+RStudio-server is one way around this.
 
@@
 
-Why computational reproducibility is difficult
+## Docker remix
 
-Recent studies have highlighted 5 reasons
+The approach of Linux Containers represented by Docker offers a technical
+solution to what is frequently seen as the primary weakness of the
+standard virtual machine approach to reproducibility - reusing and
+remixing elements.  To some extent this is already addressed by the
+"DevOpts" approach of Dockerfiles, providing a scripted description
+of the environment that can be tweaked and altered, but also includes 
+something much more fundamental to Docker.
 
-1) "Dependency Hell"
-  a) Imprecise documentation. Small holes in the documentation can be major barriers for "novices" (Sensu ..., where novices may still be experts in nearby langauages). 
-  b) Code rot. Dependency evolution breaks existing code
+The challenge to reusing virtual machines can be summarized as "you can't install an image for every pipeline you want..." [Watson 2012].
+In contrast, this is exactly how Docker containers are designed to work.
+There are at least two ways in which Docker supports this kind
+of extensibility.
 
-Solution to this is scripted installations: DevOpts. At the Operating System level!
+First, Docker makes it immensely easy to build one container on top of
+another.  Rather than copy a Dockerfile and then start adding more lines
+to the bottom, we can declare a new Dockerfile is built on an old one
+using the `FROM` directive. Here is an example Dockerfile that adds the
+R statistical computing environment to a basic Ubuntu Linux distribution.
 
-2) Complexity of existing reproducibility options 
-  a) Virtual Machines
-  b) Workflow software
-  c) Continuous Integration / software development practices
-Barriers to entry / learning
-Barriers to reuse / remix
+```
+FROM ubuntu:latest
+RUN apt-get update && apt-get -y install r-recommended
+```
 
-Docker does 4 new things:
+This acts like a software dependency; but unlike other software, a Dockerfile
+must have exactly one dependency (one `FROM` line). Note that a particular
+version of the dependency can be specified using the `:`.  We can in fact
+be much more precise, declaring not only version numbers like `ubuntu:14:04`,
+but specific cryptographic hashes that ensure we get the exact same image
+every time.
+
+We can now build this Dockerfile and give it name, by running in the working directory:
+
+```bash
+docker build -t r-recommended .
+```
+
+The key point here is that other researchers can easily build off this 
+new image we have just created, extending our work directly, rather than
+having to go back to the original image.
+
+However, before others can re-use our image we need a way to share it.
+
+### Docker hub
+
+While Docker images tend to be much smaller than equivalent virtual machines, moving 
+around even 100s of gigabytes can be a challenge.  In their work on virtual machines,
+[Dudley 2014] recommend only running these tools on cloud servers such as the Amazon
+EC2 system. As most researchers still develop software locally and may not have ready
+access to these resources, such a requirement adds an additional barrier to reuse.
+Fortunately in the case of Docker we have a somewhat more elegant solution through
+the use of the Docker Hub [^1].  Docker provides a convenient way to share any image
+publicly or privately through the Hub after creating a free account.  Here, we share
+a public copy of the image we have just created by using the `docker push` command,
+followed by the name of the image:
+
+```bash
+docker push username/r-recommended
+```
+
+Another user can now build directly on our image, rather than on the `ubuntu:latest`
+image we used as a base:
+
+```
+FROM username/r-recommended 
+RUN apt-get update && apt-get -y install r-cran-teachingdemos
+```
+
+[^1]: Similar hubs are in fact available for full virtual machine images, such as the Vagrant hub.
+
+Though we have shown adding only a single piece of software in each step,
+clearly this approach can be particularly powerful in building up more complex
+environments.
+
+Each one acts as a building block providing just what is necessary to
+run one particular service or element, and exposing just what is necessary
+to link it together with other blocks.  For instance, we could have one
+container running a PostgreSQL database which serves data to another container
+running a python environment to analyze the data:
+
+```bash
+docker run -d --name db training/postgres
+docker run -d -P --link db:db training/webapp python app.py
+```
+
+Unlike the much more heavyweight virtual machine approach, containers
+are implemented in way such that a single computer can easily run 100s
+of such services each in their own container, making it easy to break
+computational elements down into logically reusable chunks that come,
+batteries included, with everything they need to run reproducibly.
+A researcher more familiar with python could connect our MySQL container
+to their python container, and so forth.
+
+<!-- too much detail? -->
+The implementation of this would nonetheless depend on the individual 
+researcher's willingness to break things into multiple containers, as
+it is indeed possible to run everything in a monolithic container. 
+This is not necessarily bad, as it simplifies the workflow (no need
+to execute and link separate containers), and can be refactored into
+multiple containers when appropriate. It remains to be seen whether
+linking containers through this mechanism
 
 1) Remix.  Titus has an excellent post, "Virtual Machines Considered
-Harmful for reproducibility" [1] , essentially pointing out that "you
-can't install an image for every pipeline you want...".  In contrast,
+Harmful for reproducibility" [1] , .  In contrast,
 Docker containers are designed to work exactly like that -- reusable
@@
 
-[Clark 2014]
-
-> In scientific computing the environment was commonly managed via Makfiles & Unix-y hacks, or alternatively with monolithic software like Matlab.
-More recently, centralized package management has provided curated tools that work well together.
-But as more and more essential functionality is built out across a variety of systems and languages, the value -- and also the difﬁculty -- of coordinating multiple tools continues to increase.
-Whether we are producing research results or web services, it is becoming increasingly essential to set up new languages, libraries, databases, and more.
-
-> Documentation for complex software environments is stuck between two opposing demands.
-To make things easier on novice users, documentation must explain details relevant to factors like different operating systems.
-Alternatively, to save time writing and updating documentation, developers like to abstract over such details.
-
-## A "DevOpts" approach
+Docker is not a virtual machine; containers are designed expressly to
+be remixable blocks.  You can put an R engine on one container and a
+mysql database on another and connect them. Docker philosophy aims at
+one function per container to maximize this reuse.  of course it's up
+to you to build this way rather than a single monolithic dockerfile,
+but the idea of linking containers is a technical concept at the heart
+of docker that offers a second and very different way to address the
+'remix' problem of VMs.
 
@@
 
-Hub / image delivery
-
 4) Versioned images.  In addition to version managing the Dockerfile,
@@
 
-5) Linking images 
-Docker is not a virtual machine; containers are designed expressly to be remixable blocks.
-You can put an R engine on one container and a mysql database on another and connect them.
-Docker philosophy aims at one function per container to maximize this reuse.
-of course it's up to you to build this way rather than a single monolithic dockerfile, but the idea of linking containers is a technical concept at the heart of docker that offers a second and very different way to address the 'remix' problem of VMs.
+# Further considerations
 
-Cultural challenges
+## Combining virtualization with other reproducible-research tools
+
+<!-- Package managers -->
+
+It is possible, and indeed often desirable, to tackle the problem of dependencies
+at a much higher level than that of the operating system itself (e.g. [Ooms 2014]).
+For instance, it is often sufficient to address these concerns at the
+level of a particular scripting language rather than the level of the
+operating system. Most major scripting languages have dedicated tools for
+handling dependencies and versioning (e.g. Ruby gems, R packages, Python
+eggs, perl modules, javascript node package manager); with some languages
+having the fortune or misfortune to have multiple systems. Though systems
+differ somewhat in their strategies and their effectiveness, all are
+generally vulnerable to changes in their lower-level dependencies (such as
+compilers, C libraries, and so forth) that may differ between platforms.
+More pointedly, researchers whose work is not easily isolated to a single
+such environment but combines tools from different systems (including
+combining those tools used for formatting & visual display, such as
+LaTeX or pandoc, with scripting or analysis tools such as R).
+
+<!-- Dynamic documents --> 
+Dynamic documents are another example of an approach that
+
+approach ([Leisch 2002], [Peng 2011], [Xie 2013]), which seek to embed
+the code required to re-generate the results within the manuscript
+itself. This process serves as a way to distribute computational
+code with publications, to ensure that methodological discriptions,
+parameter values, figures, and so forth remain synchronized with the code,
+and to provide a way for other researchers to both tweak the code and
+'execute' the paper to reproduce the results.  While these features
+are very desirable from the perspective of reproducible research,
+they inevitably increase the number of software dependencies involved,
+including those tools that support combining code with text (e.g. [Xie
+2013]) as well as tools merely for the handling the formating and layout
+(e.g. LaTeX, pandoc).  Consequently, such a paper can fail to execute
+successfully for reasons having nothing to do with the computations
+involved in the research, but rather due to software responsible for
+implementing these dynamic features.  This problem is largely resolved by
+providing a consistent development environmment, making the approaches
+discussed here a natural complement to that of dynamic documents.
+A similar argument can be made regarding the use of workflow systems,
+which can introduce additional softare dependencies even as they seek
+to make it easier to work across them.
+
+### Impacting cultural norms?
+
+We noted at the outset that cultural expecations responsible for a
+lack of code sharing practices in many fields are a far more extensive
+primary barrier to reproducibility than the technical barriers we have
+discussed here. Nevertheless, it may be worth considering how solutions
+to these technical barriers can influence the cultural landscape as well.
+Many researchers may be reluctant to publish code today because they
+fear a it will be primarily a one-way street: more technical savvy
+researchers then themselves can benefit from their hard work, while
+they may not benefit from the work produced by others.  Lowering the
+technical barriers to reuse provides immediate practical benefits
+that make this exchanged into a more balanced, two-way street. Another
+concern is that the difficulty imposed in preparing code to be shared,
+such as providing even semi-adequate documentation or support for other
+users to be able to install and run it in the first place is too high
+[Stodden 2010]. By lowering the barriers these barrier to re-use through
+the appropriate infrastructure first, we may also reduce certain cultural
+barriers to sharing.
 
@@
 - Dirk Eddlebuettel, Rich FitzJohn, Yihui Xie, Titus Brown, John Stanton-Geddes and others for helpful discussions about reproducibility, virtualization, and Docker that have helped shape this manuscript.
+
+# References
+
+[Garijo 2013]:     http://doi.org/10.1371/journal.pone.0080278
+[Lapp 2014]: https://storify.com/hlapp/reproducibility-repeatability-bigthink
+[Gilbert 2012]: 10.1111/j.1365-294X.2012.05754.x
+[Collberg 2014]: http://reproducibility.cs.arizona.edu/v1/tr.pdf
+[Brown 2014]: http://cs.brown.edu/~sk/Memos/Examining-Reproducibility/
+
+[Vandewalle 2009]: http://doi.org/10.1109/MSP.2009.932122
+[Merali 2010]: http://doi.org/10.1038/467775a "Computational Science ... error: why scientific programming does not compute"
+[Gil 2007]: http://doi.org/10.1109/MC.2007.421 "Examining the Challenges of Scientific Workflows"
+[Hull 2006]: http://doi.org/10.1093/nar/gkl320 "Taverna: a tool for building and running workflows of services"
+[Altintas 2004]: http://doi.org/10.1109/SSDM.2004.1311241 "Kepler: an extensible system for design and execution of scientific workflows"
+[Howe 2012]: http://doi.org/10.1109/MCSE.2012.62
+[Dudley 2013]: http://doi.org/10.1038/nbt1110-1181 
+[Clark 2014]: https://berkeley.app.box.com/s/w424gdjot3tgksidyyfl
+[Ooms 2013]: http://arxiv.org/abs/1303.2140v2
+[Ooms 2014]: http://arxiv.org/abs/1406.4806
+[Peng 2011]: http://doi.org/10.1126/science.1213847
+[Barnes 2010]: http://doi.org/10.1038/467753a
+[Ince 2012]: http://doi.org/10.1038/nature10836
+[Stodden 2014]: http://doi.org/10.3233/SJI-140818
+[Stodden 2010]: http://doi.org/10.2139/ssrn.1550193
+[Stodden 2013]: http://www.davidhbailey.com/dhbpapers/icerm-report.pdf "ICERM Working group report: Setting the Default to Reproducible: Reproducibility in Computational and Experimental Mathematics"
+
+[Nature editorial]: http://doi.org/10.1038/483509a
+[Nature Special Issue]: http://www.nature.com/nature/focus/reproducibility/
+
+[van der Meulen 2008]: http://doi.org/10.1109/TSE.2008.70 "The effectiveness of software diversity in a large population of programs."
+
+[Watson 2012]: https://twitter.com/BioMickWatson/status/265037994526928896

### 2014-09-15 manuscript.Rmd
small edits

@@
 
-Existing proposals to address the challenges 1-3 either involve adopting
-an relatively new workflow that may be foreign to both the researcher and
-their collaborators, as in the case of many workflow software solutions,
-or adopt a relatively black-box approach that cannot be easily extended
-by other researchers, as in the case of virtual machines.
+<!-- Obviously reproducible research barriers go much deeper than this,
+code that is not modular, not documented, not functionalized or otherwise
+easy to read and follow are all obstacles to resuse.-->
 
-------------------------------------
+<!-- A wide range of proposals, we focus on those at the operating system level -->
+Though a wide variety of approaches exists to work around these challenges,
+few operate on a low enough level to provide a general solution.  [Clark 2014]
+provide an excellent description of this situation:
 
-<!--
-# Docker
+> In scientific computing the environment was commonly managed via
+Makfiles & Unix-y hacks, or alternatively with monolithic software
+like Matlab. More recently, centralized package management has provided
+curated tools that work well together. But as more and more essential
+functionality is built out across a variety of systems and languages,
+the value -- and also the difficulty -- of coordinating multiple tools
+continues to increase. Whether we are producing research results or web
+services, it is becoming increasingly essential to set up new languages,
+libraries, databases, and more.
 
-At the heart of the computational reproducibility challenge is the 
-complexity, diversity, and rapidly changing nature of the computer software
-that underpin such work.
+There are two dominant approaches to this issue of coordinating multiple
+tools: Workflows and Virtual Machines (VMs).
 
--->
-
-# The alternatives
-
-<!-- Current approaches at the Operating Systems level -->
+# Current approaches
 
@@
 
-In short, workflow software expects a new approach to computational research
-
-In contrast, virtual machines (VMs) offer a far more brute-force approach.  If the goal
-is to reproduce everything as it ran on the researcher's own computer, well 
-then we can just take a snapshot of the entire computer: data, software, 
-operating system and all. To make this practical, [Dudley 2013] and [Howe 2012]
+In short, workflow software expects a new approach to computational research. 
+In contrast, virtual machines (VMs) offer a more direct approach. Since
+the computer Operating System (OS) already provides the software layer
+responsible for coodinating all the different elements running on the computer,
+the VM approach captures the OS and everything running on it whole-cloth.
+To make this practical, [Dudley 2013] and [Howe 2012]
 propose using virtual machine images that will run on the cloud, such as Amazon's
@@
 
+<!-- Scripted provisioning: a solution to the weaknesses of virtual machines -->
 ## A "DevOpts" approach
 
-The problems highlighted here are not unique to academic software, but
-impact software development in general. While the academic research 
+The problems highlighted here are not unique to _academic_ software,
+but impact software development in general. While the academic research
 literature has frequently focused on the development of workflow software
-dedicated to particular domains, or otherwise to the use of virtual machines,
-the software development community has recently emphasized a philosophy 
-(rather than a particular tool), known as _Development_ and Systems _Operation_, or more 
-frequently just "DevOpts." The approach is characterized by scripting,
-rather than documenting, a description of the necessary dependencies 
-for software to run, usually from the Operating System (OS) on up.
-
-[Clark 2014] describes the DevOpts approach
-along with both its relevance to reproducible research and examples of
-its use in the academic research context.
-
-Their paper provides both a consise statement of the problem we
-have already discussed and the solution provided under this banner:
-
-> In scientific computing the environment was commonly managed via
-Makfiles & Unix-y hacks, or alternatively with monolithic software
-like Matlab. More recently, centralized package management has provided
-curated tools that work well together. But as more and more essential
-functionality is built out across a variety of systems and languages,
-the value -- and also the difficulty -- of coordinating multiple tools
-continues to increase. Whether we are producing research results or web
-services, it is becoming increasingly essential to set up new languages,
-libraries, databases, and more.
+dedicated to particular domains, or otherwise to the use of virtual
+machines, the software development community has recently emphasized a
+philosophy (rather than a particular tool), known as _Development_ and
+Systems _Operation_, or more frequently just "DevOpts." The approach is
+characterized by scripting, rather than documenting, a description of the
+necessary dependencies for software to run, usually from the Operating
+System (OS) on up.  [Clark 2014] describes the DevOpts approach along
+with both its relevance to reproducible research and examples of its
+use in the academic research context. They identify the difficulties
+we have discussed so far in terms of effective documentation:
 
@@
 
-The authors contrast this to the DevOpts approach:
+The authors contrast this to the DevOpts approach, where dependency 
+documentation is _scripted_:
 
@@
 
-which elegantly addresses both the demand for simplicity of use (one
+This elegantly addresses both the demand for simplicity of use (one
 executes a script instead of manually managing the environmental setup)
@@
 
-I focus the remainder of this paper on one of the most recent and rapidly
-growing among these, called Docker, which offers several promising
-developments above and beyond the tools highlighted in [Clark 2014]
-in providing a DevOpts approach to reproducible research. My goal
-in focusing on this example is not to promote a particular solution, 
-but to anchor the discussion of technical solutions to reproducibility
-challenges in concrete examples.
+I focus the remainder of this paper on one of the most recent and
+rapidly growing among these, called Docker, and the role it can play
+in reproducible research. Docker offers several promising features
+for reproducibility that go beyond the tools highlighted in [Clark
+2014]. Nevertheless, my goal in focusing on this technology is not to
+promote a particular solution, but to anchor the discussion of technical
+solutions to reproducibility challenges in concrete examples.
 
@@
 
-- Portable deployment: Docker defines a format for bundling and exchanging
-containers such that they run identically across platforms.
+- *Portable deployment*: Docker defines a format for bundling and
+exchanging containers such that they run identically across platforms.
 
-- Automatic builds: Docker defines a format (Dockerfiles) that provide
+- *Automatic builds*: Docker defines a format (Dockerfiles) that provide
 a simple, scripted way to construct containers, following the DevOpts
@@
 
-- Versioning: Docker provides git-like capibilities for tracking changes
-to a container, inspecting the history, metadata, or diff between
+- *Versioning*: Docker provides git-like capibilities for tracking
+changes to a container, inspecting the history, metadata, or diff between
 versions, committing versions, rolling back changes, and incremential
@@
 
-- Component reuse: Any container can be used as the base for any other.
+- *Component reuse*: Any container can be used as the base for any other.
 
-- Sharing: Docker provides a public registry (Docker Hub) where any
+- *Sharing*: Docker provides a public registry (Docker Hub) where any
 user can make images avaialble to any other. (The registry itself is
@@
 
+-----
+
+1. **"Dependency Hell"**
+2.  **Imprecise documentation**
+3.  **Code rot**
+4. **Barriers to adoption and reuse in existing solutions**
+
+------
+
 # Using Docker in an R environment
@@
 that it be easy to learn and fit relatively seamlessly into exsiting
-workflow patterns of domain researchers.  As we have seen, this, more
-than any other concern, can explain the relatively low uptake of previously
+workflow patterns of domain researchers.  As we have seen, this, more than
+any other concern, can explain the relatively low uptake of previously
 proposed solutions.  <!-- too strong? -->
 
-Being a new and unfamiliar tool to most domain scientists, Docker is far
-from immune to the same critique.  Nevertheless, Docker takes us several 
-key steps towards an approach that can be easily adopted in a research 
-context.
+Being a new and unfamiliar tool to most domain scientists, Docker is
+far from immune to the same critique.  Nevertheless, Docker takes us
+several key steps towards an approach that can be easily adopted in a
+research context.
 
@@
 
-In the R environment, the open source RStudio IDE provides another key component
-in making this system accessible to most domain scientists. Because the RStudio 
-IDE is written completely with standard web languages: javascript, css, and html,
-its web-based client, RStudio-Server, looks and feels identical to its popular 
-desktop IDE.  RStudio Server provides a way to interact with R on a remote environment
-without the latency, X tunnelling, etc typically involved in running R on a remote
-server.
+In the R environment, the open source RStudio IDE provides
+another key component in making this system accessible to most
+domain scientists. Because the RStudio IDE is written completely with
+standard web languages: javascript, css, and html, its web-based client,
+RStudio-Server, looks and feels identical to its popular desktop IDE.
+RStudio Server provides a way to interact with R on a remote environment
+without the latency, X tunnelling, etc typically involved in running R
+on a remote server.
 
-For our purposes, it provides users a way to interact with R, the filesystem,
-git, text editors, and graphics running on a Docker container.  Users already
-familiar with the popular IDE can thus benefit from the reproduciblity and 
-portability features provided by running R in a container environment without
-having to adapt to a new workflow for managing these things that is usually 
-necessary when moving to a cloud environment.
+For our purposes, it provides users a way to interact with R,
+the filesystem, git, text editors, and graphics running on a Docker
+container.  Users already familiar with the popular IDE can thus benefit
+from the reproduciblity and portability features provided by running R
+in a container environment without having to adapt to a new workflow
+for managing these things that is usually necessary when moving to a
+cloud environment.
 
-RStudio-Server is not the only way to provide a familiar developer environment
-to users working with a development environment running in a container on their
-local computer.  Docker supports the sharing of volumes (directories) between
-the container and the host OS.  This allows a user to rely on the familiar tools
-of the host OS for roles such as text editing, file browsing, or version control,
-while still allowing code execution to occur inside the controlled development 
-environment of the container. In this example, we launch an interactive R console
+RStudio-Server is not the only way to provide a familiar developer
+environment to users working with a development environment running in
+a container on their local computer.  Docker supports the sharing of
+volumes (directories) between the container and the host OS.  This allows
+a user to rely on the familiar tools of the host OS for roles such as
+text editing, file browsing, or version control, while still allowing
+code execution to occur inside the controlled development environment
+of the container. In this example, we launch an interactive R console
 in a container that is linked to our current working directory:
@@
 
-[^2]: External windows, such graphics windows cannot be opened directly from the container in this setup.
-Graphics would have to be saved to desk as raster or vector files and viewed on the host OS, until a better solution can be found.
-RStudio-server is one way around this.
+[^2]: External windows, such graphics windows cannot be opened directly
+from the container in this setup.  Graphics would have to be saved to
+desk as raster or vector files and viewed on the host OS, until a better
+solution can be found. RStudio-server is one way around this.
 
@@
 
-The challenge to reusing virtual machines can be summarized as "you can't install an image for every pipeline you want..." [Watson 2012].
-In contrast, this is exactly how Docker containers are designed to work.
-There are at least two ways in which Docker supports this kind
-of extensibility.
+The challenge to reusing virtual machines can be summarized as "you
+can't install an image for every pipeline you want..." [Watson 2012]. In
+contrast, this is exactly how Docker containers are designed to work.
+There are at least two ways in which Docker supports this kind of
+extensibility.
 
@@
 
-This acts like a software dependency; but unlike other software, a Dockerfile
-must have exactly one dependency (one `FROM` line). Note that a particular
-version of the dependency can be specified using the `:`.  We can in fact
-be much more precise, declaring not only version numbers like `ubuntu:14:04`,
-but specific cryptographic hashes that ensure we get the exact same image
-every time.
+This acts like a software dependency; but unlike other software, a
+Dockerfile must have exactly one dependency (one `FROM` line). Note that
+a particular version of the dependency can be specified using the `:`.
+We can in fact be much more precise, declaring not only version numbers
+like `ubuntu:14:04`, but specific cryptographic hashes that ensure we
+get the exact same image every time.
 
-We can now build this Dockerfile and give it name, by running in the working directory:
+We can now build this Dockerfile and give it name, by running in the
+working directory:
 
@@
 
-While Docker images tend to be much smaller than equivalent virtual machines, moving 
-around even 100s of gigabytes can be a challenge.  In their work on virtual machines,
-[Dudley 2014] recommend only running these tools on cloud servers such as the Amazon
-EC2 system. As most researchers still develop software locally and may not have ready
-access to these resources, such a requirement adds an additional barrier to reuse.
-Fortunately in the case of Docker we have a somewhat more elegant solution through
-the use of the Docker Hub [^1].  Docker provides a convenient way to share any image
-publicly or privately through the Hub after creating a free account.  Here, we share
-a public copy of the image we have just created by using the `docker push` command,
-followed by the name of the image:
+While Docker images tend to be much smaller than equivalent virtual
+machines, moving around even 100s of gigabytes can be a challenge.
+In their work on virtual machines, [Dudley 2014] recommend only
+running these tools on cloud servers such as the Amazon EC2 system. As
+most researchers still develop software locally and may not have ready
+access to these resources, such a requirement adds an additional barrier
+to reuse.  Fortunately in the case of Docker we have a somewhat more
+elegant solution through the use of the Docker Hub [^1].  Docker provides
+a convenient way to share any image publicly or privately through the
+Hub after creating a free account.  Here, we share a public copy of the
+image we have just created by using the `docker push` command, followed
+by the name of the image:
 
@@
 batteries included, with everything they need to run reproducibly.
-A researcher more familiar with python could connect our MySQL container
-to their python container, and so forth.
+A researcher could connect a container providing a computational 
+environment for a different language to this same MySQL container, 
+and so forth.
 
@@
 linking containers through this mechanism
+
+<!-- excerpted -->
 
@@
 'remix' problem of VMs.
-
-[Clark 2014]
 
@@
 
-It is possible, and indeed often desirable, to tackle the problem of dependencies
-at a much higher level than that of the operating system itself (e.g. [Ooms 2014]).
-For instance, it is often sufficient to address these concerns at the
-level of a particular scripting language rather than the level of the
-operating system. Most major scripting languages have dedicated tools for
-handling dependencies and versioning (e.g. Ruby gems, R packages, Python
-eggs, perl modules, javascript node package manager); with some languages
-having the fortune or misfortune to have multiple systems. Though systems
-differ somewhat in their strategies and their effectiveness, all are
-generally vulnerable to changes in their lower-level dependencies (such as
-compilers, C libraries, and so forth) that may differ between platforms.
-More pointedly, researchers whose work is not easily isolated to a single
-such environment but combines tools from different systems (including
-combining those tools used for formatting & visual display, such as
-LaTeX or pandoc, with scripting or analysis tools such as R).
+It is possible, and indeed often desirable, to tackle the problem of
+dependencies at a much higher level than that of the operating system
+itself (e.g. [Ooms 2014]).  For instance, it is often sufficient to
+address these concerns at the level of a particular scripting language
+rather than the level of the operating system. Most major scripting
+languages have dedicated tools for handling dependencies and versioning
+(e.g. Ruby gems, R packages, Python eggs, perl modules, javascript node
+package manager); with some languages having the fortune or misfortune to
+have multiple systems. Though systems differ somewhat in their strategies
+and their effectiveness, all are generally vulnerable to changes in their
+lower-level dependencies (such as compilers, C libraries, and so forth)
+that may differ between platforms.  More pointedly, researchers whose
+work is not easily isolated to a single such environment but combines
+tools from different systems (including combining those tools used for
+formatting & visual display, such as LaTeX or pandoc, with scripting or
+analysis tools such as R).
 
-<!-- Dynamic documents --> 
-Dynamic documents are another example of an approach that
+<!-- Dynamic documents --> Dynamic documents are another example of an
+approach that
 

### 2014-09-16 manuscript.Rmd
getting there. more writing

@@
 
-Docker is an open source (Apache Version 2) software platform written in
-the Go programming language.
+<!-- Docker is an open source (Apache Version 2) software platform written in
+the Go programming language.-->
 
-Docker builds on many long-familiar technologies from operating
-systems research: LXC containers, virtualization of the OS, and
-a hash-based or git-like verioning and differencing system, among
-others. Researchers familiar with LXC may be tempted to view the Docker
-system as little more than a creative application and marketing of an
-existing technology. However, the Docker approach adds several features
-crucial for its use in reproducible research (adapted from [Docker
-FAQ](https://docs.docker.com/faq/#what-does-docker-add-to-just-plain-lxc)):
+Docker is an open source project that builds on many long-familiar
+technologies from operating systems research: LXC containers,
+virtualization of the OS, and a hash-based or git-like verioning
+and differencing system, among others.
 
-- *Portable deployment*: Docker defines a format for bundling and
-exchanging containers such that they run identically across platforms.
+<!-- Define:
+- Dockerfile
+- Image 
+- Container: When we run a docker image, we get a container.
+We can run the same image multiple times, resulting in many identical containers.
+We can save the state of a container at any time, which creates a new image.
+- Dockerfile -> build -> image
+- Image -> run -> container
+-->
 
-- *Automatic builds*: Docker defines a format (Dockerfiles) that provide
-a simple, scripted way to construct containers, following the DevOpts
-philosophy.
+I introduce the most relevant concepts from Docker through the 
+context of the four challenges for reproducible research I have
+discussed above.
 
-- *Versioning*: Docker provides git-like capibilities for tracking
-changes to a container, inspecting the history, metadata, or diff between
-versions, committing versions, rolling back changes, and incremential
-uploads and downloads that send only the differences (like `git push`
-or `git pull`).
+### 1. Docker images: resolving 'Dependency Hell'
 
-- *Component reuse*: Any container can be used as the base for any other.
+A Docker based approach works similarly to a virtual machine image in
+addressing the dependency problem by providing other researchers with
+a binary image in which all the software has already been installed,
+configured and tested. (A machine image can also include all data files
+necessary for the research, which may simplify the distribution of data.)
 
-- *Sharing*: Docker provides a public registry (Docker Hub) where any
-user can make images avaialble to any other. (The registry itself is
-also open source so that it can be deployed privately).
+A key difference between Docker images and other virutal machines is 
+that the Docker images share the Linux kernel with the host machine.
+For the end user the primary consequence of this is that any Docker
+image must be based on a Linux system with Linux-compatible software, which
+includes (R, Python, Matlab, and most other scientific programming needs). [^4]
 
-Each of these elements have clear implications for reproducible research.
+Sharing the Linux kernel makes Docker much more light-weight and higher
+performance than complete virtual machines -- a typical desktop computer
+could run no more than a few virutal machines at once but 
+would have no trouble running 100s of Docker containers (a container
+is simply the term for running instance of an image). This feature
+has made Docker particularly attractive to industry and is largely
+responsible for the immense popularity of Docker. For our purposes this
+is a nice bonus, but the chief value to reproducible research lies
+in other aspects.
 
------
+[^4]: Note that re-distribution of an image in which proprietary software
+has been installed will be subject to any relevant licensing agreement.
 
-1. **"Dependency Hell"**
-2.  **Imprecise documentation**
-3.  **Code rot**
-4. **Barriers to adoption and reuse in existing solutions**
+<!-- 
+- image sharing (hub)
+- image linking
+-->
 
-------
+### 2. Dockerfiles: Resolving imprecise documentation
 
-# Using Docker in an R environment
+Though Docker images can be created interactively, this leaves little
+transparent record [^3] of what software has been installed and how.
+Dockerfiles provide a simple script (similar to a Makefile) that defines
+exactly how to build up the image, consistent with the DevOpts approach
+we mentioned previously.
+
+With a syntax that is simpler than other provisioning tools (e.g. Chef,
+Puppet, Ansible) or Continuous Integration (CI) platforms (e.g. Travis CI,
+Shippable CI); users need little more than a basic familarity with shell
+scripts and a Linux distribution software environment (e.g. Debian-based
+`apt-get`) to get started writing Dockerfiles.
+
+This approach has many advantages:
+
+- While machine images can be very large (many gigabytes), a Dockerfile
+is just a small plain text file that can be easily stored and shared.
+
+- Small plain text files are ideally suited for use with a version
+management system such as `subversion` or `git`, which can track any
+changes made to the `Dockerfile`
+
+- the `Dockerfile` provides a human readable summary of the necessary
+software dependencies, environmental variables and so forth needed to
+execute the code. There is little possiblity of the kind of holes or 
+imprecision in such a script that so frequently cause difficulty in
+manually implemented documentation of dependencies. This approach
+also avoids the burden of having to tediously document dependencies 
+at the end of a project, since they are instead documented as they 
+are installed by writing the `Dockerfile`.
+
+- Unlike a `Makefile` or other script, the `Dockerfile` includes all
+software dependencies down to the level of the OS, and is built by the
+Docker `build` tool, making it very unlikely that the resulting build
+will differ when being built on different machines. This is not to say
+that all builds of a Dockerfile are bitwise identical.  In particular,
+builds executed later will install more recent versions of the same
+software, if available, unless the package managers used are explicitly
+configured otherwise. We address this issue in the next section.
+
+- It is possible to add checks and tests following the commands for
+installing the software environment, which will verify that the setup
+has been successful.  This can be important in addressing the issue of
+code-rot which we discuss next.
+
+- It is straight-straight forward for other users to extend or customize 
+the resulting image by editing the script directly.
+
+[^3]: The situation is in fact slightly better than the virtual machine
+approach because these changes are versioned. Docker provides tools
+to inspect differences (diffs) between the images, and we can also roll
+back changes to earlier versions.
+
+### 3. Tackling code-rot with image versions
+
+As I have discussed above, changes to the dependencies, whether they are
+the result of security fixes, new features, or deprecation of old software,
+can break otherwise functioning code. These challenges can be significantly
+reduced because Docker defines the software environment to a particular operating system 
+and suite of libraries, such as the Ubuntu or Debian distribution.  Such 
+distributions use a staged release model with `stable`, `testing` and `unstable` 
+phases subjected to extensive testing to catch such potential problems [Ooms 2013],
+while also providing regular security updates to software within each stage.
+Nonetheless, this cannot completely avoid the challenge of code-rot,
+particularly when it is necessary to install software that is not (yet)
+available for a given distribution.
+
+To address this concern, one must archive a binary copy of the image
+used at the time the research was first performed.  Docker provides a
+simple utility to save an image as a portable `tarball` file that can
+be read in by any other Docker installation, providing a robust way to
+run the exact versions of all software involved.  By testing both the
+`tarball` archive and the image generated by the latest Dockerfile,
+Docker provides a simple way to confirm whether or not code rot has
+effected the function of a particular piece of code.
+
+<!-- FIXME move this later --> 
+To simplify this process, Docker also supports *Automated Builds*
+through the Docker Hub.  This acts as a kind of Continous Integration
+(CI) service that verifies the image builds correctly whenever the
+Dockerfile is updated, particularly if the Dockerfile includes checks
+for the environment. The Hub also provides a convenient distribution
+service, freely storing the prebuilt images, along with their metadata,
+for download and reuse by others. The Docker Hub is a free service and
+an open source software product so that users can run their own private
+versions of the Hub on their own servers, for instance, if security of
+the data or the longevity of the public platform is a concern.
+
+### 4. Barriers to adoption and re-use
+
+A technical solution, no matter how elegant, will be of little practical 
+use for reproducible research unless it is both easy to use and adapt to
+the existing workflow patterns of practicing domain researchers.
+
+Though most of the concerns we have discussed so far can be addressed
+through well-designed workflow software or the use of a DevOpts approach
+to provisioning virtual machines by scripts, neither approach has seen
+widespread adoption by domain researchers, who work primarily in a
+local rather than cloud-based environment using develoment tools native
+to their personal operating system. To gain more widespread adoption,
+reproducible research technologies must make it easier, not harder,
+for a researcher to perform the tasks they are already doing (before 
+considering any additional added benefits).
+
+These issues are reflected both during the original research or development
+phase and in any subsequent reuse. Another researcher may be less likely
+to build on existing work if it can only be done by using a particular
+workflow system or monolothic software platform with which they are
+unfamiliar.  Likewise, a user is more likey to make their own computational
+environment avialable for reuse if it does not involve a significant
+added effort in packaging and documenting [Stodden 2010].
+
+Though Docker is not immune to these challenges, it offers a interesting 
+example of a way forward in addressing these fundamental cocerns.
+
+- *Local environment*
+- *Modular reuse*
+
+<!---
+- The ability to work in a familiar, local environment (during one's own research or when re-using elements from another researcher)
+- An Docker container approach to both development and deployment
+- Image re-use & Image linking 
+-->
+
+# Using Docker as a local development environment
 
@@
 
+- Shared volumes
+- Host OS tools
+
 -----
@@
 domain scientists. Because the RStudio IDE is written completely with
-standard web languages: javascript, css, and html, its web-based client,
-RStudio-Server, looks and feels identical to its popular desktop IDE.
+standard web languages (javascript, css, and html), its web-based 
+RStudio Server looks and feels identical to its popular desktop IDE.
 RStudio Server provides a way to interact with R on a remote environment
@@
 
-- Interactive use, scripted use (Dockerfiles)
-- Reverting to a previous state
-- Integration with local development environment
-- Integration with cloud computing
-- Transparent & extensible
-- serial and linking containers
-
-- Best practices are still taking shape
-
-## Docker remix
+## Docker remix: Building re-usable modules
 
@@
 
-<!-- excerpted -->
+----------------
 
-1) Remix.  Titus has an excellent post, "Virtual Machines Considered
-Harmful for reproducibility" [1] , .  In contrast,
-Docker containers are designed to work exactly like that -- reusable
-building blocks you can link together with very little overhead in
-disk space or computation.  This more than anything else sets Docker
-apart from the standard VM approach.
+- Interactive use, scripted use (Dockerfiles)
+- Reverting to a previous state
+- Integration with local development environment
+- Integration with cloud computing
+- Transparent & extensible
+- serial and linking containers
+- Best practices are still taking shape
 
-Docker is not a virtual machine; containers are designed expressly to
-be remixable blocks.  You can put an R engine on one container and a
-mysql database on another and connect them. Docker philosophy aims at
-one function per container to maximize this reuse.  of course it's up
-to you to build this way rather than a single monolithic dockerfile,
-but the idea of linking containers is a technical concept at the heart
-of docker that offers a second and very different way to address the
-'remix' problem of VMs.
+----------
 
- Provisioning scripts. Docker images are not 'black boxes'. A
-"Dockerfile" is a simple make-like script which installs all the
-software necessary to re-create ("provision") the image.  This has
-many advantages: (a) The script is much smaller and more portable than
-the image. (b) the script can be version managed (c) the script gives
-a human readable (instead of binary) description of what software is
-installed and how. This also avoids pitfalls of traditional
-documentation of dependencies that may be too vague or out-of-sync.
-(d) Other users can build on, modify, or extend the script for their
-own needs.  All of this is what we call the he "DevOpts" approach to
-provisioning, and can be done with AMIs or other virtual machines
-using tools like Ansible, Chef, or Puppet coupled with things like
-Packer or Vagrant (or clever use of make and shell scripts).
+##  Best practices
 
-For a much better overview of this "DevOpts" approach in the
-reproducible research context and a gentle introduction to these
-tools, I highly recommend taking a look at Clark et al [2].
+- Using dockerfiles instead of containers
+- Using appropriate, common base images
+- Archiving tarball snapshots
+- Adding tests/checks to the Dockerfile
+- Sharing 
+- Leveraging package managers appropriately to handle software dependencies
+
+Upcoming features:
+
+- signed containers?
+
+-------------
 
@@
 
-Related challenges:  RStudio and a consistent environment. 
-- Linked volumes 
-- linked containers 
-- signed containers?
-
 4) Versioned images.  In addition to version managing the Dockerfile,
@@
 feature.
+
+--------------------------------
+
+<!-- Covered elsewhere in a more integrated fashion -->
+Researchers familiar with LXC may be tempted to view the Docker
+system as little more than a creative application and marketing of an
+existing technology. However, the Docker approach adds several features
+crucial for its use in reproducible research (adapted from [Docker
+FAQ](https://docs.docker.com/faq/#what-does-docker-add-to-just-plain-lxc)):
+
+- *Portable deployment*: Docker defines a format for bundling and
+exchanging containers such that they run identically across platforms.
+
+- *Automatic builds*: Docker defines a format (Dockerfiles) that provide
+a simple, scripted way to construct containers, following the DevOpts
+philosophy.
+
+- *Versioning*: Docker provides git-like capibilities for tracking
+changes to a container, inspecting the history, metadata, or diff between
+versions, committing versions, rolling back changes, and incremential
+uploads and downloads that send only the differences (like `git push`
+or `git pull`).
+
+- *Component reuse*: Any container can be used as the base for any other.
+
+- *Sharing*: Docker provides a public registry (Docker Hub) where any
+user can make images avaialble to any other. (The registry itself is
+also open source so that it can be deployed privately).
+
+Each of these elements have clear implications for reproducible research.
 

### 2014-09-16 manuscript.Rmd
writing

@@
 Reproducible research has recieved an increasing level of attention
-throughout the scientific community (e.g. [Peng 2011], [Nature editorial])
-and the public at large (e.g. [The Economist 2014]). All steps of the
+throughout the scientific community (e.g. [Peng 2011], [Nature 2014])
+and the public at large (e.g. [The Economist 2013]). All steps of the
 scientific process, from data collection and processing, to analyses,
@@
 
-<!-- FIXME more stuff on what this means-->
-
 <!-- Time savings. Personal reproducibility. Collaboration. Teaching. -->
@@
 4. Dynamic documents
-
 -->
@@
 environment avialable for reuse if it does not involve a significant
-added effort in packaging and documenting [Stodden 2010].
+added effort in packaging and documenting ([Stodden 2010]).
 
@@
 
-<!---
-- The ability to work in a familiar, local environment (during one's own research or when re-using elements from another researcher)
-- An Docker container approach to both development and deployment
-- Image re-use & Image linking 
--->
-
 # Using Docker as a local development environment
@@
 that it be easy to learn and fit relatively seamlessly into exsiting
-workflow patterns of domain researchers.  As we have seen, this, more than
-any other concern, can explain the relatively low uptake of previously
-proposed solutions.  <!-- too strong? -->
+workflow patterns of domain researchers.  As we have seen, this,
+more than any other concern, can explain the relatively low uptake
+of previously proposed solutions.  Being a new and unfamiliar tool to
+most domain scientists, Docker is far from immune to the same critique.
+Nevertheless, Docker takes us several key steps towards an approach that
+can be easily adopted in a research context.
 
-Being a new and unfamiliar tool to most domain scientists, Docker is
-far from immune to the same critique.  Nevertheless, Docker takes us
-several key steps towards an approach that can be easily adopted in a
-research context.
+While proponents of virtual machines for reproducible research propose
+that these machines would be available exclusively as cloud computing
+environments (e.g. [Dudley 2014]), many researchers work _locally_, that
+is, primarily with software that is installed on their laptop or desktop
+computer, and turning to cloud-based or other remote platforms only for
+certain collaborative tasks or when the work is mature enough to need
+increased computational power.  Working locally allows a researcher to
+rely more on the graphic interface tools for tasks such as managing
+files, text editing, debugging, IDEs, or interacting with version
+control systems.  Scientific computing on remote machines, by contrast,
+still relies largely on potentially less familiar text based command
+line functions for these tasks (though web based interfaces are rapidly
+filling this gap).
 
-Primary among these is the ability to run Docker locally on all major
-operating systems with minimal prior expertise.  On systems not already
-based on the Linux Kernel (such as Mac or Windows clients), this is
-accomplished through the use of a small VirtualBox-based VM running on
-the host OS called `boot2docker`.
+Docker can be easily installed on most major platforms (see
+[https://docs.docker.com/installation](https://docs.docker.com/installation);
+On systems not already based on the Linux Kernel, such as Mac or Windows,
+this is accomplished through the use of a small VirtualBox-based VM
+running on the host OS called `boot2docker`) and run locally.  Docker
+allows a user to link any directory (such as the working directory
+for a particular script or project) to the running Docker container.
+This allows a user to rely on the familiar tools of the host OS for
+roles such as text editing, file browsing, or version control, while
+still allowing code execution to occur inside the controlled development
+environment of the container.
 
-Local development offers several important advantages in encouraging
-adoption.
+For example, we launch an interactive R console in a container that
+is linked to our current working directory:
 
-- Shared volumes
-- Host OS tools
+```bash
+docker run -v $(pwd):/ -it eddelbuettel/docker-ubuntu-r /usr/bin/R
+```
 
------
+The resulting system behaves almost identically[^2] to running R on
+the command line of the host OS. (Clearly a similar command could be
+used just as well with interactive shells from other languages such as
+`ipython` or `irb`).
 
+[^2]: External windows, such graphics windows cannot be opened directly
+from the container in this setup.  Graphics would have to be saved to
+desk as raster or vector files and viewed on the host OS, until a better
+solution can be found. RStudio-server is one way around this.
+
+### Using RStudio Server
+
+An alternative approach for working locally with familar tools is to
+leverage web-based clients such as RStudio. 
 In the R environment, the open source RStudio IDE provides
@@
 RStudio Server looks and feels identical to its popular desktop IDE.
+<!--
 RStudio Server provides a way to interact with R on a remote environment
 without the latency, X tunnelling, etc typically involved in running R
-on a remote server.
+on a remote server. -->
 
-For our purposes, it provides users a way to interact with R,
+RStudio server provides users a way to interact with R,
 the filesystem, git, text editors, and graphics running on a Docker
@@
 from the reproduciblity and portability features provided by running R
-in a container environment without having to adapt to a new workflow
-for managing these things that is usually necessary when moving to a
-cloud environment.
+in a container environment without having to adapt to a new workflow.
 
-RStudio-Server is not the only way to provide a familiar developer
-environment to users working with a development environment running in
-a container on their local computer.  Docker supports the sharing of
-volumes (directories) between the container and the host OS.  This allows
-a user to rely on the familiar tools of the host OS for roles such as
-text editing, file browsing, or version control, while still allowing
-code execution to occur inside the controlled development environment
-of the container. In this example, we launch an interactive R console
-in a container that is linked to our current working directory:
+To accompany this paper, I provide a Docker image for running RStudio
+server, which can be launched as follows.  From the terminal (or
+boot2docker terminal on Mac or Windows client), run:
 
 ```bash
-docker run -v $(pwd):/ -it eddelbuettel/docker-ubuntu-r /usr/bin/R
+sudo docker run -d -p 8787:8787 cboettig/ropensci
 ```
 
-(Clearly a similar command could be used just as well with interactive
-shells from other languages such as `ipython` or `irb`). The resulting
-system behaves almost identically[^2] to running R on the command line
-of the host OS.
+That will take a while to download the image the first time you run
+it. boot2docker users will then need to determine the ip address assigned
+to boot2docker, while Linux users can just specify `http://localhost`
 
-[^2]: External windows, such graphics windows cannot be opened directly
-from the container in this setup.  Graphics would have to be saved to
-desk as raster or vector files and viewed on the host OS, until a better
-solution can be found. RStudio-server is one way around this.
+```bash
+boot2docker ip
+```
+
+Add the port `:8787` to the end of this address and paste
+it into your browser address bar, which should open to the
+RStudio welcome screen.  A user can now login with user/password
+`rstudio/rstudio`, run R scripts, install packages, use git,
+and so forth. User login and other configurations can be
+customized using environmental variables; see details at
+[https://github.com/ropensci/docker](https://github.com/ropensci/docker)
+
+### Portable computation
+
+A particular advantage of this approach is that the resulting
+computational environment is immediately _portable_.  This is useful not
+only for the purposes of reproducible research, where other users may seek
+to reconstruct the computational environment necessary to run the code,
+but is also of immediate value to the researcher themselves. For instance,
+a researcher might want to execute their code on a cloud server which
+has more memory or processing power then their local machine, or would
+want a co-author to help debug a particular problem.  In either case,
+the researcher can export a snapshot of their running container:
+
+```bash
+docker export <container-name> > container.tar
+```
+and then run this identical environment on the cloud or collaborators' machine.
 
@@
 
-First, Docker makes it immensely easy to build one container on top of
+First, Docker facilitates modular reuse by build one container on top of
 another.  Rather than copy a Dockerfile and then start adding more lines
@@
 
-<!-- too much detail? -->
+<!-- Transition -->
+
+## Versioning
+
+<!-- FIXME: Email text.  revise -->
+
+ In addition to version managing the Dockerfile,
+the images themselves are versioned using a git-like hash system
+(e.g. see `docker commit`, `docker push`/`docker pull`, `docker history`, 
+`docker diff`).  Docker images and containers have dedicated metadata specifying the date, author, parent
+image, and other details (see `docker inspect`).  We can roll back an image through the layers of history
+of its construction, then build off an earlier layer, or roll back changes
+we have made interactively in a container. For instance, here we can
+inspect recent changes made to the `ubuntu:latest` image:
+
+```bash
+docker history ubuntu:latest
+```
+
+We can identify an earlier version, and roll back to that version 
+just by adusting the docker tag to match the hash of that version. 
+For instance:
+
+```bash
+docker tag 25f ubuntu:latest
+```
+If we now inspect the history, we see it begins from this earlier poitn:
+
+```bash
+docker history ubuntu:latest
+```
+
+This same feature also means that docker can perform incremential
+uploads and downloads that send only the differences between images,
+(just like `git push` or `git pull` for git repositories), rather than
+transfer the full image each time.
+
+<!-- Step back, big picture -->
+
+----------------
+
+<!--
 The implementation of this would nonetheless depend on the individual 
@@
 to execute and link separate containers), and can be refactored into
-multiple containers when appropriate. It remains to be seen whether
-linking containers through this mechanism
-
-----------------
-
-- Interactive use, scripted use (Dockerfiles)
-- Reverting to a previous state
-- Integration with local development environment
-- Integration with cloud computing
-- Transparent & extensible
-- serial and linking containers
-- Best practices are still taking shape
-
-----------
+multiple containers when appropriate. 
+-->
 
@@
 -------------
-
-3) You can run the docker container *locally*.  I think this is huge.
-In my experience, most researchers do their primary development
-locally.  By running RStudio-server on your laptop, it isn't necessary
-for me to spin up an EC2 instance (with all the knowledge & potential
-cost that requires).  By sharing directories between Docker and the
-host OS, a user can still use everything they already know -- their
-favorite editor, moving files around with the native OS
-finder/browser, using all local configurations, etc, while still
-having the code execution occur in the container where the software is
-precisely specified and portable.  Whenever you need more power, you
-can then deploy the image on Amazon, DigitalOcean, a bigger desktop,
-your university HPC cluster, your favorite CI platform, or wherever
-else you want your code to run.  (On Mac & Windows, this uses
-something called boot2docker, and was not very seamless early on.  It
-has gotten much better and continues to improve.)
-
-4) Versioned images.  In addition to version managing the Dockerfile,
-the images themselves are versioned using a git-like hash system
-(check out: docker commit, docker push/pull, docker history, docker
-diff, etc).  They have metadata specifying the date, author, parent
-image, etc.  We can roll back an image through the layers of history
-of its construction, then build off an earlier layer.  This also
-allows docker to do all sorts of clever things, like avoiding
-downloading redundant software layers from the docker hub.  (If you
-pull a bunch of images that all build on ubuntu, you don't get n
-copies of ubuntu you have to download and store).  Oh yeah, and
-hosting your images on Docker hub is free (no need to pay for an S3
-bucket... for now?) and supports automated builds based on your
-dockerfiles, which acts as a kind of CI for your environment.
-Versioning and diff'ing images is a rather nice reproducibility
-feature.
-
---------------------------------
-
-<!-- Covered elsewhere in a more integrated fashion -->
-Researchers familiar with LXC may be tempted to view the Docker
-system as little more than a creative application and marketing of an
-existing technology. However, the Docker approach adds several features
-crucial for its use in reproducible research (adapted from [Docker
-FAQ](https://docs.docker.com/faq/#what-does-docker-add-to-just-plain-lxc)):
-
-- *Portable deployment*: Docker defines a format for bundling and
-exchanging containers such that they run identically across platforms.
-
-- *Automatic builds*: Docker defines a format (Dockerfiles) that provide
-a simple, scripted way to construct containers, following the DevOpts
-philosophy.
-
-- *Versioning*: Docker provides git-like capibilities for tracking
-changes to a container, inspecting the history, metadata, or diff between
-versions, committing versions, rolling back changes, and incremential
-uploads and downloads that send only the differences (like `git push`
-or `git pull`).
-
-- *Component reuse*: Any container can be used as the base for any other.
-
-- *Sharing*: Docker provides a public registry (Docker Hub) where any
-user can make images avaialble to any other. (The registry itself is
-also open source so that it can be deployed privately).
-
-Each of these elements have clear implications for reproducible research.
 
@@
 
-- NSF
-- rOpenSci
-- Dirk Eddlebuettel, Rich FitzJohn, Yihui Xie, Titus Brown, John Stanton-Geddes and others for helpful discussions about reproducibility, virtualization, and Docker that have helped shape this manuscript.
+CB acknowledges support from NSF grant DBI-1306697, and also from the Sloan Foundation support through the rOpenSci project.
+CB also wishes to thank Dirk Eddlebuettel, Rich FitzJohn, Yihui Xie, Titus Brown, John Stanton-Geddes and many others for helpful discussions about reproducibility, virtualization, and Docker that have helped shape this manuscript.
 
@@
 [Brown 2014]: http://cs.brown.edu/~sk/Memos/Examining-Reproducibility/
-
 [Vandewalle 2009]: http://doi.org/10.1109/MSP.2009.932122
@@
 [Clark 2014]: https://berkeley.app.box.com/s/w424gdjot3tgksidyyfl
-[Ooms 2013]: http://arxiv.org/abs/1303.2140v2
-[Ooms 2014]: http://arxiv.org/abs/1406.4806
+[Ooms 2013]: http://arxiv.org/abs/1303.2140v2 "Possible Directions for Improving Dependency Versioning in R"
+[Ooms 2014]: http://arxiv.org/abs/1406.4806 "The OpenCPU System: Towards a Universal Interface for Scientific Computing through Separation of Concerns"
 [Peng 2011]: http://doi.org/10.1126/science.1213847
@@
 [Stodden 2013]: http://www.davidhbailey.com/dhbpapers/icerm-report.pdf "ICERM Working group report: Setting the Default to Reproducible: Reproducibility in Computational and Experimental Mathematics"
+[Nature 2014]: http://doi.org/10.1038/483509a
+[The Economist 2013]: http://www.economist.com/news/leaders/21588069-scientific-research-has-changed-world-now-it-needs-change-itself-how-science-goes-wrong
+[van der Meulen 2008]: http://doi.org/10.1109/TSE.2008.70 "The effectiveness of software diversity in a large population of programs."
+[Watson 2012]: https://twitter.com/BioMickWatson/status/265037994526928896
 
-[Nature editorial]: http://doi.org/10.1038/483509a
-[Nature Special Issue]: http://www.nature.com/nature/focus/reproducibility/
+```{r include=FALSE}
+library("knitcitations")
+options(citation_format = "pandoc")
 
-[van der Meulen 2008]: http://doi.org/10.1109/TSE.2008.70 "The effectiveness of software diversity in a large population of programs."
+citep(c(
+"10.1371/journal.pone.0080278",
+"10.1126/science.1213847",
+"10.1038/467753a",
+"10.1038/nature10836",
+"10.1038/483509a",
+"10.1038/467775a",
+"10.1111/j.1365-294X.2012.05754.x",
+"10.1109/MSP.2009.932122",
+"10.1109/MC.2007.421",
+"10.1093/nar/gkl320"))
 
-[Watson 2012]: https://twitter.com/BioMickWatson/status/265037994526928896
+citep(c(
+"10.1109/SSDM.2004.1311241",
+"10.1109/MCSE.2012.62"))
+
+citep("10.1038/nbt1110-1181")
+
+citep("http://doi.org/10.3233/SJI-140818")
+
+citep(c(
+"10.2139/ssrn.1550193",
+"10.1109/TSE.2008.70"))
+
+citep(c(
+"http://arxiv.org/abs/1303.2140v2",
+"http://arxiv.org/abs/1406.4806",
+"https://storify.com/hlapp/reproducibility-repeatability-bigthink",
+"http://cs.brown.edu/~sk/Memos/Examining-Reproducibility/",
+"https://berkeley.app.box.com/s/w424gdjot3tgksidyyfl",
+"http://reproducibility.cs.arizona.edu/v1/tr.pdf",
+"http://www.davidhbailey.com/dhbpapers/icerm-report.pdf",
+"http://www.economist.com/news/leaders/21588069-scientific-research-has-changed-world-now-it-needs-change-itself-how-science-goes-wrong",
+"https://twitter.com/BioMickWatson/status/265037994526928896"))
+
+write.bibtex(file="components/references.bib")
+
+```
+
+<!-- Covered elsewhere in a more integrated fashion -->
+<!--
+
+Researchers familiar with LXC may be tempted to view the Docker
+system as little more than a creative application and marketing of an
+existing technology. However, the Docker approach adds several features
+crucial for its use in reproducible research (adapted from [Docker
+FAQ](https://docs.docker.com/faq/#what-does-docker-add-to-just-plain-lxc)):
+
+- *Portable deployment*: Docker defines a format for bundling and
+exchanging containers such that they run identically across platforms.
+
+- *Automatic builds*: Docker defines a format (Dockerfiles) that provide
+a simple, scripted way to construct containers, following the DevOpts
+philosophy.
+
+- *Versioning*: Docker provides git-like capibilities for tracking
+changes to a container, inspecting the history, metadata, or diff between
+versions, committing versions, rolling back changes, and incremential
+uploads and downloads that send only the differences (like `git push`
+or `git pull`).
+
+- *Component reuse*: Any container can be used as the base for any other.
+
+- *Sharing*: Docker provides a public registry (Docker Hub) where any
+user can make images avaialble to any other. (The registry itself is
+also open source so that it can be deployed privately).
+
+Each of these elements have clear implications for reproducible research.
+
+-->

### 2014-09-18 manuscript.Rmd
updated references, formatting to acm style, minor textual edits

@@
 ---
-layout: preprint
 title: "An introduction to Docker for reproducible research, with examples from the R environment"
@@
   - name: Carl Boettiger
-    affiliation: cstar
+    affiliation: 
+      l1: Center for Stock Assessment Research, 
+      l2: 110 Shaffer Rd, Santa Cruz, CA 95050, USA 
     email: cboettig(at)gmail.com
@@
 abstract: |
-  Docker is a promising solution to
+  As computational work becomes more and more integral to many aspects
+  of scientific research, computational reproducibility has become an
+  issue of increasing importance to computer systems researchers and
+  domain scientists alike. Though computational reproducibility seems 
+  more straight forward than replicating physical experiments, the 
+  immense, rapidly growing and frequently changing nature of computer
+  environments makes being able to reproduce work in a way that others
+  can understand and extend it a serious challenge.  In this paper, I
+  explore common reasons that code developed for a gien research project
+  cannot be successfully executed or extended by subsequent researchers.
+  I review current approaches to these issues, including virtual machines
+  and workflow systems, and their limitations. I then examine how the
+  popular emerging technology Docker combines several areas from systems
+  research - suc has operating system virtualization, cross-platform
+  portability, modular re-usable elements, versioning, and a 'DevOps'
+  philosophy, to address these challenges.  I illustrate this with several
+  examples of Docker use with a focus on the R statistical environment.
+
+Keywords:
+  - docker
+  - virtualization
+  - reproducibility
 
 bibliography: components/references.bib
-csl: components/ecology.csl
-documentclass: components/elsarticle
+csl: components/acm-sig-proceedings.csl
 
@@
   pdf_document:
-    template: components/elsarticle.latex
+    template: components/sigproc-sp.tex
     keep_tex: true
@@
 <!--
+There is increasing interest in promoting repeatable and reproducible research
+in computer science.  A community gains confidence in scientific claims when
+the experiments supporting those claims are published, examined, and repeated.
+Repeating research also helps a community expand on previous results and
+discover new directions for investigation.  Repetition requires that the
+artifacts used in an experiment be made available to, and made usable by,
+investigators.  In other words, artifacts must be shared.  Possible artifacts
+include not only binaries, source code, and datasets but also computing and
+networking infrastructure such as testbeds and clouds.
+
+Systems researchers face special challenges in the areas of repeatability and
+artifact sharing.  Some arise from the nature of systems software, which is
+typically complex and rife with internal and external dependencies.  Other
+issues arise from execution environments: for example, to reproduce performance
+and scalability results, one may need access to actual and specific hardware.
+Still more systems-specific challenges arise from the difficulty of measuring
+properties of interest, such as time and energy consumption.
 
@@
   * incentives for producing and sharing high-quality research artifacts
-
-Papers are expected to present significant results, insights, and/or directions
-for future work, and they are expected to include substantial material that has
-not already been published.
-
-SUBMISSION INSTRUCTIONS
-
-Submissions must be no longer than ten (10) pages; shorter papers are allowed
-and encouraged.  Submissions must be written in English and follow the standard
-formatting guidelines for papers appearing in Operating Systems Review (see
-<http://www.sigops.org/osr.html>).  Papers must be submitted in PDF format via
-the submission web site.  Submitted papers will be reviewed by the guest editor
-and the review committee.  Papers will be evaluated on technical quality,
-originality, relevance, and presentation.  Accepted papers will be published in
-the January 2015 issue of Operating Systems Review.
-
 -->
 
-Reproducible research has recieved an increasing level of attention
-throughout the scientific community (e.g. [Peng 2011], [Nature 2014])
-and the public at large (e.g. [The Economist 2013]). All steps of the
+# Introduction
+
+Reproducible research has received an increasing level of attention
+throughout the scientific community [e.g. @Peng_2011; @Nature_2012]
+and the public at large [e.g. @The_Economist_2013]. All steps of the
 scientific process, from data collection and processing, to analyses,
 visualizations and conclusions depend ever more on computation and
-algorithms, _computational reproducibility_ has recieved particular
-attention [Merali 2010].  Though in principle this algorithmic dependence
+algorithms, _computational reproducibility_ has received particular
+attention [@Merali_2010].  Though in principle this algorithmic dependence
 should make such research easier to reproduce -- computer codes being
@@
 
-# A cultural problem
+### Systems research & reproducibility
+
+Systems research has long concerned itself with the issues of
+computational reproducibility and the technologies that can facilitate
+those objectives [e.g. @Eide_2010, @Harji_2013]. Docker is a new but already
+very popular open source tool that combines many of these approaches
+in a user friendly implementation, including: (1) performing Linux
+container (LXC) based operating system (OS) level virtualization, (2)
+portable deployment of containers across platforms, (3) component reuse,
+(4) sharing, (5) archiving, and (6) versioning of container images. While
+Docker's market success has largely focused on the needs of businesses
+in deploying web applications and the potential for a lightweight
+alternative to full virtualization, these features have potentially
+important implications for systems research in the area of scientific
+reproducibility. In this paper, I seek to set these issues in the context
+of reproducibility throughout the various domain sciences where 
+computation plays an ever-increasing role, but where researchers are
+largely unaware of the concerns or technologies involved in making these
+computations more reproducible, extensible, and portable to other researchers. 
+In so doing, I highlight elements of the Docker platform that should be
+of interest to both computer systems research and domain scientists.
+
+<!-- Map here! -->
+
+### A cultural problem
 
@@
 reluctance to publish the code used to generating the results in the
-first place ([Barnes 2010]). Despite extensive evidence to the contrary
-(e.g. [Ince 2012]), many researchers and journals continue to assume that
+first place [@Barnes_2010]. Despite extensive evidence to the contrary
+[e.g. @Ince_2012], many researchers and journals continue to assume that
 summary descriptions or pseudo-code provide a sufficient description of
@@
 visualization, or analysis.  Until such code is available in the first
-place, we cannot even begin to encounter the problems that the approaches
+place, one cannot even begin to encounter the problems that the approaches
 discussed here set out to solve. As a result, few domain researchers
@@
 A lack of requirements or incentives no doubt plays a crucial role in
-discouraging sharing ([Barnes 2010], [Stodden 2013]). Nevertheless,
-it is easy to underestimate the siginficant barriers raised by a lack
+discouraging sharing [@Barnes_2010; @Stodden_2013]. Nevertheless,
+it is easy to underestimate the significant barriers raised by a lack
 of familiar, intuitive, and widely adopted tools for addressing the
@@
 that a lack of time, more than innate opposition to sharing, discourages
-researchers from providing code ([Stodden 2010], [FitzJohn 2014]).
+researchers from providing code [@Stodden_2010; @FitzJohn_2014].
 
-# The four technical challenges of computational reproducibility
+## Four technical challenges
 
 By restricting ourselves to studies of where code has been made
-available, we may sidestep for the moment the cultural challenges
-to reproducibility so that we may focus on the technical ones; in
+available, I will sidestep for the moment the cultural challenges
+to reproducibility so that I may focus on the technical ones; in
 particular, those challenges for which improved tools and techniques
@@
 barriers to reproducing the original results or building on that code
-(e.g. [Lapp 2014], [Garijo 2013], [Gilbert 2012], [Collberg 2014],
-[Brown 2014]), which I attempt to summarize as follows.
+[e.g. @Lapp_2014; @Garijo_2013; @Gilbert_2012; @Collberg_2014;
+@Krishnamurthi_2014], which I attempt to summarize as follows.
 
-<!-- Is this a permitted term? -->
-
-1. **"Dependency Hell"**
+### 1. **"Dependency Hell"**
 
@@
 less than 50% of software could even be successfully built or installed
-[Collberg 2014], and in an ongoing effort by other researchers to
-replicate that study [Brown 2014]. Installing or building software
+[@Collberg_2014], and in an ongoing effort by other researchers to
+replicate that study [@Krishnamurthi_2014]. Installing or building software
 necessary to run the code in question assumes the ability to recreate
 the computational environment of the original researchers. 
-
 <!-- Time savings. Personal reproducibility. Collaboration. Teaching. --> 
-
 <!-- Figure 1: technical reproducibility stack
@@
 -->
+<!-- Omit this? -->
 
-<!-- Omit this? -->
 Differences in numerical evaluation, such as arise in floating point
@@
 differing results between or even within the same computational platform
-([Ince 2012]). Such issues make it difficult to restrict the true dependencies
+[@Ince_2012]. Such issues make it difficult to restrict the true dependencies
 of the code to higher level environments such as that of a given scripting
-language, indpendent of the underlying OS or even hardware itself.
+language, independent of the underlying OS or even hardware itself.
 
-2.  **Imprecise documentation**
+### 2.  **Imprecise documentation**
 
 Documentation on how to install and run code associated with published
-research is another frequent barrier to replication.  [Lapp 2014] found
+research is another frequent barrier to replication.  @Lapp_2014 found
 this impairs a researcher's ability to install and build the software
 necessary, as even small holes in the documentation were found to be
-major barriers, particularly for "novices" (Sensu [Garijo 2013] --
-where novices may be experts in nearby langauages but unfamiliar with
+major barriers, particularly for "novices" [Sensu @Garijo_2013] --
+where novices may be experts in nearby languages but unfamiliar with
 the package managers and other tools of the language involved). This
-same problem is discussed in [Clark 2014].  Imprecise documentation
+same problem is discussed in [@Clark_2014].  Imprecise documentation
 goes well beyond issues of the software environment itself: incomplete
 documentation of parameters involved meant as few as 30% of analyses
-(n=34) using the popular software STRUCTURE could be reporoduced in the
-study of [Gilbert 2012].
+($n=34$) using the popular software STRUCTURE could be reproduced in the
+study of [@Gilbert_2012].
 
-3.  **Code rot**
+### 3.  **Code rot**
 
-Software dependencies are not static elements, but recieve regular
+Software dependencies are not static elements, but receive regular
 updates that may fix bugs, add new features or deprecate old features
@@
 as "code rot."  Researchers will want to know if the results are robust
-to the changes. The case studies in [Lapp 2014] provide examples of
+to the changes. The case studies in [@Lapp_2014] provide examples of
 these problems.
 
-4. **Barriers to adoption and reuse in existing solutions**
+### 4. **Barriers to adoption and reuse in existing solutions**
 
@@
 development would address many of the issues frequently frustrating
-reproducibility. However, researchers face siginificant barriers to entry
+reproducibility. However, researchers face significant barriers to entry
 in learning these tools and approaches which are not part of their typical
-curriculum, or lack incentives co-measurate with the effort required
-([Joppa 2013], [FitzJohn 2014]).
+curriculum, or lack incentives commensurate with the effort required
+[@Joppa_2013; @FitzJohn_2014].
 
@@
 
-<!-- A wide range of proposals, we focus on those at the operating system level -->
+<!-- A wide range of proposals, I focus on those at the operating system level -->
 Though a wide variety of approaches exists to work around these challenges,
-few operate on a low enough level to provide a general solution.  [Clark 2014]
+few operate on a low enough level to provide a general solution.  [@Clark_2014]
 provide an excellent description of this situation:
@@
 > In scientific computing the environment was commonly managed via
-Makfiles & Unix-y hacks, or alternatively with monolithic software
+Makefiles & Unix-y hacks, or alternatively with monolithic software
 like Matlab. More recently, centralized package management has provided
@@
 
+<!-- FIXME Doesn't say much about the approaches. Not clear what a workflow is... -->
+
 Two dominant paradigms have emerged to address these issues so far:
-workflow software (e.g. [Atintas 2004], [Hull 2006])  and
-cloud computing on virtual machines (e.g. [Dudley 2013], [Howe 2012]). 
+workflow software [e.g. @Altintas_2004; @Hull_2006]  and
+cloud computing on virtual machines [e.g. @Dudley_2010; @Howe_2012]. 
 Workflow software provides very elegant technical solutions to the
@@
 scientists, and can be very successful in the communities within which 
-they recieve substatianal adoption. Nonetheless, most workflow systems 
-struggle with relatively low total adoption overall ([Gin 2007, Dudley 2013]).
+they receive substantial adoption. Nonetheless, most workflow systems 
+struggle with relatively low total adoption overall [@Gil_2007; @Dudley_2010].
 
-[Dudley 2013] gives several reasons that such comprehensive workflow
+[@Dudley_2010] gives several reasons that such comprehensive workflow
 systems have not been more successful:
@@
 the computer Operating System (OS) already provides the software layer
-responsible for coodinating all the different elements running on the computer,
+responsible for coordinating all the different elements running on the computer,
 the VM approach captures the OS and everything running on it whole-cloth.
-To make this practical, [Dudley 2013] and [Howe 2012]
+To make this practical, [@Dudley_2010] and [@Howe_2012]
 propose using virtual machine images that will run on the cloud, such as Amazon's
@@
 Critics of the use of VMs to support reproducibility highlight that the 
-approach is too much of a black box and thus ill suited for reproducibility [Watson 2012].
+approach is too much of a black box and thus ill suited for reproducibility [@Watson_2012].
 While the approach sidesteps the need to either install or even document the
@@
 philosophy (rather than a particular tool), known as _Development_ and
-Systems _Operation_, or more frequently just "DevOpts." The approach is
+Systems _Operation_, or more frequently just "DevOps." The approach is
 characterized by scripting, rather than documenting, a description of the
 necessary dependencies for software to run, usually from the Operating
-System (OS) on up.  [Clark 2014] describes the DevOpts approach along
+System (OS) on up.  @Clark_2014 describes the DevOps approach along
 with both its relevance to reproducible research and examples of its
 use in the academic research context. They identify the difficulties
-we have discussed so far in terms of effective documentation:
+I have discussed so far in terms of effective documentation:
 
@@
 
-The authors contrast this to the DevOpts approach, where dependency 
+The authors contrast this to the DevOps approach, where dependency 
 documentation is _scripted_:
@@
 executes a script instead of manually managing the environmental setup)
-and comprehensiveness of implementation. [Clark 2014] are careful to note
+and comprehensiveness of implementation. @Clark_2014 are careful to note
 that this is not so much a technological shift as a philosophical one:
@@
 purpose have rapidly replaced the use of general purpose tools (such
-as Makefiles, bash scripts) to become synonomous with the "DevOpts"
-philosophy.  [Clark 2014] reviews many of these "DevOpts" tools, their
+as Makefiles, bash scripts) to become synonymous with the DevOps
+philosophy.  @Clark_2014 reviews many of these DevOps tools, their
 different roles, and their application in reproducible research.
@@
 in reproducible research. Docker offers several promising features
-for reproducibility that go beyond the tools highlighted in [Clark
-2014]. Nevertheless, my goal in focusing on this technology is not to
+for reproducibility that go beyond the tools highlighted in @Clark_2014.
+Nevertheless, my goal in focusing on this technology is not to
 promote a particular solution, but to anchor the discussion of technical
@@
 technologies from operating systems research: LXC containers,
-virtualization of the OS, and a hash-based or git-like verioning
+virtualization of the OS, and a hash-based or git-like versioning
 and differencing system, among others.
-
-<!-- Define:
-- Dockerfile
-- Image 
-- Container: When we run a docker image, we get a container.
-We can run the same image multiple times, resulting in many identical containers.
-We can save the state of a container at any time, which creates a new image.
-- Dockerfile -> build -> image
-- Image -> run -> container
--->
 
@@
 
-A key difference between Docker images and other virutal machines is 
+A key difference between Docker images and other virtual machines is 
 that the Docker images share the Linux kernel with the host machine.
@@
 performance than complete virtual machines -- a typical desktop computer
-could run no more than a few virutal machines at once but 
-would have no trouble running 100s of Docker containers (a container
+could run no more than a few virtual machines at once but 
+would have no trouble running 100's of Docker containers (a container
 is simply the term for running instance of an image). This feature
@@
 Dockerfiles provide a simple script (similar to a Makefile) that defines
-exactly how to build up the image, consistent with the DevOpts approach
-we mentioned previously.
+exactly how to build up the image, consistent with the DevOps approach
+I mentioned previously.
 
@@
 Puppet, Ansible) or Continuous Integration (CI) platforms (e.g. Travis CI,
-Shippable CI); users need little more than a basic familarity with shell
+Shippable CI); users need little more than a basic familiarity with shell
 scripts and a Linux distribution software environment (e.g. Debian-based
@@
 software dependencies, environmental variables and so forth needed to
-execute the code. There is little possiblity of the kind of holes or 
+execute the code. There is little possibility of the kind of holes or 
 imprecision in such a script that so frequently cause difficulty in
@@
 software, if available, unless the package managers used are explicitly
-configured otherwise. We address this issue in the next section.
+configured otherwise. I address this issue in the next section.
 
@@
 has been successful.  This can be important in addressing the issue of
-code-rot which we discuss next.
+code-rot which I discuss next.
 
@@
 approach because these changes are versioned. Docker provides tools
-to inspect differences (diffs) between the images, and we can also roll
+to inspect differences (diffs) between the images, and I can also roll
 back changes to earlier versions.
@@
 distributions use a staged release model with `stable`, `testing` and `unstable` 
-phases subjected to extensive testing to catch such potential problems [Ooms 2013],
+phases subjected to extensive testing to catch such potential problems [@Ooms_2013],
 while also providing regular security updates to software within each stage.
@@
 To simplify this process, Docker also supports *Automated Builds*
-through the Docker Hub.  This acts as a kind of Continous Integration
+through the Docker Hub.  This acts as a kind of Continuous Integration
 (CI) service that verifies the image builds correctly whenever the
@@
 for the environment. The Hub also provides a convenient distribution
-service, freely storing the prebuilt images, along with their metadata,
+service, freely storing the pre-built images, along with their metadata,
 for download and reuse by others. The Docker Hub is a free service and
@@
 
-Though most of the concerns we have discussed so far can be addressed
-through well-designed workflow software or the use of a DevOpts approach
+Though most of the concerns I have discussed so far can be addressed
+through well-designed workflow software or the use of a DevOps approach
 to provisioning virtual machines by scripts, neither approach has seen
 widespread adoption by domain researchers, who work primarily in a
-local rather than cloud-based environment using develoment tools native
+local rather than cloud-based environment using development tools native
 to their personal operating system. To gain more widespread adoption,
@@
 to build on existing work if it can only be done by using a particular
-workflow system or monolothic software platform with which they are
-unfamiliar.  Likewise, a user is more likey to make their own computational
-environment avialable for reuse if it does not involve a significant
-added effort in packaging and documenting ([Stodden 2010]).
+workflow system or monolithic software platform with which they are
+unfamiliar.  Likewise, a user is more likely to make their own computational
+environment available for reuse if it does not involve a significant
+added effort in packaging and documenting [@Stodden_2010].
 
 Though Docker is not immune to these challenges, it offers a interesting 
-example of a way forward in addressing these fundamental cocerns.
+example of a way forward in addressing these fundamental concerns.
 
@@
 
-# Using Docker as a local development environment
+## Using Docker as a local development environment
 
 Perhaps the most important feature of a reproducible research tool is
-that it be easy to learn and fit relatively seamlessly into exsiting
-workflow patterns of domain researchers.  As we have seen, this,
+that it be easy to learn and fit relatively seamlessly into existing
+workflow patterns of domain researchers.  This,
 more than any other concern, can explain the relatively low uptake
@@
 that these machines would be available exclusively as cloud computing
-environments (e.g. [Dudley 2014]), many researchers work _locally_, that
+environments [e.g. @Dudley_2010], many researchers work _locally_, that
 is, primarily with software that is installed on their laptop or desktop
@@
 
-For example, we launch an interactive R console in a container that
-is linked to our current working directory:
+For example, one launches an interactive R console in a container that
+is linked to our current working directory like so:
 
 ```bash
-docker run -v $(pwd):/ -it eddelbuettel/docker-ubuntu-r /usr/bin/R
+docker run -v $(pwd):/ -it cboettig/rstudio /usr/bin/R
 ```
@@
 
-An alternative approach for working locally with familar tools is to
+An alternative approach for working locally with familiar tools is to
 leverage web-based clients such as RStudio. 
@@
 domain scientists. Because the RStudio IDE is written completely with
-standard web languages (javascript, css, and html), its web-based 
+standard web languages (Javascript, CSS, and HTML), its web-based 
 RStudio Server looks and feels identical to its popular desktop IDE.
@@
 RStudio server provides users a way to interact with R,
-the filesystem, git, text editors, and graphics running on a Docker
+the file system, git, text editors, and graphics running on a Docker
 container.  Users already familiar with the popular IDE can thus benefit
-from the reproduciblity and portability features provided by running R
+from the reproducibility and portability features provided by running R
 in a container environment without having to adapt to a new workflow.
@@
 That will take a while to download the image the first time you run
-it. boot2docker users will then need to determine the ip address assigned
-to boot2docker, while Linux users can just specify `http://localhost`
+it. A boot2docker user will then need to determine the ip address assigned
+to boot2docker, while a Linux user can just specify `http://localhost`:
 
@@
 remixing elements.  To some extent this is already addressed by the
-"DevOpts" approach of Dockerfiles, providing a scripted description
+DevOps approach of Dockerfiles, providing a scripted description
 of the environment that can be tweaked and altered, but also includes
@@
 The challenge to reusing virtual machines can be summarized as "you
-can't install an image for every pipeline you want..." [Watson 2012]. In
+can't install an image for every pipeline you want..." [@Watson_2012]. In
 contrast, this is exactly how Docker containers are designed to work.
@@
 another.  Rather than copy a Dockerfile and then start adding more lines
-to the bottom, we can declare a new Dockerfile is built on an old one
+to the bottom, one can declare a new Dockerfile is built on an old one
 using the `FROM` directive. Here is an example Dockerfile that adds the
@@
 a particular version of the dependency can be specified using the `:`.
-We can in fact be much more precise, declaring not only version numbers
+One can in fact be much more precise, declaring not only version numbers
 like `ubuntu:14:04`, but specific cryptographic hashes that ensure we
@@
 
-We can now build this Dockerfile and give it name, by running in the
+One can now build this Dockerfile and give it name, by running in the
 working directory:
@@
 The key point here is that other researchers can easily build off this
-new image we have just created, extending our work directly, rather than
-having to go back to the original image.
+image just created, extending our work directly, rather than having
+to go back to the original image. Before others can re-use our image,
+it must be made available. Docker provides a convenient way to do this
+using the Docker Hub.
 
-However, before others can re-use our image we need a way to share it.
-
-### Docker hub
+### Docker Hub
 
 While Docker images tend to be much smaller than equivalent virtual
-machines, moving around even 100s of gigabytes can be a challenge.
-In their work on virtual machines, [Dudley 2014] recommend only
-running these tools on cloud servers such as the Amazon EC2 system. As
-most researchers still develop software locally and may not have ready
-access to these resources, such a requirement adds an additional barrier
-to reuse.  Fortunately in the case of Docker we have a somewhat more
-elegant solution through the use of the Docker Hub [^1].  Docker provides
-a convenient way to share any image publicly or privately through the
-Hub after creating a free account.  Here, we share a public copy of the
-image we have just created by using the `docker push` command, followed
-by the name of the image:
+machines, moving around even 100's of gigabytes can be a challenge.
+In their work on virtual machines, @Dudley_2010 recommend only running
+these tools on cloud servers such as the Amazon EC2 system. As most
+researchers still develop software locally and may not have ready access
+to these resources, such a requirement adds an additional barrier to
+reuse.  Docker provides a convenient way to share any image publicly
+or privately through the Hub after creating a free account.  For instance,
+one can  share a public copy of the image just created by using the
+`docker push` command, followed by the name of the image using the command:
 
@@
 Another user can now build directly on our image, rather than on the `ubuntu:latest`
-image we used as a base:
+image used as a base in the first Dockerfile.  To do so, one simply specifies a
+different image in the `FROM` directive of the Dockerfile, followed by the lines
+required to add any additional software required.
 
@@
 
-[^1]: Similar hubs are in fact available for full virtual machine images, such as the Vagrant hub.
-
-Though we have shown adding only a single piece of software in each step,
-clearly this approach can be particularly powerful in building up more complex
-environments.
+Though this shows only very minimal examples that add a single piece of
+software in each step, clearly this approach can be particularly powerful
+in building up more complex environments.
 
@@
 run one particular service or element, and exposing just what is necessary
-to link it together with other blocks.  For instance, we could have one
+to link it together with other blocks.  For instance, one could have one
 container running a PostgreSQL database which serves data to another container
@@
 
-<!-- FIXME: Email text.  revise -->
-
- In addition to version managing the Dockerfile,
-the images themselves are versioned using a git-like hash system
-(e.g. see `docker commit`, `docker push`/`docker pull`, `docker history`, 
-`docker diff`).  Docker images and containers have dedicated metadata specifying the date, author, parent
-image, and other details (see `docker inspect`).  We can roll back an image through the layers of history
-of its construction, then build off an earlier layer, or roll back changes
-we have made interactively in a container. For instance, here we can
-inspect recent changes made to the `ubuntu:latest` image:
+In addition to version managing the Dockerfile, the images themselves are
+versioned using a git-like hash system (e.g. see `docker commit`, `docker
+push`/`docker pull`, `docker history`, `docker diff`).  Docker images and
+containers have dedicated metadata specifying the date, author, parent
+image, and other details (see `docker inspect`). One can roll back an
+image through the layers of history of its construction, then build off
+an earlier layer, or roll back changes made interactively in a
+container. For instance, here I inspect recent changes made to the
+`ubuntu:latest` image:
 
@@
 
-We can identify an earlier version, and roll back to that version 
-just by adusting the docker tag to match the hash of that version. 
+One can identify an earlier version, and roll back to that version 
+just by adjusting the docker tag to match the hash of that version. 
 For instance:
@@
 ```
-If we now inspect the history, we see it begins from this earlier poitn:
+If one now inspects the history, which shows that it now begins from this earlier point:
 
@@
 
-This same feature also means that docker can perform incremential
+This same feature also means that docker can perform incremental
 uploads and downloads that send only the differences between images,
@@
 
+<!--
+
 ##  Best practices
 
-- Using dockerfiles instead of containers
+- Using Dockerfiles instead of containers
 - Using appropriate, common base images
@@
 
+-->
+
 # Further considerations
 
-## Combining virtualization with other reproducible-research tools
+### Combining virtualization with other reproducible-research tools
 
@@
 dependencies at a much higher level than that of the operating system
-itself (e.g. [Ooms 2014]).  For instance, it is often sufficient to
+itself [e.g. @Ooms_2014].  For instance, it is often sufficient to
 address these concerns at the level of a particular scripting language
@@
 languages have dedicated tools for handling dependencies and versioning
-(e.g. Ruby gems, R packages, Python eggs, perl modules, javascript node
+(e.g. Ruby gems, R packages, Python eggs, Perl modules, JavaScript node
 package manager); with some languages having the fortune or misfortune to
@@
 
-approach ([Leisch 2002], [Peng 2011], [Xie 2013]), which seek to embed
+approach [@Leisch_2002; @Peng_2011; @Xie_2013], which seek to embed
 the code required to re-generate the results within the manuscript
 itself. This process serves as a way to distribute computational
-code with publications, to ensure that methodological discriptions,
+code with publications, to ensure that methodological descriptions,
 parameter values, figures, and so forth remain synchronized with the code,
@@
 they inevitably increase the number of software dependencies involved,
-including those tools that support combining code with text (e.g. [Xie
-2013]) as well as tools merely for the handling the formating and layout
+including those tools that support combining code with text [e.g. @Xie_2013]
+as well as tools merely for the handling the formating and layout
 (e.g. LaTeX, pandoc).  Consequently, such a paper can fail to execute
@@
 implementing these dynamic features.  This problem is largely resolved by
-providing a consistent development environmment, making the approaches
+providing a consistent development environment, making the approaches
 discussed here a natural complement to that of dynamic documents.
 A similar argument can be made regarding the use of workflow systems,
-which can introduce additional softare dependencies even as they seek
+which can introduce additional software dependencies even as they seek
 to make it easier to work across them.
@@
 
-We noted at the outset that cultural expecations responsible for a
+I noted at the outset that cultural expectations responsible for a
 lack of code sharing practices in many fields are a far more extensive
-primary barrier to reproducibility than the technical barriers we have
+primary barrier to reproducibility than the technical barriers
 discussed here. Nevertheless, it may be worth considering how solutions
@@
 users to be able to install and run it in the first place is too high
-[Stodden 2010]. By lowering the barriers these barrier to re-use through
-the appropriate infrastructure first, we may also reduce certain cultural
+[@Stodden_2010]. Thus, lowering the barriers these barrier to re-use through
+the appropriate infrastructure may also reduce certain cultural
 barriers to sharing.
@@
 
-CB acknowledges support from NSF grant DBI-1306697, and also from the Sloan Foundation support through the rOpenSci project.
-CB also wishes to thank Dirk Eddlebuettel, Rich FitzJohn, Yihui Xie, Titus Brown, John Stanton-Geddes and many others for helpful discussions about reproducibility, virtualization, and Docker that have helped shape this manuscript.
+CB acknowledges support from NSF grant DBI-1306697, and also from the
+Sloan Foundation support through the rOpenSci project. CB also wishes
+to thank Dirk Eddelbuettel, Rich FitzJohn, Yihui Xie, Titus Brown,
+John Stanton-Geddes and many others for helpful discussions about
+reproducibility, virtualization, and Docker that have helped shape
+this manuscript.
 
@@
 
-[Garijo 2013]: http://doi.org/10.1371/journal.pone.0080278
-[Lapp 2014]: https://storify.com/hlapp/reproducibility-repeatability-bigthink
-[Gilbert 2012]: 10.1111/j.1365-294X.2012.05754.x
-[Collberg 2014]: http://reproducibility.cs.arizona.edu/v1/tr.pdf
-[Brown 2014]: http://cs.brown.edu/~sk/Memos/Examining-Reproducibility/
-[Vandewalle 2009]: http://doi.org/10.1109/MSP.2009.932122
-[Merali 2010]: http://doi.org/10.1038/467775a "Computational Science ... error: why scientific programming does not compute"
-[Gil 2007]: http://doi.org/10.1109/MC.2007.421 "Examining the Challenges of Scientific Workflows"
-[Hull 2006]: http://doi.org/10.1093/nar/gkl320 "Taverna: a tool for building and running workflows of services"
-[Altintas 2004]: http://doi.org/10.1109/SSDM.2004.1311241 "Kepler: an extensible system for design and execution of scientific workflows"
-[Howe 2012]: http://doi.org/10.1109/MCSE.2012.62
-[Dudley 2013]: http://doi.org/10.1038/nbt1110-1181 
-[Clark 2014]: https://berkeley.app.box.com/s/w424gdjot3tgksidyyfl
-[Ooms 2013]: http://arxiv.org/abs/1303.2140v2 "Possible Directions for Improving Dependency Versioning in R"
-[Ooms 2014]: http://arxiv.org/abs/1406.4806 "The OpenCPU System: Towards a Universal Interface for Scientific Computing through Separation of Concerns"
-[Peng 2011]: http://doi.org/10.1126/science.1213847
-[Barnes 2010]: http://doi.org/10.1038/467753a
-[Ince 2012]: http://doi.org/10.1038/nature10836
-[Stodden 2014]: http://doi.org/10.3233/SJI-140818
-[Stodden 2010]: http://doi.org/10.2139/ssrn.1550193
-[Stodden 2013]: http://www.davidhbailey.com/dhbpapers/icerm-report.pdf "ICERM Working group report: Setting the Default to Reproducible: Reproducibility in Computational and Experimental Mathematics"
-[Nature 2014]: http://doi.org/10.1038/483509a
-[The Economist 2013]: http://www.economist.com/news/leaders/21588069-scientific-research-has-changed-world-now-it-needs-change-itself-how-science-goes-wrong
-[van der Meulen 2008]: http://doi.org/10.1109/TSE.2008.70 "The effectiveness of software diversity in a large population of programs."
-[Watson 2012]: https://twitter.com/BioMickWatson/status/265037994526928896
+[@Lapp_2014]: https://storify.com/hlapp/reproducibility-repeatability-bigthink
+[@Gilbert_2012]: 10.1111/j.1365-294X.2012.05754.x
+[@Collberg_2014]: http://reproducibility.cs.arizona.edu/v1/tr.pdf
+[@Brown_2014]: http://cs.brown.edu/~sk/Memos/Examining-Reproducibility/
+[@Vandewalle_2009]: http://doi.org/10.1109/MSP.2009.932122
+[@Merali_2010]: http://doi.org/10.1038/467775a "Computational Science ... error: why scientific programming does not compute"
+[@Gil_2007]: http://doi.org/10.1109/MC.2007.421 "Examining the Challenges of Scientific Workflows"
+[@Hull_2006]: http://doi.org/10.1093/nar/gkl320 "Taverna: a tool for building and running workflows of services"
+[@Altintas_2004]: http://doi.org/10.1109/SSDM.2004.1311241 "Kepler: an extensible system for design and execution of scientific workflows"
+[@Howe_2012]: http://doi.org/10.1109/MCSE.2012.62
+[@Dudley_2010]: http://doi.org/10.1038/nbt1110-1181 
+[@Clark_2014]: https://berkeley.app.box.com/s/w424gdjot3tgksidyyfl
+[@Ooms_2013]: http://arxiv.org/abs/1303.2140v2 "Possible Directions for Improving Dependency Versioning in R"
+[@Ooms_2014]: http://arxiv.org/abs/1406.4806 "The OpenCPU System: Towards a Universal Interface for Scientific Computing through Separation of Concerns"
+[@Peng_2011]: http://doi.org/10.1126/science.1213847
+[@Barnes_2010]: http://doi.org/10.1038/467753a
+[@Ince_2012]: http://doi.org/10.1038/nature10836
+[@Stodden_2014]: http://doi.org/10.3233/SJI-140818
+[@Stodden_2010]: http://doi.org/10.2139/ssrn.1550193
+[@Stodden_2013]: http://www.davidhbailey.com/dhbpapers/icerm-report.pdf "ICERM Working group report: Setting the Default to Reproducible: Reproducibility in Computational and Experimental Mathematics"
+[@Nature_2012]: http://doi.org/10.1038/483509a
+[@The_Economist_2013]: http://www.economist.com/news/leaders/21588069-scientific-research-has-changed-world-now-it-needs-change-itself-how-science-goes-wrong
+[@van_der_Meulen_2008]: http://doi.org/10.1109/TSE.2008.70 "The effectiveness of software diversity in a large population of programs."
+[@Watson_2012]: https://twitter.com/BioMickWatson/status/265037994526928896
 
-```{r include=FALSE}
+```{r include=FALSE, eval=FALSE}
 library("knitcitations")
 options(citation_format = "pandoc")
-
 citep(c(
@@
 "10.1109/MC.2007.421",
-"10.1093/nar/gkl320"))
-
-citep(c(
+"10.1093/nar/gkl320",
 "10.1109/SSDM.2004.1311241",
-"10.1109/MCSE.2012.62"))
-
-citep("10.1038/nbt1110-1181")
-
+"10.1109/MCSE.2012.62",
+"10.1038/nbt1110-1181"))
 citep("http://doi.org/10.3233/SJI-140818")
-
 citep(c(
 "10.2139/ssrn.1550193",
-"10.1109/TSE.2008.70"))
-
-citep(c(
+"10.1109/TSE.2008.70",
 "http://arxiv.org/abs/1303.2140v2",

### 2014-09-18 manuscript.Rmd
writing

@@
 Though Docker is not immune to these challenges, it offers a interesting 
-example of a way forward in addressing these fundamental concerns.
-
-- *Local environment*
-- *Modular reuse*
+example of a way forward in addressing these fundamental concerns. Here 
+I highlight five of these features in turn:
+
+- Local environment
+- Modular reuse
+- Portable environments 
+- public repository for sharing
+- Versioning
 
@@
 
-Perhaps the most important feature of a reproducible research tool is
-that it be easy to learn and fit relatively seamlessly into existing
-workflow patterns of domain researchers.  This,
-more than any other concern, can explain the relatively low uptake
-of previously proposed solutions.  Being a new and unfamiliar tool to
-most domain scientists, Docker is far from immune to the same critique.
-Nevertheless, Docker takes us several key steps towards an approach that
-can be easily adopted in a research context.
+Perhaps the most important feature of a reproducible research tool is that
+it be easy to learn and fit relatively seamlessly into existing workflow
+patterns of domain researchers.  This, more than any other concern,
+can explain the relatively low uptake of previously proposed solutions.
+Being a new and unfamiliar tool to most domain scientists, Docker is
+far from immune to the same critique.  Nevertheless, Docker takes us
+several key steps towards an approach that can be easily adopted in a
+research context.
 
@@
 
-### Portable computation
+## Portable computation & sharing
 
 A particular advantage of this approach is that the resulting
-computational environment is immediately _portable_.  This is useful not
-only for the purposes of reproducible research, where other users may seek
-to reconstruct the computational environment necessary to run the code,
-but is also of immediate value to the researcher themselves. For instance,
-a researcher might want to execute their code on a cloud server which
-has more memory or processing power then their local machine, or would
-want a co-author to help debug a particular problem.  In either case,
-the researcher can export a snapshot of their running container:
+computational environment is immediately _portable_. LXC containers
+by themselves are unlikely to run in the same way, if at all, across
+different machines, due to differences in networking, storage,
+logging and so forth. Docker handles the packaging and execution of
+a container so that it works identically across different machines,
+while exposing the necessary interfaces for networking ports, volumes,
+and so forth.  This is useful not only for the purposes of reproducible
+research, where other users may seek to reconstruct the computational
+environment necessary to run the code, but is also of immediate value
+to the researcher themselves. For instance, a researcher might want to
+execute their code on a cloud server which has more memory or processing
+power then their local machine, or would want a co-author to help debug a
+particular problem.  In either case, the researcher can export a snapshot
+of their running container:
 
@@
 ```
+
 and then run this identical environment on the cloud or collaborators' machine.
 
-## Docker remix: Building re-usable modules
+Sharing these images is further facilitated by the Docker Hub technology. 
+While Docker images tend to be much smaller than equivalent virtual
+machines, moving around even 100's of gigabytes can be a challenge.
+In their work on virtual machines, @Dudley_2010 recommend only running
+these tools on cloud servers such as the Amazon EC2 system. As most
+researchers still develop software locally and may not have ready access
+to these resources, such a requirement adds an additional barrier to
+reuse.  Docker provides a convenient way to share any image publicly
+or privately through the Hub after creating a free account.  Docker Hub
+is both a publicly available service and also available as 
+a separate open source platform that can be deployed on private servers.
+
+One can  share a public copy of the image just created by using the
+`docker push` command, followed by the name of the image using the command:
+
+```bash
+docker push username/r-recommended
+```
+
+If a Dockerfile is made available on a public code repository such as
+[Github](https://github.com) or [Bitbucket](https://bitbucket.org),
+the Hub can automatically build the image whenever a change is made to
+the Dockerfile, making the `push` command unnecessary. A user can update
+their local image using the `docker pull <imagename>`, which downloads
+any changes that have since been made to the copy of the image on the Hub.
+
+## Re-usable modules
 
@@
 image just created, extending our work directly, rather than having
-to go back to the original image. Before others can re-use our image,
-it must be made available. Docker provides a convenient way to do this
-using the Docker Hub.
-
-### Docker Hub
-
-While Docker images tend to be much smaller than equivalent virtual
-machines, moving around even 100's of gigabytes can be a challenge.
-In their work on virtual machines, @Dudley_2010 recommend only running
-these tools on cloud servers such as the Amazon EC2 system. As most
-researchers still develop software locally and may not have ready access
-to these resources, such a requirement adds an additional barrier to
-reuse.  Docker provides a convenient way to share any image publicly
-or privately through the Hub after creating a free account.  For instance,
-one can  share a public copy of the image just created by using the
-`docker push` command, followed by the name of the image using the command:
-
-```bash
-docker push username/r-recommended
-```
-
-Another user can now build directly on our image, rather than on the `ubuntu:latest`
-image used as a base in the first Dockerfile.  To do so, one simply specifies a
-different image in the `FROM` directive of the Dockerfile, followed by the lines
-required to add any additional software required.
+to go back to the original image.
+
+If this image is shared (either directly or through a Docker Hub),
+another user can now build directly on our image, rather than on the
+`ubuntu:latest` image used as a base in the first Dockerfile.  To do
+so, one simply specifies a different image in the `FROM` directive of
+the Dockerfile, followed by the lines required to add any additional
+software required.
 
@@
 
-----------------
-
-<!--
-The implementation of this would nonetheless depend on the individual 
-researcher's willingness to break things into multiple containers, as
-it is indeed possible to run everything in a monolithic container. 
-This is not necessarily bad, as it simplifies the workflow (no need
-to execute and link separate containers), and can be refactored into
-multiple containers when appropriate. 
--->
-
-<!--
-
-##  Best practices
-
-- Using Dockerfiles instead of containers
-- Using appropriate, common base images
-- Archiving tarball snapshots
-- Adding tests/checks to the Dockerfile
-- Sharing 
-- Leveraging package managers appropriately to handle software dependencies
+# Conclusions
+
+## Best Practices
+
+The effectiveness of this approach for supporting reproducible research
+nonetheless depends on how each of these features are adopted and 
+implemented by individual researchers. I summarize a few of the choices
+
+- Use docker containers during development
+
+A key feature of the Docker approach is the ability to mimic as closely
+as possible the current workflow and development practices of the user.
+Code executing inside a container on a local machine can appear identical
+to code running natively, but with the added benefit that one can simply
+recreate or snapshot and share the entire computational environment with
+a few simple commands.  This works best if researchers set up their computational
+environment in a container from the outset of the project.
+
+- Write Dockerfiles instead installing interactive sessions.
+
+As we have noted already, Docker can be used in a purely interactive
+manner to record and distribute changes to a computational environment.
+However, the approach is most useful for reproducible research when
+researchers begin by defining their environment explicitly in the DevOps
+fashion by writing a Dockerfile.
+
+- Adding tests or checks to the Dockerfile
+
+Dockerfile commands need not be limited to installing software, but
+can also include execution.  This can help verify that an image has 
+build successfully with all the software necessary to run the 
+research code of interest.
+
+- Use and provide appropriate base images
+
+Though Docker supports modular design, it remains up to the researchers
+to take advantage of it. An appropriate workflow might involve one
+Dockerfile that includes all the software dependencies a researcher
+usually uses in the course of their development, which can then be
+extended by separate docker images for particular projects.  Re-using
+existing images reduces the effort required to set up an environment,
+contributes to the standardization of computational environments within
+a field, and best leverages the ability of Docker's distribution system
+to download only differences.
+
+- Share Docker images and Dockerfiles
+
+The Docker Hub significantly reduces the barriers for making even
+large images (which can exceed the file size limits of journals common
+scientific data repositories such as Dryad and Figshare) readily available
+to other researchers.
+
+- Archive tarball snapshots
+
+Despite similar semantics to git, Docker's versioning system works rather 
+differently than version management of code.  Docker can roll back AUFS
+layers that have been added to an image, but not revert to the earlier 
+state of a particular layer.  In consequence, to preserve a bitwise identical
+snapshot of a container used to generate a given set of results, it is
+necessary to archive the image tarball itself -- one can not simply
+rely on the docker history to recover an earlier state.
+
+## Limitations and future developments
+
+Docker has the potential to address shortcomings of certain existing
+approaches to reproducible research challenges that stem from recreating
+complex computational environments.  Docker also provides a promising
+case study in other issues. Its versioning, modular design, portable
+containers, and simple interface have proven successful in industry and
+could have promising implications for reproducible research in scientific
+communities. Nonetheless, these advances raise questions and challenges 
+of their own.
+
+Docker does not provide complete virtualization but relies on the 
+Linux kernel provided by the host.  Systems research can provide insight
+on what limitations to reproducibility this introduces [e.g.  @Harji_2013].
+
+Docker is limited to 64 bit host machines, making it impossible to run on
+older hardware.
+
+On Mac and Windows machines Docker must still be run in a fully
+virtualized environment. Though the boot2docker tool streamlines this
+process, it remains to be seen if the performance and integration with
+the host machine's OS is sufficiently seamless or creates a barrier to
+adoption by users on of these systems.
+
+It remains to be seen if Docker will be significantly adopted by any scientific community.
 
@@
 
--------------
-
--->
-
-# Further considerations
+## Further considerations
+
+<!-- Is this really necessary? -->
 
@@
 
-<!-- Dynamic documents --> Dynamic documents are another example of an
-approach that
-
-approach [@Leisch_2002; @Peng_2011; @Xie_2013], which seek to embed
+<!-- Dynamic documents --> 
+Dynamic documents are another example of an approach that to reproducible
+research that can be facilitated through the use of Docker containers. 
+Dynamic Documents [e.g. @Leisch_2002; @Peng_2011; @Xie_2013] seek to embed
 the code required to re-generate the results within the manuscript

### 2014-09-19 docker-paper/manuscript.Rmd
separate dir for manuscript

--- 2014-09-18 manuscript.Rmd
@@
 abstract: |
-  As computational work becomes more and more integral to many aspects
-  of scientific research, computational reproducibility has become an
-  issue of increasing importance to computer systems researchers and
-  domain scientists alike. Though computational reproducibility seems 
-  more straight forward than replicating physical experiments, the 
-  immense, rapidly growing and frequently changing nature of computer
-  environments makes being able to reproduce work in a way that others
-  can understand and extend it a serious challenge.  In this paper, I
-  explore common reasons that code developed for a gien research project
+  As computational work becomes more and more integral to many aspects of
+  scientific research, computational reproducibility has become an issue
+  of increasing importance to computer systems researchers and domain
+  scientists alike. Though computational reproducibility seems more
+  straight forward than replicating physical experiments, the complex
+  and rapidly changing nature of computer environments makes being able
+  to reproduce and extend such work a serious challenge.  In this paper,
+  I explore common reasons that code developed for one research project
   cannot be successfully executed or extended by subsequent researchers.
@@
   popular emerging technology Docker combines several areas from systems
-  research - suc has operating system virtualization, cross-platform
+  research - such as operating system virtualization, cross-platform
   portability, modular re-usable elements, versioning, and a 'DevOps'
@@
 ```bash
-docker export <container-name> > container.tar
+docker export container-name > container.tar
 ```
@@
 FROM ubuntu:latest
-RUN apt-get update && apt-get -y install r-recommended
+RUN apt-get update 
+RUN apt-get -y install r-recommended
 ```
@@
 ```bash
-docker build -t r-recommended .
+docker build -t username/r-recommended .
 ```
@@
 FROM username/r-recommended 
-RUN apt-get update && apt-get -y install r-cran-teachingdemos
+RUN apt-get update
+RUN apt-get -y install r-cran-matrix
 ```
@@
 nonetheless depends on how each of these features are adopted and 
-implemented by individual researchers. I summarize a few of the choices
-
-- Use docker containers during development
-
-A key feature of the Docker approach is the ability to mimic as closely
-as possible the current workflow and development practices of the user.
-Code executing inside a container on a local machine can appear identical
-to code running natively, but with the added benefit that one can simply
-recreate or snapshot and share the entire computational environment with
-a few simple commands.  This works best if researchers set up their computational
+implemented by individual researchers. I summarize a few of these practices here:
+
+- _Use docker containers during development_. A key feature of the Docker
+approach is the ability to mimic as closely as possible the current
+workflow and development practices of the user.  Code executing inside
+a container on a local machine can appear identical to code running
+natively, but with the added benefit that one can simply recreate or
+snapshot and share the entire computational environment with a few simple
+commands.  This works best if researchers set up their computational
 environment in a container from the outset of the project.
 
-- Write Dockerfiles instead installing interactive sessions.
-
-As we have noted already, Docker can be used in a purely interactive
-manner to record and distribute changes to a computational environment.
-However, the approach is most useful for reproducible research when
-researchers begin by defining their environment explicitly in the DevOps
-fashion by writing a Dockerfile.
-
-- Adding tests or checks to the Dockerfile
-
-Dockerfile commands need not be limited to installing software, but
-can also include execution.  This can help verify that an image has 
-build successfully with all the software necessary to run the 
-research code of interest.
-
-- Use and provide appropriate base images
-
-Though Docker supports modular design, it remains up to the researchers
-to take advantage of it. An appropriate workflow might involve one
-Dockerfile that includes all the software dependencies a researcher
-usually uses in the course of their development, which can then be
-extended by separate docker images for particular projects.  Re-using
-existing images reduces the effort required to set up an environment,
-contributes to the standardization of computational environments within
-a field, and best leverages the ability of Docker's distribution system
-to download only differences.
-
-- Share Docker images and Dockerfiles
-
-The Docker Hub significantly reduces the barriers for making even
-large images (which can exceed the file size limits of journals common
-scientific data repositories such as Dryad and Figshare) readily available
-to other researchers.
-
-- Archive tarball snapshots
-
-Despite similar semantics to git, Docker's versioning system works rather 
-differently than version management of code.  Docker can roll back AUFS
-layers that have been added to an image, but not revert to the earlier 
-state of a particular layer.  In consequence, to preserve a bitwise identical
-snapshot of a container used to generate a given set of results, it is
-necessary to archive the image tarball itself -- one can not simply
-rely on the docker history to recover an earlier state.
+- _Write Dockerfiles instead installing interactive sessions_. As we
+have noted already, Docker can be used in a purely interactive manner to
+record and distribute changes to a computational environment.  However,
+the approach is most useful for reproducible research when researchers
+begin by defining their environment explicitly in the DevOps fashion by
+writing a Dockerfile.
+
+- _Adding tests or checks to the Dockerfile_. Dockerfile commands need
+not be limited to installing software, but can also include execution.
+This can help verify that an image has build successfully with all the
+software necessary to run the research code of interest.
+
+- _Use and provide appropriate base images_. Though Docker supports
+modular design, it remains up to the researchers to take advantage of
+it. An appropriate workflow might involve one Dockerfile that includes
+all the software dependencies a researcher usually uses in the course of
+their development, which can then be extended by separate docker images
+for particular projects.  Re-using existing images reduces the effort
+required to set up an environment, contributes to the standardization
+of computational environments within a field, and best leverages the
+ability of Docker's distribution system to download only differences.
+
+- _Share Docker images and Dockerfiles_. The Docker Hub significantly
+reduces the barriers for making even large images (which can exceed the
+file size limits of journals common scientific data repositories such
+as Dryad and Figshare) readily available to other researchers.
+
+- _Archive tarball snapshots_. Despite similar semantics to git, Docker's
+versioning system works rather differently than version management
+of code.  Docker can roll back AUFS layers that have been added to
+an image, but not revert to the earlier state of a particular layer.
+In consequence, to preserve a bitwise identical snapshot of a container
+used to generate a given set of results, it is necessary to archive the
+image tarball itself -- one can not simply rely on the docker history
+to recover an earlier state.
 
@@
 
-Docker does not provide complete virtualization but relies on the 
+- Docker does not provide complete virtualization but relies on the 
 Linux kernel provided by the host.  Systems research can provide insight
@@
 
-Docker is limited to 64 bit host machines, making it impossible to run on
+- Docker is limited to 64 bit host machines, making it impossible to run on
 older hardware.
 
-On Mac and Windows machines Docker must still be run in a fully
+- On Mac and Windows machines Docker must still be run in a fully
 virtualized environment. Though the boot2docker tool streamlines this
@@
 
-It remains to be seen if Docker will be significantly adopted by any scientific community.
+- Most importantly, it remains to be seen if Docker will be significantly
+adopted by any scientific research or teaching community.
+
+<!--
 
@@
 
+-->
+
 ## Further considerations
 
-<!-- Is this really necessary? -->
+<!-- Is this really necessary? Can it be more consise? -->
 
@@
 Sloan Foundation support through the rOpenSci project. CB also wishes
-to thank Dirk Eddelbuettel, Rich FitzJohn, Yihui Xie, Titus Brown,
-John Stanton-Geddes and many others for helpful discussions about
-reproducibility, virtualization, and Docker that have helped shape
+to thank Scott Chamberlain, Dirk Eddelbuettel, Rich FitzJohn, Yihui Xie,
+Titus Brown, John Stanton-Geddes and many others for helpful discussions
+about reproducibility, virtualization, and Docker that have helped shape
 this manuscript.

### 2014-09-19 docker-paper/manuscript.Rmd
tweaking

@@
 Keywords:
-  - docker
+  - Docker
   - virtualization
@@
 Reproducible research has received an increasing level of attention
-throughout the scientific community [e.g. @Peng_2011; @Nature_2012]
-and the public at large [e.g. @The_Economist_2013]. All steps of the
+throughout the scientific community [_e.g._ @Peng_2011; @Nature_2012]
+and the public at large [_e.g._ @The_Economist_2013]. All steps of the
 scientific process, from data collection and processing, to analyses,
@@
 computational reproducibility and the technologies that can facilitate
-those objectives [e.g. @Eide_2010, @Harji_2013]. Docker is a new but already
+those objectives [_e.g._ @Eide_2010, @Harji_2013]. Docker is a new but already
 very popular open source tool that combines many of these approaches
@@
 first place [@Barnes_2010]. Despite extensive evidence to the contrary
-[e.g. @Ince_2012], many researchers and journals continue to assume that
+[_e.g._ @Ince_2012], many researchers and journals continue to assume that
 summary descriptions or pseudo-code provide a sufficient description of
@@
 barriers to reproducing the original results or building on that code
-[e.g. @Lapp_2014; @Garijo_2013; @Gilbert_2012; @Collberg_2014;
+[_e.g._ @Lapp_2014; @Garijo_2013; @Gilbert_2012; @Collberg_2014;
 @Krishnamurthi_2014], which I attempt to summarize as follows.
@@
 Two dominant paradigms have emerged to address these issues so far:
-workflow software [e.g. @Altintas_2004; @Hull_2006]  and
-cloud computing on virtual machines [e.g. @Dudley_2010; @Howe_2012]. 
+workflow software [_e.g._ @Altintas_2004; @Hull_2006]  and
+cloud computing on virtual machines [_e.g._ @Dudley_2010; @Howe_2012]. 
 Workflow software provides very elegant technical solutions to the
@@
 
-<!-- Link on how to install docker -->
-<!-- Examples here are not meant as a tutorial to docker, but 
+<!-- Link on how to install Docker -->
+<!-- Examples here are not meant as a tutorial to Docker, but 
 rather to illustrate concepts a reader should be able to follow whether 
-or not they have any previous knowledge of docker and its commands -->
+or not they have any previous knowledge of Docker and its commands -->
 
@@
 
-With a syntax that is simpler than other provisioning tools (e.g. Chef,
-Puppet, Ansible) or Continuous Integration (CI) platforms (e.g. Travis CI,
+With a syntax that is simpler than other provisioning tools (_e.g._ Chef,
+Puppet, Ansible) or Continuous Integration (CI) platforms (_e.g._ Travis CI,
 Shippable CI); users need little more than a basic familiarity with shell
-scripts and a Linux distribution software environment (e.g. Debian-based
+scripts and a Linux distribution software environment (_e.g._ Debian-based
 `apt-get`) to get started writing Dockerfiles.
@@
 that these machines would be available exclusively as cloud computing
-environments [e.g. @Dudley_2010], many researchers work _locally_, that
+environments [_e.g._ @Dudley_2010], many researchers work _locally_, that
 is, primarily with software that is installed on their laptop or desktop
@@
 Docker can be easily installed on most major platforms (see
-[https://docs.docker.com/installation](https://docs.docker.com/installation);
+[https://docs.Docker.com/installation](https://docs.docker.com/installation);
 On systems not already based on the Linux Kernel, such as Mac or Windows,
@@
 customized using environmental variables; see details at
-[https://github.com/ropensci/docker](https://github.com/ropensci/docker)
+[https://github.com/ropensci/Docker](https://github.com/ropensci/docker)
 
@@
 One can  share a public copy of the image just created by using the
-`docker push` command, followed by the name of the image using the command:
+`Docker push` command, followed by the name of the image using the command:
 
@@
 the Dockerfile, making the `push` command unnecessary. A user can update
-their local image using the `docker pull <imagename>`, which downloads
+their local image using the `Docker pull <imagename>`, which downloads
 any changes that have since been made to the copy of the image on the Hub.
@@
 ```bash
-docker build -t username/r-recommended .
+Docker build -t username/r-recommended .
 ```
@@
 In addition to version managing the Dockerfile, the images themselves are
-versioned using a git-like hash system (e.g. see `docker commit`, `docker
-push`/`docker pull`, `docker history`, `docker diff`).  Docker images and
+versioned using a git-like hash system (_e.g._ see `Docker commit`, `docker
+push`/`Docker pull`, `docker history`, `docker diff`).  Docker images and
 containers have dedicated metadata specifying the date, author, parent
-image, and other details (see `docker inspect`). One can roll back an
+image, and other details (see `Docker inspect`). One can roll back an
 image through the layers of history of its construction, then build off
@@
 One can identify an earlier version, and roll back to that version 
-just by adjusting the docker tag to match the hash of that version. 
+just by adjusting the Docker tag to match the hash of that version. 
 For instance:
@@
 
-This same feature also means that docker can perform incremental
+This same feature also means that Docker can perform incremental
 uploads and downloads that send only the differences between images,
@@
 
-- _Use docker containers during development_. A key feature of the Docker
+- _Use Docker containers during development_. A key feature of the Docker
 approach is the ability to mimic as closely as possible the current
@@
 all the software dependencies a researcher usually uses in the course of
-their development, which can then be extended by separate docker images
+their development, which can then be extended by separate Docker images
 for particular projects.  Re-using existing images reduces the effort
@@
 used to generate a given set of results, it is necessary to archive the
-image tarball itself -- one can not simply rely on the docker history
+image tarball itself -- one can not simply rely on the Docker history
 to recover an earlier state.
@@
 Linux kernel provided by the host.  Systems research can provide insight
-on what limitations to reproducibility this introduces [e.g.  @Harji_2013].
+on what limitations to reproducibility this introduces [_e.g._  @Harji_2013].
 
@@
 
-It is possible, and indeed often desirable, to tackle the problem of
-dependencies at a much higher level than that of the operating system
-itself [e.g. @Ooms_2014].  For instance, it is often sufficient to
-address these concerns at the level of a particular scripting language
-rather than the level of the operating system. Most major scripting
-languages have dedicated tools for handling dependencies and versioning
-(e.g. Ruby gems, R packages, Python eggs, Perl modules, JavaScript node
-package manager); with some languages having the fortune or misfortune to
-have multiple systems. Though systems differ somewhat in their strategies
-and their effectiveness, all are generally vulnerable to changes in their
-lower-level dependencies (such as compilers, C libraries, and so forth)
-that may differ between platforms.  More pointedly, researchers whose
-work is not easily isolated to a single such environment but combines
-tools from different systems (including combining those tools used for
-formatting & visual display, such as LaTeX or pandoc, with scripting or
-analysis tools such as R).
-
-<!-- Dynamic documents --> 
-Dynamic documents are another example of an approach that to reproducible
-research that can be facilitated through the use of Docker containers. 
-Dynamic Documents [e.g. @Leisch_2002; @Peng_2011; @Xie_2013] seek to embed
-the code required to re-generate the results within the manuscript
-itself. This process serves as a way to distribute computational
-code with publications, to ensure that methodological descriptions,
-parameter values, figures, and so forth remain synchronized with the code,
-and to provide a way for other researchers to both tweak the code and
-'execute' the paper to reproduce the results.  While these features
-are very desirable from the perspective of reproducible research,
-they inevitably increase the number of software dependencies involved,
-including those tools that support combining code with text [e.g. @Xie_2013]
-as well as tools merely for the handling the formating and layout
-(e.g. LaTeX, pandoc).  Consequently, such a paper can fail to execute
-successfully for reasons having nothing to do with the computations
-involved in the research, but rather due to software responsible for
-implementing these dynamic features.  This problem is largely resolved by
-providing a consistent development environment, making the approaches
-discussed here a natural complement to that of dynamic documents.
-A similar argument can be made regarding the use of workflow systems,
-which can introduce additional software dependencies even as they seek
-to make it easier to work across them.
+Using Docker containers to distribute reproducible research should be
+seen as an approach that is synergistic with, rather than an alternative
+to, other technical tools for ensuring computational reproducibility.
+Existing tools for managing dependencies for a particular language [R
+packages, Ruby gems, etc, see @Ooms_2014] can easily be employed within a
+Docker-based approach, allowing the operating-systems level virtualization
+to sidestep potential issues such as external library dependencies or
+conflicts with existing user libraries.  Other approaches that facilitate
+reproducible research also introduce additional software dependencies
+and possible points of failure [@FitzJohn_2014]. One example includes
+dynamic documents [_e.g._ @Leisch_2002; @Peng_2011; @Xie_2013] which embed
+the code required to re-generate the results within the manuscript.  As a
+result, it is necessary to package the appropriate typesetting libaries
+(_e.g._ \LaTeX) along with the code libaries such that the document executes
+successfully for different researchers and platforms.
 
@@
 users to be able to install and run it in the first place is too high
-[@Stodden_2010]. Thus, lowering the barriers these barrier to re-use through
+[@Stodden_2010]. Thus, lowering these barrier to re-use through
 the appropriate infrastructure may also reduce certain cultural
@@
 crucial for its use in reproducible research (adapted from [Docker
-FAQ](https://docs.docker.com/faq/#what-does-docker-add-to-just-plain-lxc)):
+FAQ](https://docs.Docker.com/faq/#what-does-docker-add-to-just-plain-lxc)):
 

### 2014-09-19 docker-paper/manuscript.Rmd
more minor edits

@@
 
-That will take a while to download the image the first time you run
-it. A boot2docker user will then need to determine the ip address assigned
-to boot2docker, while a Linux user can just specify `http://localhost`:
+That will take a while to download the image the first time you run it. A
+boot2docker user will then need to determine the ip address assigned
+to boot2docker as shown below, while a Linux user can just specify
+`http://localhost`.
 
@@
 Dockerfile must have exactly one dependency (one `FROM` line). Note that
-a particular version of the dependency can be specified using the `:`.
+a particular version of the dependency can be specified using the `:` notation.
 One can in fact be much more precise, declaring not only version numbers
@@
 to, other technical tools for ensuring computational reproducibility.
-Existing tools for managing dependencies for a particular language [R
+Existing tools for managing dependencies for a particular language [_e.g._ R
 packages, Ruby gems, etc, see @Ooms_2014] can easily be employed within a
@@
 # References
+
+<!-- These links don't appear in either pdf or rendered version, but provide some automatic linking on the Github rendering 
+It would be nice if pandoc's reference syntax were closer to markdown's standard link syntax (e.g. always use [], bracket each reference)
+-->
 

### 2014-09-19 docker-paper/manuscript.Rmd
tweaking

@@
 Documentation on how to install and run code associated with published
-research is another frequent barrier to replication.  @Lapp_2014 found
+research is another frequent barrier to replication. A study by Lapp [@Lapp_2014] found
 this impairs a researcher's ability to install and build the software
@@
 necessary dependencies for software to run, usually from the Operating
-System (OS) on up.  @Clark_2014 describes the DevOps approach along
+System (OS) on up. Clark _et al._ [@Clark_2014] describe the DevOps approach along
 with both its relevance to reproducible research and examples of its
@@
 executes a script instead of manually managing the environmental setup)
-and comprehensiveness of implementation. @Clark_2014 are careful to note
+and comprehensiveness of implementation. Clark _et al._ [@Clark_2014] are careful to note
 that this is not so much a technological shift as a philosophical one:
@@
 as Makefiles, bash scripts) to become synonymous with the DevOps
-philosophy.  @Clark_2014 reviews many of these DevOps tools, their
+philosophy.  Clark _et al._ [@Clark_2014] reviews many of these DevOps tools, their
 different roles, and their application in reproducible research.
@@
 in reproducible research. Docker offers several promising features
-for reproducibility that go beyond the tools highlighted in @Clark_2014.
+for reproducibility that go beyond the tools highlighted in [@Clark_2014].
 Nevertheless, my goal in focusing on this technology is not to
@@
 machines, moving around even 100's of gigabytes can be a challenge.
-In their work on virtual machines, @Dudley_2010 recommend only running
+In their work on virtual machines, Dudley _et al._ [@Dudley_2010] recommend only running
 these tools on cloud servers such as the Amazon EC2 system. As most

### 2014-09-19 docker-paper/manuscript.Rmd
tweaking reference formatting

@@
 Though a wide variety of approaches exists to work around these challenges,
-few operate on a low enough level to provide a general solution.  [@Clark_2014]
+few operate on a low enough level to provide a general solution. Clark _et al._ [@Clark_2014]
 provide an excellent description of this situation:
@@
 
-[@Dudley_2010] gives several reasons that such comprehensive workflow
+Dudley & Butte [@Dudley_2010] give several reasons that such comprehensive workflow
 systems have not been more successful:
@@
 the VM approach captures the OS and everything running on it whole-cloth.
-To make this practical, [@Dudley_2010] and [@Howe_2012]
+To make this practical, Dudley & Butte [@Dudley_2010] and Howe [@Howe_2012] both
 propose using virtual machine images that will run on the cloud, such as Amazon's

### 2014-09-23 docker-paper/manuscript.Rmd
final edits

@@
 to do with the technological approaches discussed here, but stems rather from a
-reluctance to publish the code used to generating the results in the
+reluctance to publish the code used in generating the results in the
 first place [@Barnes_2010]. Despite extensive evidence to the contrary
@@
 less than 50% of software could even be successfully built or installed
-[@Collberg_2014], and in an ongoing effort by other researchers to
+[@Collberg_2014] and similar results are seen in an ongoing effort by other researchers to
 replicate that study [@Krishnamurthi_2014]. Installing or building software
@@
 Differences in numerical evaluation, such as arise in floating point
-arithmetic or computers or even ambiguities in standardized programming
+arithmetic or even ambiguities in standardized programming
 languages ("order-of-evaluation" problems) can be responsible for
@@
 where novices may be experts in nearby languages but unfamiliar with
-the package managers and other tools of the language involved). This
+the package managers and other tools of the language involved. This
 same problem is discussed in [@Clark_2014].  Imprecise documentation
@@
 workflow software [_e.g._ @Altintas_2004; @Hull_2006]  and
-cloud computing on virtual machines [_e.g._ @Dudley_2010; @Howe_2012]. 
+virtual machines [_e.g._ @Dudley_2010; @Howe_2012]. 
 Workflow software provides very elegant technical solutions to the
 challenges of communication between diverse software tools, capturing
-provenance in graphically driven interfaces, and handling issues from of
+provenance in graphically driven interfaces, and handling issues from 
 versioning dependencies to data access.  Workflow solutions are often 
@@
 <!-- Scripted provisioning: a solution to the weaknesses of virtual machines -->
-## A "DevOpts" approach
+## A "DevOps" approach
 
@@
 Sharing the Linux kernel makes Docker much more light-weight and higher
-performance than complete virtual machines -- a typical desktop computer
+performing than complete virtual machines -- a typical desktop computer
 could run no more than a few virtual machines at once but 
@@
 
-- It is straight-straight forward for other users to extend or customize 
+- It is straightforward for other users to extend or customize 
 the resulting image by editing the script directly.
@@
 To simplify this process, Docker also supports *Automated Builds*
-through the Docker Hub.  This acts as a kind of Continuous Integration
+through the Docker Hub ([hub.docker.com](http://hub.docker.com)).  This acts as a kind of Continuous Integration
 (CI) service that verifies the image builds correctly whenever the
@@
 - Portable environments 
-- public repository for sharing
+- Public repository for sharing
 - Versioning
@@
 still relies largely on potentially less familiar text based command
-line functions for these tasks (though web based interfaces are rapidly
-filling this gap).
+line functions for these tasks (though web based interfaces like RStudio
+Server are rapidly filling this gap).
 
@@
 from the container in this setup.  Graphics would have to be saved to
-desk as raster or vector files and viewed on the host OS, until a better
-solution can be found. RStudio-server is one way around this.
+disk as raster or vector files and viewed on the host OS, until a better
+solution can be found. RStudio Server ([rstudio.com/products/rstudio/#Server](http://rstudio.com/products/rstudio/#Server)) is one way around this.
 
@@
 RStudio Server looks and feels identical to its popular desktop IDE.
-<!--
 RStudio Server provides a way to interact with R on a remote environment
-without the latency, X tunnelling, etc typically involved in running R
-on a remote server. -->
+without the latency, X tunnelling, and so forth typically involved in running R
+on a remote server.
 
@@
 A researcher could connect a container providing a computational 
-environment for a different language to this same MySQL container, 
+environment for a different language to this same PostgreSQL container, 
 and so forth.
@@
 
-- _Write Dockerfiles instead installing interactive sessions_. As we
+- _Write Dockerfiles instead of installing interactive sessions_. As we
 have noted already, Docker can be used in a purely interactive manner to
@@
 - Docker is limited to 64 bit host machines, making it impossible to run on
-older hardware.
+older hardware (at this time).
 
@@
 
+- Potential computer security issues may still need to be evaluated. Among other
+changes, future support for digitally signing Docker images may make it easier
+to build off of only trusted binaries.
+
 - Most importantly, it remains to be seen if Docker will be significantly
 adopted by any scientific research or teaching community.
-
-<!--
-
-Upcoming features:
-
-- signed containers?
-
--->
 
@@
 technical barriers to reuse provides immediate practical benefits
-that make this exchanged into a more balanced, two-way street. Another
+that make this exchange into a more balanced, two-way street. Another
 concern is that the difficulty imposed in preparing code to be shared,
@@
 users to be able to install and run it in the first place is too high
-[@Stodden_2010]. Thus, lowering these barrier to re-use through
+[@Stodden_2010]. Thus, lowering these barriers to re-use through
 the appropriate infrastructure may also reduce certain cultural
@@
 - *Automatic builds*: Docker defines a format (Dockerfiles) that provide
-a simple, scripted way to construct containers, following the DevOpts
+a simple, scripted way to construct containers, following the DevOps
 philosophy.

### 2014-09-30 docker-paper/manuscript.Rmd
updating to reflect rocker work, etc

@@
     footnote: Corresponding author
-address: 
-  - code: cstar
-    address: | 
-      Center for Stock Assessment Research,
-      110 Shaffer Rd,
-      Santa Cruz, CA 95050, USA 
 abstract: |

### 2014-09-30 rstudio/userconf.sh

@@ -1,9 +1,0 @@
-#!/bin/bash
-
-useradd -m $USER && echo "$USER:$PASSWORD" | chpasswd
-git config --global user.name $USER
-git config --global user.email $EMAIL
-git config --global credential.helper cache
-cp /.gitconfig /home/$USER
-rsync -a /tmp/ /home/$USER/
-chown -R $USER:$USER /home/$USER

### 2014-09-30 setup.sh

@@ -1,0 +1,46 @@
+#!/bin/bash
+
+## Stay updated 
+sudo apt-get -q update && sudo apt-get -qy dist-upgrade
+
+
+## UFW -- Uncomplicated firewall configuration
+sudo apt-get install ufw
+
+
+## Docker needs to ufw to allow forwarding and port 2375
+sudo sed -i s/DEFAULT_FORWARD_POLICY="DROP"/DEFAULT_FORWARD_POLICY="ACCEPT" /etc/default/ufw
+sudo ufw reload
+sudo ufw allow 2375/tcp
+
+## Allow services: 
+sudo ufw allow $DRONE/tcp
+sudo ufw allow $SSH/tcp
+sudo ufw allow $SSH-CONTAINER/tcp
+sudo ufw allow $GITLAB-SSH/tcp
+sudo ufw allow $GITLAB-WEB/tcp
+
+
+sudo apt-get install fail2ban 
+sudo cp /etc/fail2ban/jail.conf /etc/fail2ban/jail.local
+# adjust [ssh] port section to monitor non-standard $SSH port in use.
+
+## Requires interactive work
+sudo apt-get install -y tripwire 
+
+
+
+# Allocate some swap space for a digitalocean image
+sudo fallocate -l 1G /swapfile
+sudo chmod 600 /swapfile
+sudo mkswap /swapfile
+sudo swapon /swapfile
+sudo echo "/swapfile       none    swap    sw      0       0" >> /etc/fstab
+
+# Redundant? 
+sudo chown root:root /swapfile
+sudo chmod 0600 /swapfile
+
+
+# Install docker
+curl -sSL https://get.docker.io/ubuntu/ | sudo sh

### 2014-11-04 setup.sh
update makefile, setup

@@ -132,10 +132,10 @@ apt-get install -y \
 
 ###########################################
 
-## Install nsenter:
-docker run -v /usr/local/bin:/target jpetazzo/nsenter
+## Install nsenter: -- not needed now, docker 1.3 has docker exec -it <containerid> bash
+# docker run -v /usr/local/bin:/target jpetazzo/nsenter
 ## alias appropriately
-echo 'function dock { sudo nsenter -m -u -n -i -p -t `docker inspect --format {{.State.Pid}} "$1"` /bin/bash; }' >> ~/.bashrc
+# echo 'function dock { sudo nsenter -m -u -n -i -p -t `docker inspect --format {{.State.Pid}} "$1"` /bin/bash; }' >> ~/.bashrc
 
 
 # Launch containerized services: 

### 2014-11-04 docker-paper/manuscript.Rmd
addressed some of the minor edits from reviewers, see issue #1

@@
 ---
-title: "An introduction to Docker for reproducible research, with examples from the R environment"
+title: "An introduction to Docker for reproducible research"
 author: 
@@
 
-<!--
+<!-- Excerpt from the CALL FOR PAPERS:
 There is increasing interest in promoting repeatable and reproducible research
@@
 important implications for systems research in the area of scientific
-reproducibility. In this paper, I seek to set these issues in the context
-of reproducibility throughout the various domain sciences where 
-computation plays an ever-increasing role, but where researchers are
-largely unaware of the concerns or technologies involved in making these
-computations more reproducible, extensible, and portable to other researchers. 
-In so doing, I highlight elements of the Docker platform that should be
-of interest to both computer systems research and domain scientists.
-
-<!-- Map here! -->
+reproducibility.
+
+In this paper, I seek to address two audiences. First, that of the domain
+scientist, those conducting research in ecology, bioinformatics, economics,
+psychology and so many other disciplines in which computation plays an 
+ever-increasing role. I seek to help this audience become more aware of the 
+concerns and challenges in making these computations more reproducible, extensible, 
+and portable to other researchers, and highlight Docker as platform to address 
+these challenges. The second audience is that of the computer systems researchers:
+readers familiar with both the challenges to reproducibility (versioning,
+software dependencies, etc.) and the technical components underlying Docker (virtualization,
+LXC containers, jails, hashes, etc.), but who may be unfamiliar with Docker software. 
+I hope the domain scientist will be motivated to try Docker as a tool to address
+reproducibility. Meanwhile, I expect the systems researcher to see Docker technology as
+an area ripe for further systems research (e.g. to what extent can differences between 
+Linux kernels frustrate reproducibility in a Docker image?) rather than as a tool
+for their own study (because Docker applies only to software at the level above
+the Linux kernel, it will be of little use in making studies of differences in 
+kernel behavior, hardware performance, execution times, and so forth reproducible.)
 
@@
 cannot easily build on the virtual machine in a consistent and scalable way. 
-If each study provided it's own virtual machine, any pipeline combining the 
+If each study provided its own virtual machine, any pipeline combining the 
 tools of multiple studies would quickly become impractical or impossible 
@@
 
-Though Docker is not immune to these challenges, it offers a interesting 
+Though Docker is not immune to these challenges, it offers an interesting 
 example of a way forward in addressing these fundamental concerns. Here 
@@
 
-First, Docker facilitates modular reuse by build one container on top of
+First, Docker facilitates modular reuse by building one container on top of
 another.  Rather than copy a Dockerfile and then start adding more lines
@@
 the code required to re-generate the results within the manuscript.  As a
-result, it is necessary to package the appropriate typesetting libaries
-(_e.g._ \LaTeX) along with the code libaries such that the document executes
+result, it is necessary to package the appropriate typesetting libraries
+(_e.g._ \LaTeX) along with the code libraries such that the document executes
 successfully for different researchers and platforms.
@@
 fear a it will be primarily a one-way street: more technical savvy
-researchers then themselves can benefit from their hard work, while
+researchers than themselves can benefit from their hard work, while
 they may not benefit from the work produced by others.  Lowering the

### 2014-11-04 docker-paper/manuscript.Rmd
ticking off more edits

@@
 less than 50% of software could even be successfully built or installed
-[@Collberg_2014] and similar results are seen in an ongoing effort by other researchers to
-replicate that study [@Krishnamurthi_2014]. Installing or building software
-necessary to run the code in question assumes the ability to recreate
-the computational environment of the original researchers. 
-<!-- Time savings. Personal reproducibility. Collaboration. Teaching. --> 
-<!-- Figure 1: technical reproducibility stack
-1. Operating system
-2. Workflow systems 
-3. Package managers 
-4. Dynamic documents
--->
-<!-- Omit this? -->
+[@Collberg_2014]. Though sufficiently knowledgeable users may be able
+to overcome the issues in at least some of these cases [^6], similar
+results are seen in an ongoing effort by other researchers to replicate
+that study [@Krishnamurthi_2014], and has also been observed in other
+indepedent studies [@Lapp_2014; @Garijo_2013].  Installing or building
+software necessary to run the code in question assumes the ability to
+recreate the computational environment of the original researchers.
+
+[^6]: See discussion at https://gist.github.com/pbailis/9629050 and https://gist.github.com/samth/9641364
 
@@
 virtualization of the OS, and a hash-based or git-like versioning
-and differencing system, among others.
+and differencing system, among others (see [docs.docker.com/faq](https://docs.docker.com/faq/#what-does-docker-add-to-just-plain-lxc) for
+an excellent overview of what Docker adds to plain LXC).
 
@@
 
-<!-- FIXME move this later --> 
-To simplify this process, Docker also supports *Automated Builds*
-through the Docker Hub ([hub.docker.com](http://hub.docker.com)).  This acts as a kind of Continuous Integration
-(CI) service that verifies the image builds correctly whenever the
-Dockerfile is updated, particularly if the Dockerfile includes checks
-for the environment. The Hub also provides a convenient distribution
-service, freely storing the pre-built images, along with their metadata,
-for download and reuse by others. The Docker Hub is a free service and
-an open source software product so that users can run their own private
-versions of the Hub on their own servers, for instance, if security of
-the data or the longevity of the public platform is a concern.
-
 ### 4. Barriers to adoption and re-use
@@
 
+<!-- Part of the reason is better image distribution-->
 While proponents of virtual machines for reproducible research propose
@@
 certain collaborative tasks or when the work is mature enough to need
-increased computational power.  Working locally allows a researcher to
+increased computational power.
+
 rely more on the graphic interface tools for tasks such as managing
@@
 
+<!--
+
 ### Using RStudio Server
@@
 [https://github.com/ropensci/Docker](https://github.com/ropensci/docker)
+
+-->
 
@@
 
-Sharing these images is further facilitated by the Docker Hub technology. 
-While Docker images tend to be much smaller than equivalent virtual
-machines, moving around even 100's of gigabytes can be a challenge.
-In their work on virtual machines, Dudley _et al._ [@Dudley_2010] recommend only running
-these tools on cloud servers such as the Amazon EC2 system. As most
-researchers still develop software locally and may not have ready access
-to these resources, such a requirement adds an additional barrier to
-reuse.  Docker provides a convenient way to share any image publicly
-or privately through the Hub after creating a free account.  Docker Hub
-is both a publicly available service and also available as 
-a separate open source platform that can be deployed on private servers.
+Sharing these images is further facilitated by the Docker Hub technology
+([hub.docker.com](http://hub.docker.com)).  While Docker images tend
+to be much smaller than equivalent virtual machines, moving around
+even 100's of gigabytes can be a challenge.  The Docker Hub provides a
+convenient distribution service, freely storing the pre-built images,
+along with their metadata, for download and reuse by others. The Docker
+Hub is a free service and an open source software product so that users
+can run their own private versions of the Hub on their own servers, for
+instance, if security of the data or the longevity of the public platform
+is a concern.  Docker also supports *Automated Builds* through the Docker
+Hub.  This acts as a kind of Continuous Integration (CI) service that
+verifies the image builds correctly whenever the Dockerfile is updated,
+particularly if the Dockerfile includes checks for the environment.
 
@@
 
-The approach of Linux Containers represented by Docker offers a technical
+The approach of Docker offers a technical
 solution to what is frequently seen as the primary weakness of the
@@
 ```
-FROM ubuntu:latest
+FROM ubuntu:14.04
 RUN apt-get update 
@@
 This acts like a software dependency; but unlike other software, a
-Dockerfile must have exactly one dependency (one `FROM` line). Note that
-a particular version of the dependency can be specified using the `:` notation.
-One can in fact be much more precise, declaring not only version numbers
-like `ubuntu:14:04`, but specific cryptographic hashes that ensure we
-get the exact same image every time.
-
-One can now build this Dockerfile and give it name, by running in the
+Dockerfile must have exactly one dependency (one `FROM` line). Note
+that a particular version of the dependency can be specified using the
+`:` notation (similarly, package managers like `apt` show here can be
+directed to install specific versions, if desired. One can in fact be much
+more precise by specific cryptographic hash of a particular binary image.
+In practice, using hashes in FROM directives may be undesirable as it also
+avoids adopting security updates that may have been made to the base image.
+
+One can name and build this Dockerfile by running in the
 working directory:
@@
 ```bash
-Docker build -t username/r-recommended .
+docker build -t username/r-recommended .
 ```
@@
 another user can now build directly on our image, rather than on the
-`ubuntu:latest` image used as a base in the first Dockerfile.  To do
+`ubuntu:14.04` image used as a base in the first Dockerfile.  To do
 so, one simply specifies a different image in the `FROM` directive of
@@
 ```
+This separates the task fo running the database (first line) from
+the application used to analyze the data (second line), allowing a more
+modular approach to reuse: another researcher could use the same database
+container while connecting it to different scripts.
 
@@
 batteries included, with everything they need to run reproducibly.
-A researcher could connect a container providing a computational 
-environment for a different language to this same PostgreSQL container, 
-and so forth.
 
@@
 container. For instance, here I inspect recent changes made to the
-`ubuntu:latest` image:
+`ubuntu:14.04` image:
 
 ```bash
-docker history ubuntu:latest
+docker history ubuntu:14.04
 ```
@@
 ```bash
-docker tag 25f ubuntu:latest
+docker tag 25f ubuntu:14.04
 ```
@@
 ```bash
-docker history ubuntu:latest
+docker history ubuntu:14.04
 ```
@@
 versioning system works rather differently than version management
-of code.  Docker can roll back AUFS layers that have been added to
+of code.  Docker can roll back layers[^5] that have been added to
 an image, but not revert to the earlier state of a particular layer.
@@
 to recover an earlier state.
+
+[^5]: Technically AUFS (advanced multi layered unification filesystem )
+layers, see [wikipedia.org/wiki/aufs](http://en.wikipedia.org/wiki/aufs)
 

### 2014-11-05 docker-paper/manuscript.Rmd
texlive example

@@
 
-I introduce the most relevant concepts from Docker through the 
-context of the four challenges for reproducible research I have
-discussed above.
+I introduce the most relevant concepts from Docker through the context
+of the four challenges for reproducible research I have discussed above.
+In brief, these challenges can be addressed by distributing a Dockerfile
+capable of reconstructing the researcher's development environment,
+ideally along with depositing the corresponding binary Docker image in
+an appropriate repository.
 
@@
 image must be based on a Linux system with Linux-compatible software, which
-includes (R, Python, Matlab, and most other scientific programming needs). [^4]
-
-Sharing the Linux kernel makes Docker much more light-weight and higher
+includes R, Python, Matlab, and most other scientific programming needs.[^4]
+
+Sharing the Linux kernel makes Docker more light-weight and higher
 performing than complete virtual machines -- a typical desktop computer
@@
 Docker provides a simple way to confirm whether or not code rot has
-effected the function of a particular piece of code.
+effected the function of a particular piece of code. Binary Docker images
+can be efficiently shared through the Docker Hub as well, as I describe
+later.
 
@@
 example of a way forward in addressing these fundamental concerns. Here 
-I highlight five of these features in turn:
-
-- Local environment
+I highlight these features in turn:
+
+- Integrating into local development environments
 - Modular reuse
@@
 
-## Using Docker as a local development environment
-
-Perhaps the most important feature of a reproducible research tool is that
-it be easy to learn and fit relatively seamlessly into existing workflow
-patterns of domain researchers.  This, more than any other concern,
-can explain the relatively low uptake of previously proposed solutions.
-Being a new and unfamiliar tool to most domain scientists, Docker is
-far from immune to the same critique.  Nevertheless, Docker takes us
-several key steps towards an approach that can be easily adopted in a
-research context.
-
-<!-- Part of the reason is better image distribution-->
-While proponents of virtual machines for reproducible research propose
+## Integrating into local development environments
+
+Due in part to the difficulty of moving large VM images around a network,
+proponents of virtual machines for reproducible research often propose
 that these machines would be available exclusively as cloud computing
-environments [_e.g._ @Dudley_2010], many researchers work _locally_, that
-is, primarily with software that is installed on their laptop or desktop
-computer, and turning to cloud-based or other remote platforms only for
-certain collaborative tasks or when the work is mature enough to need
-increased computational power.
-
-rely more on the graphic interface tools for tasks such as managing
-files, text editing, debugging, IDEs, or interacting with version
-control systems.  Scientific computing on remote machines, by contrast,
-still relies largely on potentially less familiar text based command
-line functions for these tasks (though web based interfaces like RStudio
-Server are rapidly filling this gap).
-
-Docker can be easily installed on most major platforms (see
-[https://docs.Docker.com/installation](https://docs.docker.com/installation);
-On systems not already based on the Linux Kernel, such as Mac or Windows,
-this is accomplished through the use of a small VirtualBox-based VM
-running on the host OS called `boot2docker`) and run locally.  Docker
-allows a user to link any directory (such as the working directory
-for a particular script or project) to the running Docker container.
-This allows a user to rely on the familiar tools of the host OS for
-roles such as text editing, file browsing, or version control, while
-still allowing code execution to occur inside the controlled development
-environment of the container.
-
-For example, one launches an interactive R console in a container that
-is linked to our current working directory like so:
-
-```bash
-docker run -v $(pwd):/ -it cboettig/rstudio /usr/bin/R
-```
-
-The resulting system behaves almost identically[^2] to running R on
-the command line of the host OS. (Clearly a similar command could be
-used just as well with interactive shells from other languages such as
-`ipython` or `irb`).
-
-[^2]: External windows, such graphics windows cannot be opened directly
-from the container in this setup.  Graphics would have to be saved to
-disk as raster or vector files and viewed on the host OS, until a better
-solution can be found. RStudio Server ([rstudio.com/products/rstudio/#Server](http://rstudio.com/products/rstudio/#Server)) is one way around this.
-
-<!--
-
-### Using RStudio Server
-
-An alternative approach for working locally with familiar tools is to
-leverage web-based clients such as RStudio. 
-In the R environment, the open source RStudio IDE provides
-another key component in making this system accessible to most
-domain scientists. Because the RStudio IDE is written completely with
-standard web languages (Javascript, CSS, and HTML), its web-based 
-RStudio Server looks and feels identical to its popular desktop IDE.
-RStudio Server provides a way to interact with R on a remote environment
-without the latency, X tunnelling, and so forth typically involved in running R
-on a remote server.
-
-RStudio server provides users a way to interact with R,
-the file system, git, text editors, and graphics running on a Docker
-container.  Users already familiar with the popular IDE can thus benefit
-from the reproducibility and portability features provided by running R
-in a container environment without having to adapt to a new workflow.
-
-To accompany this paper, I provide a Docker image for running RStudio
-server, which can be launched as follows.  From the terminal (or
-boot2docker terminal on Mac or Windows client), run:
-
-```bash
-sudo docker run -d -p 8787:8787 cboettig/ropensci
-```
-
-That will take a while to download the image the first time you run it. A
-boot2docker user will then need to determine the ip address assigned
-to boot2docker as shown below, while a Linux user can just specify
-`http://localhost`.
-
-```bash
-boot2docker ip
-```
-
-Add the port `:8787` to the end of this address and paste
-it into your browser address bar, which should open to the
-RStudio welcome screen.  A user can now login with user/password
-`rstudio/rstudio`, run R scripts, install packages, use git,
-and so forth. User login and other configurations can be
-customized using environmental variables; see details at
-[https://github.com/ropensci/Docker](https://github.com/ropensci/docker)
-
--->
+environments [_e.g._ @Dudley_2010] rather than downloaded to a user's
+personal computer. Though cloud computing offers some advantages such as
+scalable resources for computationally intensive tasks, many researchers
+prefer to work with tools installed locally on their personal computer,
+at least during testing and development.  This can reduce costs, issues of
+network latency, the ability to work off-line, and will be most familiar
+to students and new developers.
+
+It is possible to run virtual machines locally on most common laptop and
+desktop computers, as demonstrated in the approach now being pioneered
+at UC Berkeley [@Clark_2014].  This approach provides users with a pixel-identical
+environment whether working locally or on the cloud, which has particular
+advantages for student instruction [@Clark_2014]. However, it remains to be
+seen if existing researchers will be willing to forgo native applications
+for such tasks as file browsing, text editing, or version management and 
+rely exclusively on this standardized virtual environment.
+
+In contrast, Docker's approach is optimized for a more integrated
+workflow.  Rather than replace a user's existing toolchain with a
+standardized virtual environment, editors and all, Docker is optimized at
+the level of single applications. A developer can thus rely on familiar
+tools while still ensuring that the execution of their code always occurs
+on the standardized container environment, thus ensuring its portability
+and reproducibility. This approach can be accomplished either by linking
+volumes or directories between the container and the host, or simply
+by instructing the Dockerfile to copy the code to the container (As we will
+see, modular reuse makes this latter strategy efficient).
+
+Rather than put the burden on the researcher to adopt
+a very different workflow, the researcher can thus use their familiar editors,
+etc, while immediately benefiting from the fact their computations can
+be easily deployed on the cloud or the machines of collaborators without
+any further effort than installing Docker software. The docker approach
+is particularly well suited for moving between local and cloud platforms
+when a web-based integrated development environment is available, such as
+RStudio Server, or (to lesser extent) an iPython notebook.
+
+On systems not already based on the Linux kernel (such
+as Mac or Windows platforms), Docker is installed (see see
+[docs.docker.com/installation](https://docs.docker.com/installation))
+through means of a very small (about
+24 MB) VM platform called `boot2docker`
+([github.com/docker/boot2docker](https://github.com/docker/boot2docker)),.
+While this poses an additional challenge for tight integration with desktop
+tools, recent advances in Docker are rapidly bridging this gap as well. 
+For example, Docker 1.3, released between drafts of this manuscript, supports
+shared volumes on Macs ([blog.docker.com](http://blog.docker.com/2014/10/docker-1-3-signed-images-process-injection-security-options-mac-shared-directories/)).
 
@@
 solution to what is frequently seen as the primary weakness of the
-standard virtual machine approach to reproducibility - reusing and
+standard VM approach to reproducibility - reusing and
 remixing elements.  To some extent this is already addressed by the
@@
 
-The challenge to reusing virtual machines can be summarized as "you
-can't install an image for every pipeline you want..." [@Watson_2012]. In
-contrast, this is exactly how Docker containers are designed to work.
-There are at least two ways in which Docker supports this kind of
+The challenge to reusing VMs can be summarized as "you
+can't install an image for every pipeline you want..." [@Watson_2012]. 
+While providing a VM may make it easy for other researchers to run
+a particular piece of software, it becomes very difficult to combine
+multiple software components in future research if each must run inside
+it's own VM. The lack of reusable, scalable modules in the VM model
+poses a major barrier to future reuse.  In contrast, (as the analogy to 
+shipping containers in the name might imply) Docker containers
+are optimized for this kind of stacking and modular reuse.
+There are at least three ways in which Docker supports this kind of
 extensibility.
 
-First, Docker facilitates modular reuse by building one container on top of
-another.  Rather than copy a Dockerfile and then start adding more lines
-to the bottom, one can declare a new Dockerfile is built on an old one
-using the `FROM` directive. Here is an example Dockerfile that adds the
-R statistical computing environment to a basic Ubuntu Linux distribution.
+Most primitively, because Dockerfile itself provides a script for
+creating the computational environment, future researchers can
+extend or modify the resulting machine image by simply editing the
+Dockerfile. This same approach is also possible for VM approaches that rely on 
+DevOps tools [@Clark_2014]. In contrast to VMs, however, Docker is also
+modular by design: both in how Dockerfiles are defined and how Docker
+containers can be linked.
+
+First, Docker facilitates modular reuse by building one container on top
+of another through the use of `FROM` directive in Dockerfiles. This acts
+like a software dependency; but unlike other software, a Dockerfile must
+have exactly one dependency (one `FROM` line). Particular version of
+the dependency can be specified using the `:` notation, or ommitted to
+default to the latest version. To install software, Dockerfiles leverage
+existing package managers on common Linux distributions such as `apt`
+on Debian. These also permit installing either specific or only the
+latest version.
+
+Consider the following Dockerfile, which
 
 ```
-FROM ubuntu:14.04
-RUN apt-get update 
-RUN apt-get -y install r-recommended
+FROM rocker/ropensci:latest
+RUN git clone https://github.com/ropensci/RNeXML.git /rnexml
+RUN Rscript -e 'devtools::install("/rnexml"); devtools::check("/rnexml")'
 ```
 
-This acts like a software dependency; but unlike other software, a
-Dockerfile must have exactly one dependency (one `FROM` line). Note
-that a particular version of the dependency can be specified using the
-`:` notation (similarly, package managers like `apt` show here can be
-directed to install specific versions, if desired. One can in fact be much
-more precise by specific cryptographic hash of a particular binary image.
-In practice, using hashes in FROM directives may be undesirable as it also
-avoids adopting security updates that may have been made to the base image.
-
-One can name and build this Dockerfile by running in the
-working directory:
-
-```bash
-docker build -t username/r-recommended .
-```
-
-The key point here is that other researchers can easily build off this
-image just created, extending our work directly, rather than having
-to go back to the original image.
-
-If this image is shared (either directly or through a Docker Hub),
-another user can now build directly on our image, rather than on the
-`ubuntu:14.04` image used as a base in the first Dockerfile.  To do
-so, one simply specifies a different image in the `FROM` directive of
-the Dockerfile, followed by the lines required to add any additional
-software required.
-
-```
-FROM username/r-recommended 
-RUN apt-get update
-RUN apt-get -y install r-cran-matrix
-```
-
-Though this shows only very minimal examples that add a single piece of
-software in each step, clearly this approach can be particularly powerful
-in building up more complex environments.
+A Dockerfile can have only a single `FROM` line, while sometimes it may
+be necessary to build on the computational environment provided by more
+than one container. To address this, Docker defines a syntax for linking
+multiple containers together:
 
@@
 run one particular service or element, and exposing just what is necessary
-to link it together with other blocks.  For instance, one could have one
+to link it together with other blocks.
+
+```bash
+docker run --name tex -v /usr/local/texlive rocker/texlive
+docker run -dP --volumes-from tex \
+-e PATH=$PATH:/usr/local/texlive/2014/bin/x86_64-linux/ \
+rocker/rstudio 
+```
+
+For instance, one could have one
 container running a PostgreSQL database which serves data to another container
@@
 
-Unlike the much more heavyweight virtual machine approach, containers
-are implemented in way such that a single computer can easily run 100s
-of such services each in their own container, making it easy to break
-computational elements down into logically reusable chunks that come,
-batteries included, with everything they need to run reproducibly.
+Unlike the much more heavyweight virtual machine approach, a single
+computer can easily run 100s of such services each in their own
+container. A rapidly expanding ecosystem of software around Docker,
+such as fig ([github.com/docker/fig](https://github.com/docker/fig))
+facilitates the complexity of running multiple containers.  This feature
+making it easy to break computational elements down into logically
+reusable chunks that come, batteries included, with everything they need
+to run reproducibly.
 

### 2014-11-06 docker-paper/manuscript.Rmd
further edits

@@
 
-Consider the following Dockerfile, which
+<!--
+Consider the following Dockerfile to illustrate this modular reuse. This
+is the complete Dockerfile required to create the computational environment
+required for a particular manuscript:
 
-```
+```bash
 FROM rocker/ropensci:latest
+MAINTAINER Carl Boettiger cboettig@ropensci.org
 RUN git clone https://github.com/ropensci/RNeXML.git /rnexml
-RUN Rscript -e 'devtools::install("/rnexml"); devtools::check("/rnexml")'
+RUN Rscript -e 'devtools::install("/rnexml")'
 ```
 
-A Dockerfile can have only a single `FROM` line, while sometimes it may
+The first line contains the FROM directive, indicating that this Dockerfile
+builds on the existing Docker image `rocker/ropensci`, hosted on the Docker
+Hub.  For this approach to be effective for reproducible one must take care 
+that the base image and it's Dockerfile recipe are available from a persistant
+archive.
+
+The second line is optional, declaring minimal metadata regarding the 
+contact information for the image maintainer. The third line copys the
+project onto the image and the fourth installs the necessary software that
+has not been provided by the base image. Projects using less standard
+environments may require more instructions.  
+-->
+
+While a Dockerfile can have only a single `FROM` line, sometimes it may
 be necessary to build on the computational environment provided by more
 than one container. To address this, Docker defines a syntax for linking
-multiple containers together:
+multiple containers together.  This allows each container to act as a
+building block providing just what is necessary to run one particular
+service or element, and exposing just what is needed to link it together
+with other blocks.  For instance, one could have one container running
+a PostgreSQL database which serves data to another container running a
+python environment to analyze the data:
 
-Each one acts as a building block providing just what is necessary to
-run one particular service or element, and exposing just what is necessary
-to link it together with other blocks.
-
+<!--
 ```bash
-docker run --name tex -v /usr/local/texlive rocker/texlive
-docker run -dP --volumes-from tex \
+docker run --name texlive -v /usr/local/texlive rocker/texlive
+docker run -dP --volumes-from texlive \
 -e PATH=$PATH:/usr/local/texlive/2014/bin/x86_64-linux/ \
@@
 ```
-
-For instance, one could have one
-container running a PostgreSQL database which serves data to another container
-running a python environment to analyze the data:
+-->
 

### 2014-11-11 docker-paper/manuscript.Rmd
docker for travis, maybe?

@@
 
-Docker is an open source project that builds on many long-familiar
-technologies from operating systems research: LXC containers,
-virtualization of the OS, and a hash-based or git-like versioning
-and differencing system, among others (see [docs.docker.com/faq](https://docs.docker.com/faq/#what-does-docker-add-to-just-plain-lxc) for
-an excellent overview of what Docker adds to plain LXC).
+Docker is an open source project that builds on many
+long-familiar technologies from operating systems research:
+LXC containers, virtualization of the OS, and a hash-based
+or git-like versioning and differencing system, among others (see
+[docs.docker.com/faq](https://docs.docker.com/faq/#what-does-docker-add-to-just-plain-lxc)
+for an excellent overview of what Docker adds to plain LXC). The
+official documentation [docs.docker.com](https://docs.docker.com)
+already provides a thorough introduction in how to use Docker
+software; here my focus is on describing the implications this
+has for reproducible research.  Readers may also find it helpful
+to see more detailed examples of using Docker to capture, share,
+and interact with a specific computational environments. Some such
+examples, along with more detailed documentation on use can be found at
+[github.com/rocker-org](https://github.com/rocker-org/rocker).
 
@@
 a very different workflow, the researcher can thus use their familiar editors,
-etc, while immediately benefiting from the fact their computations can
+etc., while immediately benefiting from the fact their computations can
 be easily deployed on the cloud or the machines of collaborators without
@@
 have exactly one dependency (one `FROM` line). Particular version of
-the dependency can be specified using the `:` notation, or ommitted to
+the dependency can be specified using the `:` notation, or omitted to
 default to the latest version. To install software, Dockerfiles leverage
@@
 ```
-This separates the task fo running the database (first line) from
+This separates the task for running the database (first line) from
 the application used to analyze the data (second line), allowing a more

### 2014-12-01 docker-paper/manuscript.Rmd
commit final revisions on formatting

@@
 csl: components/acm-sig-proceedings.csl
+citecolor: black
+linkcolor: black
+urlcolor: black
 
@@
 One can  share a public copy of the image just created by using the
-`Docker push` command, followed by the name of the image using the command:
+`docker push` command, followed by the name of the image using the command:
 
@@
 the Dockerfile, making the `push` command unnecessary. A user can update
-their local image using the `Docker pull <imagename>`, which downloads
+their local image using the `docker pull <imagename>`, which downloads
 any changes that have since been made to the copy of the image on the Hub.
@@
 docker run -d --name db training/postgres
-docker run -d -P --link db:db training/webapp python app.py
+docker run -d -P --link db:db training/webapp \
+  python app.py
 ```
@@
 In addition to version managing the Dockerfile, the images themselves are
-versioned using a git-like hash system (_e.g._ see `Docker commit`, `docker
-push`/`Docker pull`, `docker history`, `docker diff`).  Docker images and
+versioned using a git-like hash system (_e.g._ see `docker commit`, `docker
+push`/`docker pull`, `docker history`, `docker diff`).  Docker images and
 containers have dedicated metadata specifying the date, author, parent
-image, and other details (see `Docker inspect`). One can roll back an
+image, and other details (see `docker inspect`). One can roll back an
 image through the layers of history of its construction, then build off

### 2014-12-01 setup.sh

@@ -152,8 +152,12 @@ docker run --name='rstudio' -d -p $RSTUDIO:8787 -e USER=$USER -e PASSWORD=$PASSW
 git clone https://github.com/drone/drone.git
 docker build -t drone/drone drone/
 ## Deploy Drone CI on 8080 (Then visit localhost:8080/install)
-docker run --name='drone' -d -p $DRONE:80 --privileged drone/drone
-
+docker run --name drone -d -p 88:80 \
+-v /var/run/docker.sock:/var/run/docker.sock \
+-t \
+-e DRONE_GITHUB_CLIENT=$DRONE_GITHUB_CLIENT \
+-e DRONE_GITHUB_SECRET=$DRONE_GITHUB_SECRET \
+drone/drone
 
 ## Deploy Gitlab: we need some data containers running first
 docker run --name=postgresql -d \

### 2014-12-04 docker-paper/manuscript.tex
archiving submitted copies

--- 2014-11-11 docker-paper/manuscript.Rmd
@@
----
-title: "An introduction to Docker for reproducible research"
-author: 
-  - name: Carl Boettiger
-    affiliation:
-      l1: Center for Stock Assessment Research, 
-      l2: 110 Shaffer Rd, Santa Cruz, CA 95050, USA 
-    email: cboettig(at)gmail.com
-    footnote: Corresponding author
-abstract: |
+% THIS IS SIGPROC-SP.TEX - VERSION 3.1
+% WORKS WITH V3.2SP OF ACM_PROC_ARTICLE-SP.CLS
+% APRIL 2009
+%
+% It is an example file showing how to use the 'acm_proc_article-sp.cls' V3.2SP
+% LaTeX2e document class file for Conference Proceedings submissions.
+% ----------------------------------------------------------------------------------------------------------------
+% This .tex file (and associated .cls V3.2SP) *DOES NOT* produce:
+%       1) The Permission Statement
+%       2) The Conference (location) Info information
+%       3) The Copyright Line with ACM data
+%       4) Page numbering
+% ---------------------------------------------------------------------------------------------------------------
+% It is an example which *does* use the .bib file (from which the .bbl file
+% is produced).
+% REMEMBER HOWEVER: After having produced the .bbl file,
+% and prior to final submission,
+% you need to 'insert'  your .bbl file into your source .tex file so as to provide
+% ONE 'self-contained' source file.
+%
+% Questions regarding SIGS should be sent to
+% Adrienne Griscti ---> griscti@acm.org
+%
+% Questions/suggestions regarding the guidelines, .tex and .cls files, etc. to
+% Gerald Murray ---> murray@hq.acm.org
+%
+% For tracking purposes - this is V3.1SP - APRIL 2009
+
+\documentclass[9pt]{components/acm_proc_article-sp}
+%\documentclass{acm_proc_article-sp}
+\usepackage[hyphens]{url}
+\usepackage{graphicx}
+\usepackage{booktabs} % book-quality tables
+
+\usepackage[T1]{fontenc}
+\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
+% use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[utf8]{inputenc}
+
+\else % if luatex or xelatex
+  \usepackage{fontspec}
+  \ifxetex
+    \usepackage{xltxtra,xunicode}
+  \fi
+  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
+  \newcommand{\euro}{€}
+\fi
+
+  % use microtype if available
+\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
+\usepackage{color}
+\usepackage{fancyvrb}
+\newcommand{\VerbBar}{|}
+\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
+\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
+% Add ',fontsize=\small' for more characters per line
+\usepackage{framed}
+\definecolor{shadecolor}{RGB}{248,248,248}
+\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
+\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
+\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{{#1}}}
+\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
+\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
+\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
+\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
+\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
+\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
+\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{{#1}}}
+\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{{#1}}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
+\newcommand{\RegionMarkerTok}[1]{{#1}}
+\newcommand{\ErrorTok}[1]{\textbf{{#1}}}
+\newcommand{\NormalTok}[1]{{#1}}
+\ifxetex
+  \usepackage[setpagesize=false, % page size defined by xetex
+              unicode=false, % unicode breaks when used with xetex
+              xetex]{hyperref}
+\else
+  \usepackage[unicode=true]{hyperref}
+\fi
+\hypersetup{breaklinks=true,
+            bookmarks=true,
+            pdfauthor={},
+            pdftitle={An introduction to Docker for reproducible research},
+            colorlinks=true,
+            urlcolor=black,
+            linkcolor=black,
+            citecolor=black,
+            pdfborder={0 0 0}}
+\urlstyle{same}  % don't use monospace font for urls
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+\setlength{\emergencystretch}{3em}  % prevent overfull lines
+\setcounter{secnumdepth}{0}
+% Pandoc toggle for numbering sections (defaults to be off)
+\setcounter{secnumdepth}{0}
+% Pandoc header
+
+ %% Omit the copyright space.
+%  \makeatletter
+%  \def\@copyrightspace{}
+%  \makeatother
+
+  %% Fill in the title and author list for your paper.
+%  \title{An introduction to Docker for reproducible research}
+%  \author{ \thanks{Copyright is held by the author(s)}}
+%  \renewcommand\footnotemark{}
+
+\toappear{Copyright is held by the author(s)}
+\begin{document}
+
+\title{An introduction to Docker for reproducible research}
+\subtitle{}
+%
+% You need the command \numberofauthors to handle the 'placement
+% and alignment' of the authors beneath the title.
+%
+% For aesthetic reasons, we recommend 'three authors at a time'
+% i.e. three 'name/affiliation blocks' be placed beneath the title.
+%
+% NOTE: You are NOT restricted in how many 'rows' of
+% "name/affiliations" may appear. We just ask that you restrict
+% the number of 'columns' to three.
+%
+% Because of the available 'opening page real-estate'
+% we ask you to refrain from putting more than six authors
+% (two rows with three columns) beneath the article title.
+% More than six makes the first-page appear very cluttered indeed.
+%
+% Use the \alignauthor commands to handle the names
+% and affiliations for an 'aesthetic maximum' of six authors.
+% Add names, affiliations, addresses for
+% the seventh etc. author(s) as the argument for the
+% \additionalauthors command.
+% These 'additional authors' will be output/set for you
+% without further effort on your part as the last section in
+% the body of your article BEFORE References or any Appendices.
+
+\numberofauthors{1} %  in this sample file, there are a *total*
+% of EIGHT authors. SIX appear on the 'first-page' (for formatting
+% reasons) and the remaining two appear in the \additionalauthors section.
+%
+\author{
+% You can go ahead and credit any number of authors here,
+% e.g. one 'row of three' or two rows (consisting of one row of three
+% and a second row of one, two or three).
+%
+% The command \alignauthor (no curly braces needed) should
+% precede each author name, affiliation/snail-mail address and
+% e-mail address. Additionally, tag each line of
+% affiliation/address with \affaddr, and tag the
+% e-mail address with \email.
+%
+% 1st. author
+  \alignauthor
+  Carl Boettiger %  \titlenote{Corresponding author}
+  \\
+  \affaddr{\small Center for Stock Assessment Research,}\\
+  \affaddr{\small 110 Shaffer Rd, Santa Cruz, CA 95050, USA}\\
+  \email{\small cboettig(at)gmail.com}
+}
+\date{\today}
+
+\maketitle
+\begin{abstract}
 As computational work becomes more and more integral to many aspects of
@@
 scientists alike. Though computational reproducibility seems more
-  straight forward than replicating physical experiments, the complex
-  and rapidly changing nature of computer environments makes being able
-  to reproduce and extend such work a serious challenge.  In this paper,
-  I explore common reasons that code developed for one research project
-  cannot be successfully executed or extended by subsequent researchers.
-  I review current approaches to these issues, including virtual machines
+straight forward than replicating physical experiments, the complex and
+rapidly changing nature of computer environments makes being able to
+reproduce and extend such work a serious challenge. In this paper, I
+explore common reasons that code developed for one research project
+cannot be successfully executed or extended by subsequent researchers. I
+review current approaches to these issues, including virtual machines
 and workflow systems, and their limitations. I then examine how the
@@
 research - such as operating system virtualization, cross-platform
-  portability, modular re-usable elements, versioning, and a 'DevOps'
+portability, modular re-usable elements, versioning, and a `DevOps'
 philosophy, to address these challenges. I illustrate this with several
 examples of Docker use with a focus on the R statistical environment.
-
-Keywords:
-  - Docker
-  - virtualization
-  - reproducibility
-
-bibliography: components/references.bib
-csl: components/acm-sig-proceedings.csl
-
-output: 
-  pdf_document:
-    template: components/sigproc-sp.tex
-    keep_tex: true
-    fig_caption: true
----
-
-<!-- Excerpt from the CALL FOR PAPERS:
-There is increasing interest in promoting repeatable and reproducible research
-in computer science.  A community gains confidence in scientific claims when
-the experiments supporting those claims are published, examined, and repeated.
-Repeating research also helps a community expand on previous results and
-discover new directions for investigation.  Repetition requires that the
-artifacts used in an experiment be made available to, and made usable by,
-investigators.  In other words, artifacts must be shared.  Possible artifacts
-include not only binaries, source code, and datasets but also computing and
-networking infrastructure such as testbeds and clouds.
-
-Systems researchers face special challenges in the areas of repeatability and
-artifact sharing.  Some arise from the nature of systems software, which is
-typically complex and rife with internal and external dependencies.  Other
-issues arise from execution environments: for example, to reproduce performance
-and scalability results, one may need access to actual and specific hardware.
-Still more systems-specific challenges arise from the difficulty of measuring
-properties of interest, such as time and energy consumption.
-
-Suggested paper topics include, but are not limited to:
-
-  * novel ways to manage systems research artifacts, dependencies, workflows,
-    and data collection
-  * testbeds for repeatable research and/or publishing executable artifacts
-  * "rich publications" that connect papers to software and data
-  * repeatability in distributed and nondeterministic systems
-  * mitigating obstacles to artifact sharing, repeatability, and reproduction
-  * repeatability challenges specific to systems
-  * measurement bias in systems experiments
-  * experiences and outcomes from systems repetition and reproduction studies
-  * improving scientific practice in systems research
-  * incentives for producing and sharing high-quality research artifacts
--->
-
-# Introduction
+\end{abstract}
+
+% A category with the (minimum) three required fields
+%\category{H.4}{Information Systems Applications}{Miscellaneous}
+%A category including the fourth, optional field follows...
+%\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]
+
+%\terms{Systems, Reproducible Research}
+%
+
+\section{Introduction}\label{introduction}
 
 Reproducible research has received an increasing level of attention
-throughout the scientific community [_e.g._ @Peng_2011; @Nature_2012]
-and the public at large [_e.g._ @The_Economist_2013]. All steps of the
-scientific process, from data collection and processing, to analyses,
-visualizations and conclusions depend ever more on computation and
-algorithms, _computational reproducibility_ has received particular
-attention [@Merali_2010].  Though in principle this algorithmic dependence
-should make such research easier to reproduce -- computer codes being
-both more portable and potentially more precise to exchange and run
-than experimental methods -- in practice this has led to an ever larger
-and more complex black box that stands between what was actually done
-and what is described in the literature.  Crucial scientific processes
-such as replicating the results, extending the approach or testing the
-conclusions in other contexts, or even merely installing the software
-used by the original researchers can become immensely time-consuming if
-not impossible.
-
-### Systems research & reproducibility
+throughout the scientific community {[}19, 22{]} and the public at large
+{[}25{]}. All steps of the scientific process, from data collection and
+processing, to analyses, visualizations and conclusions depend ever more
+on computation and algorithms, \emph{computational reproducibility} has
+received particular attention {[}18{]}. Though in principle this
+algorithmic dependence should make such research easier to reproduce --
+computer codes being both more portable and potentially more precise to
+exchange and run than experimental methods -- in practice this has led
+to an ever larger and more complex black box that stands between what
+was actually done and what is described in the literature. Crucial
+scientific processes such as replicating the results, extending the
+approach or testing the conclusions in other contexts, or even merely
+installing the software used by the original researchers can become
+immensely time-consuming if not impossible.
+
+\subsubsection{Systems research \&
+reproducibility}\label{systems-research-reproducibility}
 
@@
 computational reproducibility and the technologies that can facilitate
-those objectives [_e.g._ @Eide_2010, @Harji_2013]. Docker is a new but already
-very popular open source tool that combines many of these approaches
-in a user friendly implementation, including: (1) performing Linux
-container (LXC) based operating system (OS) level virtualization, (2)
-portable deployment of containers across platforms, (3) component reuse,
-(4) sharing, (5) archiving, and (6) versioning of container images. While
-Docker's market success has largely focused on the needs of businesses
-in deploying web applications and the potential for a lightweight
-alternative to full virtualization, these features have potentially
-important implications for systems research in the area of scientific
-reproducibility.
-
-In this paper, I seek to address two audiences. First, that of the domain
-scientist, those conducting research in ecology, bioinformatics, economics,
-psychology and so many other disciplines in which computation plays an 
-ever-increasing role. I seek to help this audience become more aware of the 
-concerns and challenges in making these computations more reproducible, extensible, 
-and portable to other researchers, and highlight Docker as platform to address 
-these challenges. The second audience is that of the computer systems researchers:
-readers familiar with both the challenges to reproducibility (versioning,
-software dependencies, etc.) and the technical components underlying Docker (virtualization,
-LXC containers, jails, hashes, etc.), but who may be unfamiliar with Docker software. 
-I hope the domain scientist will be motivated to try Docker as a tool to address
-reproducibility. Meanwhile, I expect the systems researcher to see Docker technology as
-an area ripe for further systems research (e.g. to what extent can differences between 
-Linux kernels frustrate reproducibility in a Docker image?) rather than as a tool
-for their own study (because Docker applies only to software at the level above
-the Linux kernel, it will be of little use in making studies of differences in 
-kernel behavior, hardware performance, execution times, and so forth reproducible.)
-
-### A cultural problem
+those objectives {[}6{]}. Docker is a new but already very popular open
+source tool that combines many of these approaches in a user friendly
+implementation, including: (1) performing Linux container (LXC) based
+operating system (OS) level virtualization, (2) portable deployment of
+containers across platforms, (3) component reuse, (4) sharing, (5)
+archiving, and (6) versioning of container images. While Docker's market
+success has largely focused on the needs of businesses in deploying web
+applications and the potential for a lightweight alternative to full
+virtualization, these features have potentially important implications
+for systems research in the area of scientific reproducibility.
+
+In this paper, I seek to address two audiences. First, that of the
+domain scientist, those conducting research in ecology, bioinformatics,
+economics, psychology and so many other disciplines in which computation
+plays an ever-increasing role. I seek to help this audience become more
+aware of the concerns and challenges in making these computations more
+reproducible, extensible, and portable to other researchers, and
+highlight Docker as platform to address these challenges. The second
+audience is that of the computer systems researchers: readers familiar
+with both the challenges to reproducibility (versioning, software
+dependencies, etc.) and the technical components underlying Docker
+(virtualization, LXC containers, jails, hashes, etc.), but who may be
+unfamiliar with Docker software. I hope the domain scientist will be
+motivated to try Docker as a tool to address reproducibility. Meanwhile,
+I expect the systems researcher to see Docker technology as an area ripe
+for further systems research (e.g.~to what extent can differences
+between Linux kernels frustrate reproducibility in a Docker image?)
+rather than as a tool for their own study (because Docker applies only
+to software at the level above the Linux kernel, it will be of little
+use in making studies of differences in kernel behavior, hardware
+performance, execution times, and so forth reproducible.)
+
+\subsubsection{A cultural problem}\label{a-cultural-problem}
 
 It is worth observing from the outset that the primary barrier to
-computational reproducibility in many domain sciences has nothing
-to do with the technological approaches discussed here, but stems rather from a
-reluctance to publish the code used in generating the results in the
-first place [@Barnes_2010]. Despite extensive evidence to the contrary
-[_e.g._ @Ince_2012], many researchers and journals continue to assume that
-summary descriptions or pseudo-code provide a sufficient description of
+computational reproducibility in many domain sciences has nothing to do
+with the technological approaches discussed here, but stems rather from
+a reluctance to publish the code used in generating the results in the
+first place {[}2{]}. Despite extensive evidence to the contrary
+{[}14{]}, many researchers and journals continue to assume that summary
+descriptions or pseudo-code provide a sufficient description of
 computational methods used in data gathering, processing, simulation,
 visualization, or analysis. Until such code is available in the first
-place, one cannot even begin to encounter the problems that the approaches
-discussed here set out to solve. As a result, few domain researchers
-may be fully aware of the challenges involved in effectively re-using
-published code.
+place, one cannot even begin to encounter the problems that the
+approaches discussed here set out to solve. As a result, few domain
+researchers may be fully aware of the challenges involved in effectively
+re-using published code.
 
 A lack of requirements or incentives no doubt plays a crucial role in
-discouraging sharing [@Barnes_2010; @Stodden_2013]. Nevertheless,
-it is easy to underestimate the significant barriers raised by a lack
-of familiar, intuitive, and widely adopted tools for addressing the
-challenges of computational reproducibility. Surveys and case studies find
-that a lack of time, more than innate opposition to sharing, discourages
-researchers from providing code [@Stodden_2010; @FitzJohn_2014].
-
-## Four technical challenges
+discouraging sharing {[}2, 24{]}. Nevertheless, it is easy to
+underestimate the significant barriers raised by a lack of familiar,
+intuitive, and widely adopted tools for addressing the challenges of
+computational reproducibility. Surveys and case studies find that a lack
+of time, more than innate opposition to sharing, discourages researchers
+from providing code {[}7, 23{]}.
+
+\subsection{Four technical challenges}\label{four-technical-challenges}
 
 By restricting ourselves to studies of where code has been made
-available, I will sidestep for the moment the cultural challenges
-to reproducibility so that I may focus on the technical ones; in
+available, I will sidestep for the moment the cultural challenges to
+reproducibility so that I may focus on the technical ones; in
 particular, those challenges for which improved tools and techniques
@@
 barriers to reproducing the original results or building on that code
-[_e.g._ @Lapp_2014; @Garijo_2013; @Gilbert_2012; @Collberg_2014;
-@Krishnamurthi_2014], which I attempt to summarize as follows.
-
-### 1. **"Dependency Hell"**
+{[}4, 8, 10, 16, 27{]}, which I attempt to summarize as follows.
+
+\subsubsection{1. \textbf{``Dependency Hell''}}\label{dependency-hell}
 
 A recent study by researchers at the University of Arizona found that
-less than 50% of software could even be successfully built or installed
-[@Collberg_2014]. Though sufficiently knowledgeable users may be able
-to overcome the issues in at least some of these cases [^6], similar
-results are seen in an ongoing effort by other researchers to replicate
-that study [@Krishnamurthi_2014], and has also been observed in other
-indepedent studies [@Lapp_2014; @Garijo_2013].  Installing or building
-software necessary to run the code in question assumes the ability to
-recreate the computational environment of the original researchers.
-
-[^6]: See discussion at https://gist.github.com/pbailis/9629050 and https://gist.github.com/samth/9641364
+less than 50\% of software could even be successfully built or installed
+{[}4{]}. Though sufficiently knowledgeable users may be able to overcome
+the issues in at least some of these cases,\footnote{See discussion at
+  \url{https://gist.github.com/pbailis/9629050} and
+  \url{https://gist.github.com/samth/9641364}} similar results are seen
+in an ongoing effort by other researchers to replicate that study
+{[}27{]}, and has also been observed in other indepedent studies {[}8,
+16{]}. Installing or building software necessary to run the code in
+question assumes the ability to recreate the computational environment
+of the original researchers.
 
 Differences in numerical evaluation, such as arise in floating point
-arithmetic or even ambiguities in standardized programming
-languages ("order-of-evaluation" problems) can be responsible for
-differing results between or even within the same computational platform
-[@Ince_2012]. Such issues make it difficult to restrict the true dependencies
-of the code to higher level environments such as that of a given scripting
+arithmetic or even ambiguities in standardized programming languages
+(``order-of-evaluation'' problems) can be responsible for differing
+results between or even within the same computational platform {[}14{]}.
+Such issues make it difficult to restrict the true dependencies of the
+code to higher level environments such as that of a given scripting
 language, independent of the underlying OS or even hardware itself.
 
-### 2.  **Imprecise documentation**
+\subsubsection{2. \textbf{Imprecise
+documentation}}\label{imprecise-documentation}
 
 Documentation on how to install and run code associated with published
-research is another frequent barrier to replication. A study by Lapp [@Lapp_2014] found
-this impairs a researcher's ability to install and build the software
-necessary, as even small holes in the documentation were found to be
-major barriers, particularly for "novices" [Sensu @Garijo_2013] --
-where novices may be experts in nearby languages but unfamiliar with
-the package managers and other tools of the language involved. This
-same problem is discussed in [@Clark_2014].  Imprecise documentation
-goes well beyond issues of the software environment itself: incomplete
-documentation of parameters involved meant as few as 30% of analyses
-($n=34$) using the popular software STRUCTURE could be reproduced in the
-study of [@Gilbert_2012].
-
-### 3.  **Code rot**
+research is another frequent barrier to replication. A study by Lapp
+{[}16{]} found this impairs a researcher's ability to install and build
+the software necessary, as even small holes in the documentation were
+found to be major barriers, particularly for ``novices'' {[}8{]} --
+where novices may be experts in nearby languages but unfamiliar with the
+package managers and other tools of the language involved. This same
+problem is discussed in {[}3{]}. Imprecise documentation goes well
+beyond issues of the software environment itself: incomplete
+documentation of parameters involved meant as few as 30\% of analyses
+(\(n=34\)) using the popular software STRUCTURE could be reproduced in
+the study of {[}10{]}.
+
+\subsubsection{3. \textbf{Code rot}}\label{code-rot}
 
@@
 potentially change the results generated by the code. As some of these
-changes may indeed resolve valid bugs or earlier problems with underlying
-code, it will often be insufficient to demonstrate that results can be
-reproduced when using the original versions, a problem sometimes known
-as "code rot."  Researchers will want to know if the results are robust
-to the changes. The case studies in [@Lapp_2014] provide examples of
-these problems.
-
-### 4. **Barriers to adoption and reuse in existing solutions**
+changes may indeed resolve valid bugs or earlier problems with
+underlying code, it will often be insufficient to demonstrate that
+results can be reproduced when using the original versions, a problem
+sometimes known as ``code rot.'' Researchers will want to know if the
+results are robust to the changes. The case studies in {[}16{]} provide
+examples of these problems.
+
+\subsubsection{4. \textbf{Barriers to adoption and reuse in existing
+solutions}}\label{barriers-to-adoption-and-reuse-in-existing-solutions}
 
@@
 reproducibility. However, researchers face significant barriers to entry
-in learning these tools and approaches which are not part of their typical
-curriculum, or lack incentives commensurate with the effort required
-[@Joppa_2013; @FitzJohn_2014].
-
-<!-- Obviously reproducible research barriers go much deeper than this,
-code that is not modular, not documented, not functionalized or otherwise
-easy to read and follow are all obstacles to resuse.-->
-
-<!-- A wide range of proposals, I focus on those at the operating system level -->
-Though a wide variety of approaches exists to work around these challenges,
-few operate on a low enough level to provide a general solution. Clark _et al._ [@Clark_2014]
-provide an excellent description of this situation:
-
-> In scientific computing the environment was commonly managed via
-Makefiles & Unix-y hacks, or alternatively with monolithic software
+in learning these tools and approaches which are not part of their
+typical curriculum, or lack incentives commensurate with the effort
+required {[}7, 15{]}.
+
+Though a wide variety of approaches exists to work around these
+challenges, few operate on a low enough level to provide a general
+solution. Clark \emph{et al.} {[}3{]} provide an excellent description
+of this situation:
+
+\begin{quote}
+In scientific computing the environment was commonly managed via
+Makefiles \& Unix-y hacks, or alternatively with monolithic software
 like Matlab. More recently, centralized package management has provided
@@
 libraries, databases, and more.
+\end{quote}
 
@@
 
-# Current approaches
-
-<!-- FIXME Doesn't say much about the approaches. Not clear what a workflow is... -->
+\section{Current approaches}\label{current-approaches}
 
 Two dominant paradigms have emerged to address these issues so far:
-workflow software [_e.g._ @Altintas_2004; @Hull_2006]  and
-virtual machines [_e.g._ @Dudley_2010; @Howe_2012]. 
-Workflow software provides very elegant technical solutions to the
-challenges of communication between diverse software tools, capturing
-provenance in graphically driven interfaces, and handling issues from 
-versioning dependencies to data access.  Workflow solutions are often 
-built by well-funded collaborations between domain scientists and computer
+workflow software {[}1, 13{]} and virtual machines {[}5, 12{]}. Workflow
+software provides very elegant technical solutions to the challenges of
+communication between diverse software tools, capturing provenance in
+graphically driven interfaces, and handling issues from versioning
+dependencies to data access. Workflow solutions are often built by
+well-funded collaborations between domain scientists and computer
 scientists, and can be very successful in the communities within which
 they receive substantial adoption. Nonetheless, most workflow systems
-struggle with relatively low total adoption overall [@Gil_2007; @Dudley_2010].
-
-Dudley & Butte [@Dudley_2010] give several reasons that such comprehensive workflow
-systems have not been more successful:
-
-> (i) efforts are not rewarded by the current academic research and
-funding environment; (ii) commercial software vendors tend to protect
-their markets through proprietary formats and interfaces; (iii)
-investigators naturally tend to want to 'own' and control their
+struggle with relatively low total adoption overall {[}5, 9{]}.
+
+Dudley \& Butte {[}5{]} give several reasons that such comprehensive
+workflow systems have not been more successful:
+
+\begin{quote}
+\begin{enumerate}
+\def\labelenumi{(\roman{enumi})}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  efforts are not rewarded by the current academic research and funding
+  environment; (ii) commercial software vendors tend to protect their
+  markets through proprietary formats and interfaces; (iii)
+  investigators naturally tend to want to `own' and control their
   research tools; (iv) even the most generalized software will not be
   able to meet the needs of every researcher in a field; and finally (v)
-the need to derive and publish results as quickly as possible precludes
-the often slower standards-based development path.
-
-In short, workflow software expects a new approach to computational research. 
-In contrast, virtual machines (VMs) offer a more direct approach. Since
-the computer Operating System (OS) already provides the software layer
-responsible for coordinating all the different elements running on the computer,
-the VM approach captures the OS and everything running on it whole-cloth.
-To make this practical, Dudley & Butte [@Dudley_2010] and Howe [@Howe_2012] both
-propose using virtual machine images that will run on the cloud, such as Amazon's
-EC2 system, which is already based upon this kind of virtualization.
+  the need to derive and publish results as quickly as possible
+  precludes the often slower standards-based development path.
+\end{enumerate}
+\end{quote}
+
+In short, workflow software expects a new approach to computational
+research. In contrast, virtual machines (VMs) offer a more direct
+approach. Since the computer Operating System (OS) already provides the
+software layer responsible for coordinating all the different elements
+running on the computer, the VM approach captures the OS and everything
+running on it whole-cloth. To make this practical, Dudley \& Butte
+{[}5{]} and Howe {[}12{]} both propose using virtual machine images that
+will run on the cloud, such as Amazon's EC2 system, which is already
+based upon this kind of virtualization.
 
 Critics of the use of VMs to support reproducibility highlight that the
-approach is too much of a black box and thus ill suited for reproducibility [@Watson_2012].
-While the approach sidesteps the need to either install or even document the
-dependencies, this also makes it more difficult for other researchers to
-understand, evaluate, or alter those dependencies.  Moreover, other research
-cannot easily build on the virtual machine in a consistent and scalable way. 
-If each study provided its own virtual machine, any pipeline combining the 
-tools of multiple studies would quickly become impractical or impossible 
-to implement.
-
-<!-- Scripted provisioning: a solution to the weaknesses of virtual machines -->
-## A "DevOps" approach
-
-The problems highlighted here are not unique to _academic_ software,
-but impact software development in general. While the academic research
-literature has frequently focused on the development of workflow software
-dedicated to particular domains, or otherwise to the use of virtual
-machines, the software development community has recently emphasized a
-philosophy (rather than a particular tool), known as _Development_ and
-Systems _Operation_, or more frequently just "DevOps." The approach is
-characterized by scripting, rather than documenting, a description of the
-necessary dependencies for software to run, usually from the Operating
-System (OS) on up. Clark _et al._ [@Clark_2014] describe the DevOps approach along
-with both its relevance to reproducible research and examples of its
-use in the academic research context. They identify the difficulties
-I have discussed so far in terms of effective documentation:
-
-> Documentation for complex software environments is stuck between two
+approach is too much of a black box and thus ill suited for
+reproducibility {[}28{]}. While the approach sidesteps the need to
+either install or even document the dependencies, this also makes it
+more difficult for other researchers to understand, evaluate, or alter
+those dependencies. Moreover, other research cannot easily build on the
+virtual machine in a consistent and scalable way. If each study provided
+its own virtual machine, any pipeline combining the tools of multiple
+studies would quickly become impractical or impossible to implement.
+
+\subsection{A ``DevOps'' approach}\label{a-devops-approach}
+
+The problems highlighted here are not unique to \emph{academic}
+software, but impact software development in general. While the academic
+research literature has frequently focused on the development of
+workflow software dedicated to particular domains, or otherwise to the
+use of virtual machines, the software development community has recently
+emphasized a philosophy (rather than a particular tool), known as
+\emph{Development} and Systems \emph{Operation}, or more frequently just
+``DevOps.'' The approach is characterized by scripting, rather than
+documenting, a description of the necessary dependencies for software to
+run, usually from the Operating System (OS) on up. Clark \emph{et al.}
+{[}3{]} describe the DevOps approach along with both its relevance to
+reproducible research and examples of its use in the academic research
+context. They identify the difficulties I have discussed so far in terms
+of effective documentation:
+
+\begin{quote}
+Documentation for complex software environments is stuck between two
 opposing demands. To make things easier on novice users, documentation
@@
 developers like to abstract over such details.
+\end{quote}
 
 The authors contrast this to the DevOps approach, where dependency
-documentation is _scripted_:
-
-> A DevOps approach to "documenting" an application might consist of
-providing brief descriptions of various install paths, along with scripts
-or "recipes" that automate setup.
+documentation is \emph{scripted}:
+
+\begin{quote}
+A DevOps approach to ``documenting'' an application might consist of
+providing brief descriptions of various install paths, along with
+scripts or ``recipes'' that automate setup.
+\end{quote}
 
@@
 executes a script instead of manually managing the environmental setup)
-and comprehensiveness of implementation. Clark _et al._ [@Clark_2014] are careful to note
-that this is not so much a technological shift as a philosophical one:
-
-> The primary shift that's required is not one of new tooling, as most
+and comprehensiveness of implementation. Clark \emph{et al.} {[}3{]} are
+careful to note that this is not so much a technological shift as a
+philosophical one:
+
+\begin{quote}
+The primary shift that's required is not one of new tooling, as most
 developers already have the basic tooling they need. Rather, the needed
 shift is one of philosophy.
+\end{quote}
 
 Nevertheless, a growing suite of tools designed explicitly for this
-purpose have rapidly replaced the use of general purpose tools (such
-as Makefiles, bash scripts) to become synonymous with the DevOps
-philosophy.  Clark _et al._ [@Clark_2014] reviews many of these DevOps tools, their
-different roles, and their application in reproducible research.
+purpose have rapidly replaced the use of general purpose tools (such as
+Makefiles, bash scripts) to become synonymous with the DevOps
+philosophy. Clark \emph{et al.} {[}3{]} reviews many of these DevOps
+tools, their different roles, and their application in reproducible
+research.
 
 I focus the remainder of this paper on one of the most recent and
-rapidly growing among these, called Docker, and the role it can play
-in reproducible research. Docker offers several promising features
-for reproducibility that go beyond the tools highlighted in [@Clark_2014].
-Nevertheless, my goal in focusing on this technology is not to
-promote a particular solution, but to anchor the discussion of technical
-solutions to reproducibility challenges in concrete examples.
-
-<!-- Link on how to install Docker -->
-<!-- Examples here are not meant as a tutorial to Docker, but 
-rather to illustrate concepts a reader should be able to follow whether 
-or not they have any previous knowledge of Docker and its commands -->
-
-# Docker
-
-<!-- Docker is an open source (Apache Version 2) software platform written in
-the Go programming language.-->
-
-Docker is an open source project that builds on many
-long-familiar technologies from operating systems research:
-LXC containers, virtualization of the OS, and a hash-based
-or git-like versioning and differencing system, among others (see
-[docs.docker.com/faq](https://docs.docker.com/faq/#what-does-docker-add-to-just-plain-lxc)
+rapidly growing among these, called Docker, and the role it can play in
+reproducible research. Docker offers several promising features for
+reproducibility that go beyond the tools highlighted in {[}3{]}.
+Nevertheless, my goal in focusing on this technology is not to promote a
+particular solution, but to anchor the discussion of technical solutions
+to reproducibility challenges in concrete examples.
+
+\section{Docker}\label{docker}
+
+Docker is an open source project that builds on many long-familiar
+technologies from operating systems research: LXC containers,
+virtualization of the OS, and a hash-based or git-like versioning and
+differencing system, among others (see
+\href{https://docs.docker.com/faq/\#what-does-docker-add-to-just-plain-lxc}{docs.docker.com/faq}
 for an excellent overview of what Docker adds to plain LXC). The
-official documentation [docs.docker.com](https://docs.docker.com)
-already provides a thorough introduction in how to use Docker
-software; here my focus is on describing the implications this
-has for reproducible research.  Readers may also find it helpful
-to see more detailed examples of using Docker to capture, share,
-and interact with a specific computational environments. Some such
-examples, along with more detailed documentation on use can be found at
-[github.com/rocker-org](https://github.com/rocker-org/rocker).
+official documentation \href{https://docs.docker.com}{docs.docker.com}
+already provides a thorough introduction in how to use Docker software;
+here my focus is on describing the implications this has for
+reproducible research. Readers may also find it helpful to see more
+detailed examples of using Docker to capture, share, and interact with a
+specific computational environments. Some such examples, along with more
+detailed documentation on use can be found at
+\href{https://github.com/rocker-org/rocker}{github.com/rocker-org}.
 
@@
 
-### 1. Docker images: resolving 'Dependency Hell'
+\subsubsection{1. Docker images: resolving `Dependency
+Hell'}\label{docker-images-resolving-dependency-hell}
 
 A Docker based approach works similarly to a virtual machine image in
-addressing the dependency problem by providing other researchers with
-a binary image in which all the software has already been installed,
+addressing the dependency problem by providing other researchers with a
+binary image in which all the software has already been installed,
 configured and tested. (A machine image can also include all data files
-necessary for the research, which may simplify the distribution of data.)
+necessary for the research, which may simplify the distribution of
+data.)
 
 A key difference between Docker images and other virtual machines is
-that the Docker images share the Linux kernel with the host machine.
-For the end user the primary consequence of this is that any Docker
-image must be based on a Linux system with Linux-compatible software, which
-includes R, Python, Matlab, and most other scientific programming needs.[^4]
+that the Docker images share the Linux kernel with the host machine. For
+the end user the primary consequence of this is that any Docker image
+must be based on a Linux system with Linux-compatible software, which
+includes R, Python, Matlab, and most other scientific programming
+needs.\footnote{Note that re-distribution of an image in which
+  proprietary software has been installed will be subject to any
+  relevant licensing agreement.}
 
@@
 performing than complete virtual machines -- a typical desktop computer
-could run no more than a few virtual machines at once but 
-would have no trouble running 100's of Docker containers (a container
-is simply the term for running instance of an image). This feature
-has made Docker particularly attractive to industry and is largely
-responsible for the immense popularity of Docker. For our purposes this
-is a nice bonus, but the chief value to reproducible research lies
-in other aspects.
-
-[^4]: Note that re-distribution of an image in which proprietary software
-has been installed will be subject to any relevant licensing agreement.
-
-<!-- 
-- image sharing (hub)
-- image linking
--->
-
-### 2. Dockerfiles: Resolving imprecise documentation
+could run no more than a few virtual machines at once but would have no
+trouble running 100's of Docker containers (a container is simply the
+term for running instance of an image). This feature has made Docker
+particularly attractive to industry and is largely responsible for the
+immense popularity of Docker. For our purposes this is a nice bonus, but
+the chief value to reproducible research lies in other aspects.
+
+\subsubsection{2. Dockerfiles: Resolving imprecise
+documentation}\label{dockerfiles-resolving-imprecise-documentation}
 
 Though Docker images can be created interactively, this leaves little
-transparent record [^3] of what software has been installed and how.
-Dockerfiles provide a simple script (similar to a Makefile) that defines
-exactly how to build up the image, consistent with the DevOps approach
-I mentioned previously.
-
-With a syntax that is simpler than other provisioning tools (_e.g._ Chef,
-Puppet, Ansible) or Continuous Integration (CI) platforms (_e.g._ Travis CI,
-Shippable CI); users need little more than a basic familiarity with shell
-scripts and a Linux distribution software environment (_e.g._ Debian-based
-`apt-get`) to get started writing Dockerfiles.
+transparent record\footnote{The situation is in fact slightly better
+  than the virtual machine approach because these changes are versioned.
+  Docker provides tools to inspect differences (diffs) between the
+  images, and I can also roll back changes to earlier versions.} of what
+software has been installed and how. Dockerfiles provide a simple script
+(similar to a Makefile) that defines exactly how to build up the image,
+consistent with the DevOps approach I mentioned previously.
+
+With a syntax that is simpler than other provisioning tools (\emph{e.g.}
+Chef, Puppet, Ansible) or Continuous Integration (CI) platforms
+(\emph{e.g.} Travis CI, Shippable CI); users need little more than a
+basic familiarity with shell scripts and a Linux distribution software
+environment (\emph{e.g.} Debian-based \texttt{apt-get}) to get started
+writing Dockerfiles.
 
@@
 
-- While machine images can be very large (many gigabytes), a Dockerfile
+\begin{itemize}
+\item
+  While machine images can be very large (many gigabytes), a Dockerfile
   is just a small plain text file that can be easily stored and shared.
-
-- Small plain text files are ideally suited for use with a version
-management system such as `subversion` or `git`, which can track any
-changes made to the `Dockerfile`
-
-- the `Dockerfile` provides a human readable summary of the necessary
-software dependencies, environmental variables and so forth needed to
-execute the code. There is little possibility of the kind of holes or 
-imprecision in such a script that so frequently cause difficulty in
-manually implemented documentation of dependencies. This approach
-also avoids the burden of having to tediously document dependencies 
-at the end of a project, since they are instead documented as they 
-are installed by writing the `Dockerfile`.
-
-- Unlike a `Makefile` or other script, the `Dockerfile` includes all
-software dependencies down to the level of the OS, and is built by the
-Docker `build` tool, making it very unlikely that the resulting build
-will differ when being built on different machines. This is not to say
-that all builds of a Dockerfile are bitwise identical.  In particular,
-builds executed later will install more recent versions of the same
-software, if available, unless the package managers used are explicitly
-configured otherwise. I address this issue in the next section.
-
-- It is possible to add checks and tests following the commands for
+\item
+  Small plain text files are ideally suited for use with a version
+  management system such as \texttt{subversion} or \texttt{git}, which
+  can track any changes made to the \texttt{Dockerfile}
+\item
+  the \texttt{Dockerfile} provides a human readable summary of the
+  necessary software dependencies, environmental variables and so forth
+  needed to execute the code. There is little possibility of the kind of
+  holes or imprecision in such a script that so frequently cause
+  difficulty in manually implemented documentation of dependencies. This
+  approach also avoids the burden of having to tediously document
+  dependencies at the end of a project, since they are instead
+  documented as they are installed by writing the \texttt{Dockerfile}.
+\item
+  Unlike a \texttt{Makefile} or other script, the \texttt{Dockerfile}
+  includes all software dependencies down to the level of the OS, and is
+  built by the Docker \texttt{build} tool, making it very unlikely that
+  the resulting build will differ when being built on different
+  machines. This is not to say that all builds of a Dockerfile are
+  bitwise identical. In particular, builds executed later will install
+  more recent versions of the same software, if available, unless the
+  package managers used are explicitly configured otherwise. I address
+  this issue in the next section.
+\item
+  It is possible to add checks and tests following the commands for
   installing the software environment, which will verify that the setup
@@
   code-rot which I discuss next.
-
-- It is straightforward for other users to extend or customize 
-the resulting image by editing the script directly.
-
-[^3]: The situation is in fact slightly better than the virtual machine
-approach because these changes are versioned. Docker provides tools
-to inspect differences (diffs) between the images, and I can also roll
-back changes to earlier versions.
-
-### 3. Tackling code-rot with image versions
+\item
+  It is straightforward for other users to extend or customize the
+  resulting image by editing the script directly.
+\end{itemize}
+
+\subsubsection{3. Tackling code-rot with image
+versions}\label{tackling-code-rot-with-image-versions}
 
 As I have discussed above, changes to the dependencies, whether they are
-the result of security fixes, new features, or deprecation of old software,
-can break otherwise functioning code. These challenges can be significantly
-reduced because Docker defines the software environment to a particular operating system 
-and suite of libraries, such as the Ubuntu or Debian distribution.  Such 
-distributions use a staged release model with `stable`, `testing` and `unstable` 
-phases subjected to extensive testing to catch such potential problems [@Ooms_2013],
-while also providing regular security updates to software within each stage.
-Nonetheless, this cannot completely avoid the challenge of code-rot,
-particularly when it is necessary to install software that is not (yet)
-available for a given distribution.
+the result of security fixes, new features, or deprecation of old
+software, can break otherwise functioning code. These challenges can be
+significantly reduced because Docker defines the software environment to
+a particular operating system and suite of libraries, such as the Ubuntu
+or Debian distribution. Such distributions use a staged release model
+with \texttt{stable}, \texttt{testing} and \texttt{unstable} phases
+subjected to extensive testing to catch such potential problems
+{[}20{]}, while also providing regular security updates to software
+within each stage. Nonetheless, this cannot completely avoid the
+challenge of code-rot, particularly when it is necessary to install
+software that is not (yet) available for a given distribution.
 
@@
 used at the time the research was first performed. Docker provides a
-simple utility to save an image as a portable `tarball` file that can
-be read in by any other Docker installation, providing a robust way to
-run the exact versions of all software involved.  By testing both the
-`tarball` archive and the image generated by the latest Dockerfile,
-Docker provides a simple way to confirm whether or not code rot has
-effected the function of a particular piece of code. Binary Docker images
-can be efficiently shared through the Docker Hub as well, as I describe
-later.
-
-### 4. Barriers to adoption and re-use
+simple utility to save an image as a portable \texttt{tarball} file that
+can be read in by any other Docker installation, providing a robust way
+to run the exact versions of all software involved. By testing both the
+\texttt{tarball} archive and the image generated by the latest
+Dockerfile, Docker provides a simple way to confirm whether or not code
+rot has effected the function of a particular piece of code. Binary
+Docker images can be efficiently shared through the Docker Hub as well,
+as I describe later.
+
+\subsubsection{4. Barriers to adoption and
+re-use}\label{barriers-to-adoption-and-re-use}
 
@@
 to provisioning virtual machines by scripts, neither approach has seen
-widespread adoption by domain researchers, who work primarily in a
-local rather than cloud-based environment using development tools native
-to their personal operating system. To gain more widespread adoption,
-reproducible research technologies must make it easier, not harder,
-for a researcher to perform the tasks they are already doing (before 
+widespread adoption by domain researchers, who work primarily in a local
+rather than cloud-based environment using development tools native to
+their personal operating system. To gain more widespread adoption,
+reproducible research technologies must make it easier, not harder, for
+a researcher to perform the tasks they are already doing (before
 considering any additional added benefits).
 
-These issues are reflected both during the original research or development
-phase and in any subsequent reuse. Another researcher may be less likely
-to build on existing work if it can only be done by using a particular
-workflow system or monolithic software platform with which they are
-unfamiliar.  Likewise, a user is more likely to make their own computational
-environment available for reuse if it does not involve a significant
-added effort in packaging and documenting [@Stodden_2010].
-
-Though Docker is not immune to these challenges, it offers an interesting 
-example of a way forward in addressing these fundamental concerns. Here 
-I highlight these features in turn:
-
-- Integrating into local development environments
-- Modular reuse
-- Portable environments 
-- Public repository for sharing
-- Versioning
-
-## Integrating into local development environments
-
-Due in part to the difficulty of moving large VM images around a network,
-proponents of virtual machines for reproducible research often propose
-that these machines would be available exclusively as cloud computing
-environments [_e.g._ @Dudley_2010] rather than downloaded to a user's
+These issues are reflected both during the original research or
+development phase and in any subsequent reuse. Another researcher may be
+less likely to build on existing work if it can only be done by using a
+particular workflow system or monolithic software platform with which
+they are unfamiliar. Likewise, a user is more likely to make their own
+computational environment available for reuse if it does not involve a
+significant added effort in packaging and documenting {[}23{]}.
+
+Though Docker is not immune to these challenges, it offers an
+interesting example of a way forward in addressing these fundamental
+concerns. Here I highlight these features in turn:
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  Integrating into local development environments
+\item
+  Modular reuse
+\item
+  Portable environments
+\item
+  Public repository for sharing
+\item
+  Versioning
+\end{itemize}
+
+\subsection{Integrating into local development
+environments}\label{integrating-into-local-development-environments}
+
+Due in part to the difficulty of moving large VM images around a
+network, proponents of virtual machines for reproducible research often
+propose that these machines would be available exclusively as cloud
+computing environments {[}5{]} rather than downloaded to a user's
 personal computer. Though cloud computing offers some advantages such as
@@
 prefer to work with tools installed locally on their personal computer,
-at least during testing and development.  This can reduce costs, issues of
-network latency, the ability to work off-line, and will be most familiar
-to students and new developers.
+at least during testing and development. This can reduce costs, issues
+of network latency, the ability to work off-line, and will be most
+familiar to students and new developers.
 
@@
 desktop computers, as demonstrated in the approach now being pioneered
-at UC Berkeley [@Clark_2014].  This approach provides users with a pixel-identical
-environment whether working locally or on the cloud, which has particular
-advantages for student instruction [@Clark_2014]. However, it remains to be
-seen if existing researchers will be willing to forgo native applications
-for such tasks as file browsing, text editing, or version management and 
-rely exclusively on this standardized virtual environment.
+at UC Berkeley {[}3{]}. This approach provides users with a
+pixel-identical environment whether working locally or on the cloud,
+which has particular advantages for student instruction {[}3{]}.
+However, it remains to be seen if existing researchers will be willing
+to forgo native applications for such tasks as file browsing, text
+editing, or version management and rely exclusively on this standardized
+virtual environment.
 
@@
 workflow. Rather than replace a user's existing toolchain with a
-standardized virtual environment, editors and all, Docker is optimized at
-the level of single applications. A developer can thus rely on familiar
-tools while still ensuring that the execution of their code always occurs
-on the standardized container environment, thus ensuring its portability
-and reproducibility. This approach can be accomplished either by linking
-volumes or directories between the container and the host, or simply
-by instructing the Dockerfile to copy the code to the container (As we will
-see, modular reuse makes this latter strategy efficient).
-
-Rather than put the burden on the researcher to adopt
-a very different workflow, the researcher can thus use their familiar editors,
-etc., while immediately benefiting from the fact their computations can
-be easily deployed on the cloud or the machines of collaborators without
+standardized virtual environment, editors and all, Docker is optimized
+at the level of single applications. A developer can thus rely on
+familiar tools while still ensuring that the execution of their code
+always occurs on the standardized container environment, thus ensuring
+its portability and reproducibility. This approach can be accomplished
+either by linking volumes or directories between the container and the
+host, or simply by instructing the Dockerfile to copy the code to the
+container (As we will see, modular reuse makes this latter strategy
+efficient).
+
+Rather than put the burden on the researcher to adopt a very different
+workflow, the researcher can thus use their familiar editors, etc.,
+while immediately benefiting from the fact their computations can be
+easily deployed on the cloud or the machines of collaborators without
 any further effort than installing Docker software. The docker approach
 is particularly well suited for moving between local and cloud platforms
-when a web-based integrated development environment is available, such as
-RStudio Server, or (to lesser extent) an iPython notebook.
-
-On systems not already based on the Linux kernel (such
-as Mac or Windows platforms), Docker is installed (see see
-[docs.docker.com/installation](https://docs.docker.com/installation))
-through means of a very small (about
-24 MB) VM platform called `boot2docker`
-([github.com/docker/boot2docker](https://github.com/docker/boot2docker)),.
-While this poses an additional challenge for tight integration with desktop
-tools, recent advances in Docker are rapidly bridging this gap as well. 
-For example, Docker 1.3, released between drafts of this manuscript, supports
-shared volumes on Macs ([blog.docker.com](http://blog.docker.com/2014/10/docker-1-3-signed-images-process-injection-security-options-mac-shared-directories/)).
-
-## Portable computation & sharing
+when a web-based integrated development environment is available, such
+as RStudio Server, or (to lesser extent) an iPython notebook.
+
+On systems not already based on the Linux kernel (such as Mac or Windows
+platforms), Docker is installed (see see
+\href{https://docs.docker.com/installation}{docs.docker.com/installation})
+through means of a very small (about 24 MB) VM platform called
+\texttt{boot2docker}
+(\href{https://github.com/docker/boot2docker}{github.com/docker/boot2docker}),.
+While this poses an additional challenge for tight integration with
+desktop tools, recent advances in Docker are rapidly bridging this gap
+as well. For example, Docker 1.3, released between drafts of this
+manuscript, supports shared volumes on Macs
+(\href{http://blog.docker.com/2014/10/docker-1-3-signed-images-process-injection-security-options-mac-shared-directories/}{blog.docker.com}).
+
+\subsection{Portable computation \&
+sharing}\label{portable-computation-sharing}
 
 A particular advantage of this approach is that the resulting
-computational environment is immediately _portable_. LXC containers
+computational environment is immediately \emph{portable}. LXC containers
 by themselves are unlikely to run in the same way, if at all, across
-different machines, due to differences in networking, storage,
-logging and so forth. Docker handles the packaging and execution of
-a container so that it works identically across different machines,
-while exposing the necessary interfaces for networking ports, volumes,
-and so forth.  This is useful not only for the purposes of reproducible
-research, where other users may seek to reconstruct the computational
-environment necessary to run the code, but is also of immediate value
-to the researcher themselves. For instance, a researcher might want to
-execute their code on a cloud server which has more memory or processing
-power then their local machine, or would want a co-author to help debug a
+different machines, due to differences in networking, storage, logging
+and so forth. Docker handles the packaging and execution of a container
+so that it works identically across different machines, while exposing
+the necessary interfaces for networking ports, volumes, and so forth.
+This is useful not only for the purposes of reproducible research, where
+other users may seek to reconstruct the computational environment
+necessary to run the code, but is also of immediate value to the
+researcher themselves. For instance, a researcher might want to execute
+their code on a cloud server which has more memory or processing power
+then their local machine, or would want a co-author to help debug a
 particular problem. In either case, the researcher can export a snapshot
@@
 
-```bash
-docker export container-name > container.tar
-```
-
-and then run this identical environment on the cloud or collaborators' machine.
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{docker} \NormalTok{export container-name }\KeywordTok{>} \NormalTok{container.tar}
+\end{Highlighting}
+\end{Shaded}
+
+and then run this identical environment on the cloud or collaborators'
+machine.
 
 Sharing these images is further facilitated by the Docker Hub technology
-([hub.docker.com](http://hub.docker.com)).  While Docker images tend
-to be much smaller than equivalent virtual machines, moving around
-even 100's of gigabytes can be a challenge.  The Docker Hub provides a
+(\href{http://hub.docker.com}{hub.docker.com}). While Docker images tend
+to be much smaller than equivalent virtual machines, moving around even
+100's of gigabytes can be a challenge. The Docker Hub provides a
 convenient distribution service, freely storing the pre-built images,
@@
 can run their own private versions of the Hub on their own servers, for
-instance, if security of the data or the longevity of the public platform
-is a concern.  Docker also supports *Automated Builds* through the Docker
-Hub.  This acts as a kind of Continuous Integration (CI) service that
-verifies the image builds correctly whenever the Dockerfile is updated,
-particularly if the Dockerfile includes checks for the environment.
+instance, if security of the data or the longevity of the public
+platform is a concern. Docker also supports \emph{Automated Builds}
+through the Docker Hub. This acts as a kind of Continuous Integration
+(CI) service that verifies the image builds correctly whenever the
+Dockerfile is updated, particularly if the Dockerfile includes checks
+for the environment.
 
 One can share a public copy of the image just created by using the
-`Docker push` command, followed by the name of the image using the command:
-
-```bash
-docker push username/r-recommended
-```
+\texttt{docker push} command, followed by the name of the image using
+the command:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{docker} \NormalTok{push username/r-recommended}
+\end{Highlighting}
+\end{Shaded}
 
 If a Dockerfile is made available on a public code repository such as
-[Github](https://github.com) or [Bitbucket](https://bitbucket.org),
-the Hub can automatically build the image whenever a change is made to
-the Dockerfile, making the `push` command unnecessary. A user can update
-their local image using the `Docker pull <imagename>`, which downloads
-any changes that have since been made to the copy of the image on the Hub.
-
-## Re-usable modules
-
-The approach of Docker offers a technical
-solution to what is frequently seen as the primary weakness of the
-standard VM approach to reproducibility - reusing and
-remixing elements.  To some extent this is already addressed by the
-DevOps approach of Dockerfiles, providing a scripted description
-of the environment that can be tweaked and altered, but also includes
-something much more fundamental to Docker.
-
-The challenge to reusing VMs can be summarized as "you
-can't install an image for every pipeline you want..." [@Watson_2012]. 
-While providing a VM may make it easy for other researchers to run
-a particular piece of software, it becomes very difficult to combine
-multiple software components in future research if each must run inside
-it's own VM. The lack of reusable, scalable modules in the VM model
-poses a major barrier to future reuse.  In contrast, (as the analogy to 
-shipping containers in the name might imply) Docker containers
-are optimized for this kind of stacking and modular reuse.
-There are at least three ways in which Docker supports this kind of
-extensibility.
+\href{https://github.com}{Github} or
+\href{https://bitbucket.org}{Bitbucket}, the Hub can automatically build
+the image whenever a change is made to the Dockerfile, making the
+\texttt{push} command unnecessary. A user can update their local image
+using the \texttt{docker pull \textless{}imagename\textgreater{}}, which
+downloads any changes that have since been made to the copy of the image
+on the Hub.
+
+\subsection{Re-usable modules}\label{re-usable-modules}
+
+The approach of Docker offers a technical solution to what is frequently
+seen as the primary weakness of the standard VM approach to
+reproducibility - reusing and remixing elements. To some extent this is
+already addressed by the DevOps approach of Dockerfiles, providing a
+scripted description of the environment that can be tweaked and altered,
+but also includes something much more fundamental to Docker.
+
+The challenge to reusing VMs can be summarized as ``you can't install an
+image for every pipeline you want\ldots{}'' {[}28{]}. While providing a
+VM may make it easy for other researchers to run a particular piece of
+software, it becomes very difficult to combine multiple software
+components in future research if each must run inside it's own VM. The
+lack of reusable, scalable modules in the VM model poses a major barrier
+to future reuse. In contrast, (as the analogy to shipping containers in
+the name might imply) Docker containers are optimized for this kind of
+stacking and modular reuse. There are at least three ways in which
+Docker supports this kind of extensibility.
 
 Most primitively, because Dockerfile itself provides a script for
-creating the computational environment, future researchers can
-extend or modify the resulting machine image by simply editing the
-Dockerfile. This same approach is also possible for VM approaches that rely on 
-DevOps tools [@Clark_2014]. In contrast to VMs, however, Docker is also
+creating the computational environment, future researchers can extend or
+modify the resulting machine image by simply editing the Dockerfile.
+This same approach is also possible for VM approaches that rely on
+DevOps tools {[}3{]}. In contrast to VMs, however, Docker is also
 modular by design: both in how Dockerfiles are defined and how Docker
@@
 First, Docker facilitates modular reuse by building one container on top
-of another through the use of `FROM` directive in Dockerfiles. This acts
-like a software dependency; but unlike other software, a Dockerfile must
-have exactly one dependency (one `FROM` line). Particular version of
-the dependency can be specified using the `:` notation, or omitted to
-default to the latest version. To install software, Dockerfiles leverage
-existing package managers on common Linux distributions such as `apt`
-on Debian. These also permit installing either specific or only the
-latest version.
-
-<!--
-Consider the following Dockerfile to illustrate this modular reuse. This
-is the complete Dockerfile required to create the computational environment
-required for a particular manuscript:
-
-```bash
-FROM rocker/ropensci:latest
-MAINTAINER Carl Boettiger cboettig@ropensci.org
-RUN git clone https://github.com/ropensci/RNeXML.git /rnexml
-RUN Rscript -e 'devtools::install("/rnexml")'
-```
-
-The first line contains the FROM directive, indicating that this Dockerfile
-builds on the existing Docker image `rocker/ropensci`, hosted on the Docker
-Hub.  For this approach to be effective for reproducible one must take care 
-that the base image and it's Dockerfile recipe are available from a persistant
-archive.
-
-The second line is optional, declaring minimal metadata regarding the 
-contact information for the image maintainer. The third line copys the
-project onto the image and the fourth installs the necessary software that
-has not been provided by the base image. Projects using less standard
-environments may require more instructions.  
--->
-
-While a Dockerfile can have only a single `FROM` line, sometimes it may
-be necessary to build on the computational environment provided by more
-than one container. To address this, Docker defines a syntax for linking
-multiple containers together.  This allows each container to act as a
-building block providing just what is necessary to run one particular
-service or element, and exposing just what is needed to link it together
-with other blocks.  For instance, one could have one container running
-a PostgreSQL database which serves data to another container running a
-python environment to analyze the data:
-
-<!--
-```bash
-docker run --name texlive -v /usr/local/texlive rocker/texlive
-docker run -dP --volumes-from texlive \
--e PATH=$PATH:/usr/local/texlive/2014/bin/x86_64-linux/ \
-rocker/rstudio 
-```
--->
-
-```bash
-docker run -d --name db training/postgres
-docker run -d -P --link db:db training/webapp python app.py
-```
-This separates the task for running the database (first line) from
-the application used to analyze the data (second line), allowing a more
-modular approach to reuse: another researcher could use the same database
-container while connecting it to different scripts.
+of another through the use of \texttt{FROM} directive in Dockerfiles.
+This acts like a software dependency; but unlike other software, a
+Dockerfile must have exactly one dependency (one \texttt{FROM} line).
+Particular version of the dependency can be specified using the
+\texttt{:} notation, or omitted to default to the latest version. To
+install software, Dockerfiles leverage existing package managers on
+common Linux distributions such as \texttt{apt} on Debian. These also
+permit installing either specific or only the latest version.
+
+While a Dockerfile can have only a single \texttt{FROM} line, sometimes
+it may be necessary to build on the computational environment provided
+by more than one container. To address this, Docker defines a syntax for
+linking multiple containers together. This allows each container to act
+as a building block providing just what is necessary to run one
+particular service or element, and exposing just what is needed to link
+it together with other blocks. For instance, one could have one
+container running a PostgreSQL database which serves data to another
+container running a python environment to analyze the data:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{docker} \NormalTok{run -d --name db training/postgres}
+\KeywordTok{docker} \NormalTok{run -d -P --link db:db training/webapp \textbackslash{}}
+  \KeywordTok{python} \NormalTok{app.py}
+\end{Highlighting}
+\end{Shaded}
+
+This separates the task for running the database (first line) from the
+application used to analyze the data (second line), allowing a more
+modular approach to reuse: another researcher could use the same
+database container while connecting it to different scripts.
 
@@
 computer can easily run 100s of such services each in their own
-container. A rapidly expanding ecosystem of software around Docker,
-such as fig ([github.com/docker/fig](https://github.com/docker/fig))
+container. A rapidly expanding ecosystem of software around Docker, such
+as fig (\href{https://github.com/docker/fig}{github.com/docker/fig})
 facilitates the complexity of running multiple containers. This feature
@@
 
-<!-- Transition -->
-
-## Versioning
-
-In addition to version managing the Dockerfile, the images themselves are
-versioned using a git-like hash system (_e.g._ see `Docker commit`, `docker
-push`/`Docker pull`, `docker history`, `docker diff`).  Docker images and
+\subsection{Versioning}\label{versioning}
+
+In addition to version managing the Dockerfile, the images themselves
+are versioned using a git-like hash system (\emph{e.g.} see
+\texttt{docker commit}, \texttt{docker push}/\texttt{docker pull},
+\texttt{docker history}, \texttt{docker diff}). Docker images and
 containers have dedicated metadata specifying the date, author, parent
-image, and other details (see `Docker inspect`). One can roll back an
-image through the layers of history of its construction, then build off
-an earlier layer, or roll back changes made interactively in a
+image, and other details (see \texttt{docker inspect}). One can roll
+back an image through the layers of history of its construction, then
+build off an earlier layer, or roll back changes made interactively in a
 container. For instance, here I inspect recent changes made to the
-`ubuntu:14.04` image:
-
-```bash
-docker history ubuntu:14.04
-```
-
-One can identify an earlier version, and roll back to that version 
-just by adjusting the Docker tag to match the hash of that version. 
-For instance:
-
-```bash
-docker tag 25f ubuntu:14.04
-```
-If one now inspects the history, which shows that it now begins from this earlier point:
-
-```bash
-docker history ubuntu:14.04
-```
-
-This same feature also means that Docker can perform incremental
-uploads and downloads that send only the differences between images,
-(just like `git push` or `git pull` for git repositories), rather than
-transfer the full image each time.
-
-<!-- Step back, big picture -->
-
-# Conclusions
-
-## Best Practices
+\texttt{ubuntu:14.04} image:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{docker} \NormalTok{history ubuntu:14.04}
+\end{Highlighting}
+\end{Shaded}
+
+One can identify an earlier version, and roll back to that version just
+by adjusting the Docker tag to match the hash of that version. For
+instance:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{docker} \NormalTok{tag 25f ubuntu:14.04}
+\end{Highlighting}
+\end{Shaded}
+
+If one now inspects the history, which shows that it now begins from
+this earlier point:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{docker} \NormalTok{history ubuntu:14.04}
+\end{Highlighting}
+\end{Shaded}
+
+This same feature also means that Docker can perform incremental uploads
+and downloads that send only the differences between images, (just like
+\texttt{git push} or \texttt{git pull} for git repositories), rather
+than transfer the full image each time.
+
+\section{Conclusions}\label{conclusions}
+
+\subsection{Best Practices}\label{best-practices}
 
@@
 nonetheless depends on how each of these features are adopted and
-implemented by individual researchers. I summarize a few of these practices here:
-
-- _Use Docker containers during development_. A key feature of the Docker
-approach is the ability to mimic as closely as possible the current
-workflow and development practices of the user.  Code executing inside
-a container on a local machine can appear identical to code running
-natively, but with the added benefit that one can simply recreate or
-snapshot and share the entire computational environment with a few simple
-commands.  This works best if researchers set up their computational
-environment in a container from the outset of the project.
-
-- _Write Dockerfiles instead of installing interactive sessions_. As we
-have noted already, Docker can be used in a purely interactive manner to
-record and distribute changes to a computational environment.  However,
-the approach is most useful for reproducible research when researchers
-begin by defining their environment explicitly in the DevOps fashion by
-writing a Dockerfile.
-
-- _Adding tests or checks to the Dockerfile_. Dockerfile commands need
-not be limited to installing software, but can also include execution.
-This can help verify that an image has build successfully with all the
-software necessary to run the research code of interest.
-
-- _Use and provide appropriate base images_. Though Docker supports
+implemented by individual researchers. I summarize a few of these
+practices here:
+
+\begin{itemize}
+\item
+  \emph{Use Docker containers during development}. A key feature of the
+  Docker approach is the ability to mimic as closely as possible the
+  current workflow and development practices of the user. Code executing
+  inside a container on a local machine can appear identical to code
+  running natively, but with the added benefit that one can simply
+  recreate or snapshot and share the entire computational environment
+  with a few simple commands. This works best if researchers set up
+  their computational environment in a container from the outset of the
+  project.
+\item
+  \emph{Write Dockerfiles instead of installing interactive sessions}.
+  As we have noted already, Docker can be used in a purely interactive
+  manner to record and distribute changes to a computational
+  environment. However, the approach is most useful for reproducible
+  research when researchers begin by defining their environment
+  explicitly in the DevOps fashion by writing a Dockerfile.
+\item
+  \emph{Adding tests or checks to the Dockerfile}. Dockerfile commands
+  need not be limited to installing software, but can also include
+  execution. This can help verify that an image has build successfully
+  with all the software necessary to run the research code of interest.
+\item
+  \emph{Use and provide appropriate base images}. Though Docker supports
   modular design, it remains up to the researchers to take advantage of
   it. An appropriate workflow might involve one Dockerfile that includes
-all the software dependencies a researcher usually uses in the course of
-their development, which can then be extended by separate Docker images
-for particular projects.  Re-using existing images reduces the effort
-required to set up an environment, contributes to the standardization
-of computational environments within a field, and best leverages the
-ability of Docker's distribution system to download only differences.
-
-- _Share Docker images and Dockerfiles_. The Docker Hub significantly
-reduces the barriers for making even large images (which can exceed the
-file size limits of journals common scientific data repositories such
-as Dryad and Figshare) readily available to other researchers.
-
-- _Archive tarball snapshots_. Despite similar semantics to git, Docker's
-versioning system works rather differently than version management
-of code.  Docker can roll back layers[^5] that have been added to
-an image, but not revert to the earlier state of a particular layer.
-In consequence, to preserve a bitwise identical snapshot of a container
-used to generate a given set of results, it is necessary to archive the
-image tarball itself -- one can not simply rely on the Docker history
-to recover an earlier state.
-
-[^5]: Technically AUFS (advanced multi layered unification filesystem )
-layers, see [wikipedia.org/wiki/aufs](http://en.wikipedia.org/wiki/aufs)
-
-## Limitations and future developments
+  all the software dependencies a researcher usually uses in the course
+  of their development, which can then be extended by separate Docker
+  images for particular projects. Re-using existing images reduces the
+  effort required to set up an environment, contributes to the
+  standardization of computational environments within a field, and best
+  leverages the ability of Docker's distribution system to download only
+  differences.
+\item
+  \emph{Share Docker images and Dockerfiles}. The Docker Hub
+  significantly reduces the barriers for making even large images (which
+  can exceed the file size limits of journals common scientific data
+  repositories such as Dryad and Figshare) readily available to other
+  researchers.
+\item
+  \emph{Archive tarball snapshots}. Despite similar semantics to git,
+  Docker's versioning system works rather differently than version
+  management of code. Docker can roll back layers\footnote{Technically
+    AUFS (advanced multi layered unification filesystem ) layers, see
+    \href{http://en.wikipedia.org/wiki/aufs}{wikipedia.org/wiki/aufs}}
+  that have been added to an image, but not revert to the earlier state
+  of a particular layer. In consequence, to preserve a bitwise identical
+  snapshot of a container used to generate a given set of results, it is
+  necessary to archive the image tarball itself -- one can not simply
+  rely on the Docker history to recover an earlier state.
+\end{itemize}
+
+\subsection{Limitations and future
+developments}\label{limitations-and-future-developments}
 
@@
 containers, and simple interface have proven successful in industry and
-could have promising implications for reproducible research in scientific
-communities. Nonetheless, these advances raise questions and challenges 
-of their own.
-
-- Docker does not provide complete virtualization but relies on the 
-Linux kernel provided by the host.  Systems research can provide insight
-on what limitations to reproducibility this introduces [_e.g._  @Harji_2013].
-
-- Docker is limited to 64 bit host machines, making it impossible to run on
-older hardware (at this time).
-
-- On Mac and Windows machines Docker must still be run in a fully
+could have promising implications for reproducible research in
+scientific communities. Nonetheless, these advances raise questions and
+challenges of their own.
+
+\begin{itemize}
+\item
+  Docker does not provide complete virtualization but relies on the
+  Linux kernel provided by the host. Systems research can provide
+  insight on what limitations to reproducibility this introduces
+  {[}11{]}.
+\item
+  Docker is limited to 64 bit host machines, making it impossible to run
+  on older hardware (at this time).
+\item
+  On Mac and Windows machines Docker must still be run in a fully
   virtualized environment. Though the boot2docker tool streamlines this
@@
   adoption by users on of these systems.
-
-- Potential computer security issues may still need to be evaluated. Among other
-changes, future support for digitally signing Docker images may make it easier
-to build off of only trusted binaries.
-
-- Most importantly, it remains to be seen if Docker will be significantly
-adopted by any scientific research or teaching community.
-
-## Further considerations
-
-<!-- Is this really necessary? Can it be more consise? -->
-
-### Combining virtualization with other reproducible-research tools
-
-<!-- Package managers -->
+\item
+  Potential computer security issues may still need to be evaluated.
+  Among other changes, future support for digitally signing Docker
+  images may make it easier to build off of only trusted binaries.
+\item
+  Most importantly, it remains to be seen if Docker will be
+  significantly adopted by any scientific research or teaching
+  community.
+\end{itemize}
+
+\subsection{Further considerations}\label{further-considerations}
+
+\subsubsection{Combining virtualization with other reproducible-research
+tools}\label{combining-virtualization-with-other-reproducible-research-tools}
 
@@
 to, other technical tools for ensuring computational reproducibility.
-Existing tools for managing dependencies for a particular language [_e.g._ R
-packages, Ruby gems, etc, see @Ooms_2014] can easily be employed within a
-Docker-based approach, allowing the operating-systems level virtualization
-to sidestep potential issues such as external library dependencies or
-conflicts with existing user libraries.  Other approaches that facilitate
-reproducible research also introduce additional software dependencies
-and possible points of failure [@FitzJohn_2014]. One example includes
-dynamic documents [_e.g._ @Leisch_2002; @Peng_2011; @Xie_2013] which embed
-the code required to re-generate the results within the manuscript.  As a
-result, it is necessary to package the appropriate typesetting libraries
-(_e.g._ \LaTeX) along with the code libraries such that the document executes
-successfully for different researchers and platforms.
-
-### Impacting cultural norms?
-
-I noted at the outset that cultural expectations responsible for a
-lack of code sharing practices in many fields are a far more extensive
-primary barrier to reproducibility than the technical barriers
-discussed here. Nevertheless, it may be worth considering how solutions
-to these technical barriers can influence the cultural landscape as well.
-Many researchers may be reluctant to publish code today because they
-fear a it will be primarily a one-way street: more technical savvy
-researchers than themselves can benefit from their hard work, while
-they may not benefit from the work produced by others.  Lowering the
-technical barriers to reuse provides immediate practical benefits
-that make this exchange into a more balanced, two-way street. Another
-concern is that the difficulty imposed in preparing code to be shared,
-such as providing even semi-adequate documentation or support for other
-users to be able to install and run it in the first place is too high
-[@Stodden_2010]. Thus, lowering these barriers to re-use through
-the appropriate infrastructure may also reduce certain cultural
-barriers to sharing.
-
-# Acknowledgements
+Existing tools for managing dependencies for a particular language
+{[}21{]} can easily be employed within a Docker-based approach, allowing
+the operating-systems level virtualization to sidestep potential issues
+such as external library dependencies or conflicts with existing user
+libraries. Other approaches that facilitate reproducible research also
+introduce additional software dependencies and possible points of
+failure {[}7{]}. One example includes dynamic documents {[}17, 22, 26{]}
+which embed the code required to re-generate the results within the
+manuscript. As a result, it is necessary to package the appropriate
+typesetting libraries (\emph{e.g.} \LaTeX) along with the code libraries
+such that the document executes successfully for different researchers
+and platforms.
+
+\subsubsection{Impacting cultural
+norms?}\label{impacting-cultural-norms}
+
+I noted at the outset that cultural expectations responsible for a lack
+of code sharing practices in many fields are a far more extensive
+primary barrier to reproducibility than the technical barriers discussed
+here. Nevertheless, it may be worth considering how solutions to these
+technical barriers can influence the cultural landscape as well. Many
+researchers may be reluctant to publish code today because they fear a
+it will be primarily a one-way street: more technical savvy researchers
+than themselves can benefit from their hard work, while they may not
+benefit from the work produced by others. Lowering the technical
+barriers to reuse provides immediate practical benefits that make this
+exchange into a more balanced, two-way street. Another concern is that
+the difficulty imposed in preparing code to be shared, such as providing
+even semi-adequate documentation or support for other users to be able
+to install and run it in the first place is too high {[}23{]}. Thus,
+lowering these barriers to re-use through the appropriate infrastructure
+may also reduce certain cultural barriers to sharing.
+
+\section{Acknowledgements}\label{acknowledgements}
 
 CB acknowledges support from NSF grant DBI-1306697, and also from the
-Sloan Foundation support through the rOpenSci project. CB also wishes
-to thank Scott Chamberlain, Dirk Eddelbuettel, Rich FitzJohn, Yihui Xie,
+Sloan Foundation support through the rOpenSci project. CB also wishes to
+thank Scott Chamberlain, Dirk Eddelbuettel, Rich FitzJohn, Yihui Xie,
 Titus Brown, John Stanton-Geddes and many others for helpful discussions
@@
 
-# References
-
-<!-- These links don't appear in either pdf or rendered version, but provide some automatic linking on the Github rendering 
-It would be nice if pandoc's reference syntax were closer to markdown's standard link syntax (e.g. always use [], bracket each reference)
--->
-
-[@Lapp_2014]: https://storify.com/hlapp/reproducibility-repeatability-bigthink
-[@Gilbert_2012]: 10.1111/j.1365-294X.2012.05754.x
-[@Collberg_2014]: http://reproducibility.cs.arizona.edu/v1/tr.pdf
-[@Brown_2014]: http://cs.brown.edu/~sk/Memos/Examining-Reproducibility/
-[@Vandewalle_2009]: http://doi.org/10.1109/MSP.2009.932122
-[@Merali_2010]: http://doi.org/10.1038/467775a "Computational Science ... error: why scientific programming does not compute"
-[@Gil_2007]: http://doi.org/10.1109/MC.2007.421 "Examining the Challenges of Scientific Workflows"
-[@Hull_2006]: http://doi.org/10.1093/nar/gkl320 "Taverna: a tool for building and running workflows of services"
-[@Altintas_2004]: http://doi.org/10.1109/SSDM.2004.1311241 "Kepler: an extensible system for design and execution of scientific workflows"
-[@Howe_2012]: http://doi.org/10.1109/MCSE.2012.62
-[@Dudley_2010]: http://doi.org/10.1038/nbt1110-1181 
-[@Clark_2014]: https://berkeley.app.box.com/s/w424gdjot3tgksidyyfl
-[@Ooms_2013]: http://arxiv.org/abs/1303.2140v2 "Possible Directions for Improving Dependency Versioning in R"
-[@Ooms_2014]: http://arxiv.org/abs/1406.4806 "The OpenCPU System: Towards a Universal Interface for Scientific Computing through Separation of Concerns"
-[@Peng_2011]: http://doi.org/10.1126/science.1213847
-[@Barnes_2010]: http://doi.org/10.1038/467753a
-[@Ince_2012]: http://doi.org/10.1038/nature10836
-[@Stodden_2014]: http://doi.org/10.3233/SJI-140818
-[@Stodden_2010]: http://doi.org/10.2139/ssrn.1550193
-[@Stodden_2013]: http://www.davidhbailey.com/dhbpapers/icerm-report.pdf "ICERM Working group report: Setting the Default to Reproducible: Reproducibility in Computational and Experimental Mathematics"
-[@Nature_2012]: http://doi.org/10.1038/483509a
-[@The_Economist_2013]: http://www.economist.com/news/leaders/21588069-scientific-research-has-changed-world-now-it-needs-change-itself-how-science-goes-wrong
-[@van_der_Meulen_2008]: http://doi.org/10.1109/TSE.2008.70 "The effectiveness of software diversity in a large population of programs."
-[@Watson_2012]: https://twitter.com/BioMickWatson/status/265037994526928896
-
-```{r include=FALSE, eval=FALSE}
-library("knitcitations")
-options(citation_format = "pandoc")
-citep(c(
-"10.1371/journal.pone.0080278",
-"10.1126/science.1213847",
-"10.1038/467753a",
-"10.1038/nature10836",
-"10.1038/483509a",
-"10.1038/467775a",
-"10.1111/j.1365-294X.2012.05754.x",
-"10.1109/MSP.2009.932122",
-"10.1109/MC.2007.421",
-"10.1093/nar/gkl320",
-"10.1109/SSDM.2004.1311241",
-"10.1109/MCSE.2012.62",
-"10.1038/nbt1110-1181"))
-citep("http://doi.org/10.3233/SJI-140818")
-citep(c(
-"10.2139/ssrn.1550193",
-"10.1109/TSE.2008.70",
-"http://arxiv.org/abs/1303.2140v2",
-"http://arxiv.org/abs/1406.4806",
-"https://storify.com/hlapp/reproducibility-repeatability-bigthink",
-"http://cs.brown.edu/~sk/Memos/Examining-Reproducibility/",
-"https://berkeley.app.box.com/s/w424gdjot3tgksidyyfl",
-"http://reproducibility.cs.arizona.edu/v1/tr.pdf",
-"http://www.davidhbailey.com/dhbpapers/icerm-report.pdf",
-"http://www.economist.com/news/leaders/21588069-scientific-research-has-changed-world-now-it-needs-change-itself-how-science-goes-wrong",
-"https://twitter.com/BioMickWatson/status/265037994526928896"))
-
-write.bibtex(file="components/references.bib")
-
-```
-
-<!-- Covered elsewhere in a more integrated fashion -->
-<!--
-
-Researchers familiar with LXC may be tempted to view the Docker
-system as little more than a creative application and marketing of an
-existing technology. However, the Docker approach adds several features
-crucial for its use in reproducible research (adapted from [Docker
-FAQ](https://docs.Docker.com/faq/#what-does-docker-add-to-just-plain-lxc)):
-
-- *Portable deployment*: Docker defines a format for bundling and
-exchanging containers such that they run identically across platforms.
-
-- *Automatic builds*: Docker defines a format (Dockerfiles) that provide
-a simple, scripted way to construct containers, following the DevOps
-philosophy.
-
-- *Versioning*: Docker provides git-like capibilities for tracking
-changes to a container, inspecting the history, metadata, or diff between
-versions, committing versions, rolling back changes, and incremential
-uploads and downloads that send only the differences (like `git push`
-or `git pull`).
-
-- *Component reuse*: Any container can be used as the base for any other.
-
-- *Sharing*: Docker provides a public registry (Docker Hub) where any
-user can make images avaialble to any other. (The registry itself is
-also open source so that it can be deployed privately).
-
-Each of these elements have clear implications for reproducible research.
-
--->
+\section{References}\label{references}
+
+{[}1{]}Altintas, I. et al. 2004. Kepler: an extensible system for design
+and execution of scientific workflows. \emph{Proceedings. 16th
+international conference on scientific and statistical database
+management, 2004.} (2004).
+
+{[}2{]}Barnes, N. 2010. Publish your computer code: it is good enough.
+\emph{Nature}. 467, 7317 (Oct. 2010), 753--753.
+
+{[}3{]}Clark, D. et al. 2014. BCE: Berkeley's Common Scientific Compute
+Environment for Research and Education. \emph{Proceedings of the 13th
+Python in Science Conference (SciPy 2014)}. (2014).
+
+{[}4{]}Collberg, C. et al. 2014. \emph{Measuring Reproducibility in
+Computer Systems Research}.
+
+{[}5{]}Dudley, J.T. and Butte, A.J. 2010. In silico research in the era
+of cloud computing. \emph{Nat Biotechnol}. 28, 11 (Nov. 2010),
+1181--1185.
+
+{[}6{]}Eide, E. 2010. Toward Replayable Research in Networking and
+Systems. \emph{Archive '10, the nSF workshop on archiving experiments to
+raise scientific standards} (2010).
+
+{[}7{]}FitzJohn, R. et al. 2014. Reproducible research is still a
+challenge. \url{http://ropensci.org/blog/2014/06/09/reproducibility/}.
+
+{[}8{]}Garijo, D. et al. 2013. Quantifying reproducibility in
+computational biology: The case of the tuberculosis drugome.
+\emph{\(\lbrace\)PLoS\(\rbrace\) \(\lbrace\)ONE\(\rbrace\)}. 8, 11 (Nov.
+2013), e80278.
+
+{[}9{]}Gil, Y. et al. 2007. Examining the challenges of scientific
+workflows. \emph{Computer}. 40, 12 (2007), 24--32.
+
+{[}10{]}Gilbert, K.J. et al. 2012. Recommendations for utilizing and
+reporting population genetic analyses: the reproducibility of genetic
+clustering using the program structure. \emph{Mol Ecol}. 21, 20 (Sep.
+2012), 4925--4930.
+
+{[}11{]}Harji, A.S. et al. 2013. Our Troubles with Linux Kernel Upgrades
+and Why You Should Care. \emph{ACM SIGOPS Operating Systems Review}. 47,
+2 (2013), 66--72.
+
+{[}12{]}Howe, B. 2012. Virtual appliances, cloud computing, and
+reproducible research. \emph{Computing in Science \& Engineering}. 14, 4
+(Jul. 2012), 36--41.
+
+{[}13{]}Hull, D. et al. 2006. Taverna: a tool for building and running
+workflows of services. \emph{Nucleic Acids Research}. 34, Web Server
+(Jul. 2006), W729--W732.
+
+{[}14{]}Ince, D.C. et al. 2012. The case for open computer programs.
+\emph{Nature}. 482, 7386 (Feb. 2012), 485--488.
+
+{[}15{]}Joppa, L.N. et al. 2013. Troubling Trends in Scientific Software
+Use. \emph{Science (New York, N.Y.)}. 340, 6134 (May 2013), 814--815.
+
+{[}16{]}Lapp, Hilmar 2014. Reproducibility / repeatability bigThink
+(with tweets) @hlapp. \emph{Storify}.
+\url{http://storify.com/hlapp/reproducibility-repeatability-bigthink}.
+
+{[}17{]}Leisch, F. 2002. Sweave: Dynamic Generation of Statistical
+Reports Using Literate Data Analysis. \emph{Compstat}. W. Härdle and B.
+Rönz, eds. Physica-Verlag HD.
+
+{[}18{]}Merali, Z. 2010. Computational science: ...Error. \emph{Nature}.
+467, 7317 (Oct. 2010), 775--777.
+
+{[}19{]}Nature Editors 2012. Must try harder. \emph{Nature}. 483, 7391
+(Mar. 2012), 509--509.
+
+{[}20{]}Ooms, J. 2013. Possible directions for improving dependency
+versioning in r. \emph{arXiv.org}.
+\url{http://arxiv.org/abs/1303.2140v2}.
+
+{[}21{]}Ooms, J. 2014. The openCPU system: Towards a universal interface
+for scientific computing through separation of concerns.
+\emph{arXiv.org}. \url{http://arxiv.org/abs/1406.4806}.
+
+{[}22{]}Peng, R.D. 2011. Reproducible research in computational science.
+\emph{Science}. 334, 6060 (Dec. 2011), 1226--1227.
+
+{[}23{]}Stodden, V. 2010. The scientific method in practice:
+Reproducibility in the computational sciences. \emph{SSRN Journal}.
+(2010).
+
+{[}24{]}Stodden, V. et al. 2013. Setting the Default to Reproducible.
+(2013), 1--19.
+
+{[}25{]}The Economist 2013. How science goes wrong. \emph{The
+Economist}.
+\url{http://www.economist.com/news/leaders/21588069-scientific-research-has-changed-world-now-it-needs-change-itself-how-science-goes-wrong}.
+
+{[}26{]}Xie, Y. 2013. \emph{Dynamic documents with R and knitr}.
+Chapman; Hall/CRC.
+
+{[}27{]}2014. Examining reproducibility in computer science.
+\url{http://cs.brown.edu/~sk/Memos/Examining-Reproducibility/}.
+
+{[}28{]}2012. Mick Watson on Twitter: @ewanbirney @pathogenomenick
+@ctitusbrown you can't install an image for every pipeline you want...
+\url{https://twitter.com/BioMickWatson/status/265037994526928896}.
+
+\end{document}
